---
title: "Multi-Assembly Comparison Report"
params:
  comparison_tsv: "__COMPARISON_TSV__"
  completeness_tsv: "__COMPLETENESS_TSV__"
  ref_lengths_tsv: "__REF_LENGTHS_TSV__"
  per_assembly_summary_tsvs: "__PER_ASM_TSVS__"
  assembly_names: "__ASM_NAMES__"
  per_assembly_macro_blocks_tsvs: "__PER_ASM_MACRO_TSVS__"
  per_assembly_rdna_arrays_tsvs: "__PER_ASM_RDNA_TSVS__"
  per_assembly_contaminants_tsvs: "__PER_ASM_CONTAM_TSVS__"
  pairwise_macro_blocks_tsvs: "__PAIRWISE_MACRO_TSVS__"
  pairwise_names: "__PAIRWISE_NAMES__"
  reference_name: "__REFNAME__"
  synteny_mode: "__SYNTENY_MODE__"
  chr_like_minlen: "__CHRLIKE__"
  out_prefix: "__OUTPREFIX__"
output:
  html_document:
    self_contained: true
    css: "__COMMON_CSS__"
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, dpi = 150)

source("__COMMON_R__")
pacman::p_load(ggridges)

# ---------------------------------------------------------------------------
# Comparison-report-specific helpers
# ---------------------------------------------------------------------------
fmt_pct <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.1f%%", x * 100))

# Assembly color palette (one color per assembly)
asm_pal <- c(oi[5], oi[1], oi[3], oi[6], oi[7], oi[4], oi[2], oi[8], oi[9])

# ---------------------------------------------------------------------------
# Subgenome detection and color setup
# ---------------------------------------------------------------------------
ref_lengths <- read_tsv(params$ref_lengths_tsv, show_col_types = FALSE,
                        col_types = cols(subgenome = col_character()))
list2env(setup_subgenomes(ref_lengths), envir = environment())

# ---------------------------------------------------------------------------
# Load data
# ---------------------------------------------------------------------------
comp_df <- read_tsv(params$comparison_tsv, show_col_types = FALSE)
compl_df <- read_tsv(params$completeness_tsv, show_col_types = FALSE)

n_asm <- nrow(comp_df)
asm_names <- comp_df$assembly
asm_colors <- setNames(rep_len(asm_pal, n_asm), asm_names)

# Load per-assembly summary TSVs
asm_tsv_paths <- strsplit(params$per_assembly_summary_tsvs, ";")[[1]]
asm_name_list <- strsplit(params$assembly_names, ";")[[1]]

per_asm_summaries <- list()
for (i in seq_along(asm_tsv_paths)) {
  p <- asm_tsv_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(assigned_subgenome = col_character()))
    df$assembly <- asm_name_list[i]
    per_asm_summaries[[i]] <- df
  }
}
all_summary_df <- if (length(per_asm_summaries) > 0) bind_rows(per_asm_summaries) else NULL
has_per_asm <- !is.null(all_summary_df) && nrow(all_summary_df) > 0

has_any_chrom <- any(comp_df$n_chrom_assigned > 0, na.rm = TRUE)
has_any_contam <- any(comp_df$n_contaminants > 0, na.rm = TRUE)
has_any_depth <- any(!is.na(comp_df$mean_chrom_depth) & comp_df$mean_chrom_depth > 0)

chr_like_minlen <- as.numeric(params$chr_like_minlen)

# PDF output paths
pdf_classbar <- paste0(params$out_prefix, ".classification_comparison.pdf")
pdf_heatmap  <- paste0(params$out_prefix, ".chromosome_completeness.pdf")

# ---------------------------------------------------------------------------
# Load per-assembly macro_blocks and pairwise macro_blocks for synteny plot
# ---------------------------------------------------------------------------
per_asm_macro_paths <- if (nzchar(params$per_assembly_macro_blocks_tsvs))
  strsplit(params$per_assembly_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_macro_paths <- if (nzchar(params$pairwise_macro_blocks_tsvs))
  strsplit(params$pairwise_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_pair_names <- if (nzchar(params$pairwise_names))
  strsplit(params$pairwise_names, ";")[[1]] else character(0)

# Load per-assembly macro_blocks (ref->asm) and tag with assembly name
per_asm_macro_list <- list()
for (i in seq_along(per_asm_macro_paths)) {
  p <- per_asm_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      per_asm_macro_list[[i]] <- df
    }
  }
}
per_asm_macro_df <- if (length(per_asm_macro_list) > 0) bind_rows(per_asm_macro_list) else NULL

# Load pairwise macro_blocks (asm->asm)
pairwise_macro_list <- list()
for (i in seq_along(pairwise_macro_paths)) {
  p <- pairwise_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0 && i <= length(pairwise_pair_names)) {
      df$pair_name <- pairwise_pair_names[i]
      # Detect rescue pairs (prefixed with "rescue:")
      raw_name <- pairwise_pair_names[i]
      df$is_rescue <- startsWith(raw_name, "rescue:")
      clean_name <- sub("^rescue:", "", raw_name)
      # Extract left and right assembly names from pair name
      parts <- strsplit(clean_name, "_vs_", fixed = TRUE)[[1]]
      if (length(parts) >= 2) {
        df$left_assembly <- parts[1]
        df$right_assembly <- paste(parts[-1], collapse = "_vs_")
      } else {
        warning(sprintf("Cannot parse pairwise pair name: %s", pairwise_pair_names[i]))
        next
      }
      pairwise_macro_list[[i]] <- df
    }
  }
}
pairwise_macro_df <- if (length(pairwise_macro_list) > 0) bind_rows(pairwise_macro_list) else NULL

# Synteny data available for both nucleotide and protein modes
has_synteny_data <- !is.null(per_asm_macro_df) && nrow(per_asm_macro_df) > 0

# ---------------------------------------------------------------------------
# Load per-assembly rDNA arrays TSVs
# ---------------------------------------------------------------------------
rdna_arrays_paths <- if (nzchar(params$per_assembly_rdna_arrays_tsvs))
  strsplit(params$per_assembly_rdna_arrays_tsvs, ";")[[1]] else character(0)

rdna_arrays_list <- list()
for (i in seq_along(rdna_arrays_paths)) {
  p <- rdna_arrays_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE)
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      rdna_arrays_list[[i]] <- df
    }
  }
}
all_rdna_arrays_df <- if (length(rdna_arrays_list) > 0) bind_rows(rdna_arrays_list) else NULL
has_rdna_arrays <- !is.null(all_rdna_arrays_df) && nrow(all_rdna_arrays_df) > 0

# ---------------------------------------------------------------------------
# Load per-assembly contaminants TSVs (for Top Taxa section)
# ---------------------------------------------------------------------------
contam_tsv_paths <- if (nzchar(params$per_assembly_contaminants_tsvs))
  strsplit(params$per_assembly_contaminants_tsvs, ";")[[1]] else character(0)

contam_detail_list <- list()
for (i in seq_along(contam_tsv_paths)) {
  p <- contam_tsv_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE)
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      contam_detail_list[[i]] <- df
    }
  }
}
all_contam_detail_df <- if (length(contam_detail_list) > 0) bind_rows(contam_detail_list) else NULL
has_contam_detail <- !is.null(all_contam_detail_df) && nrow(all_contam_detail_df) > 0
```

---

<p class="section-intro">
**Reference**: `r if (nchar(params$reference_name) > 0) params$reference_name else "\u2014"` &emsp;
**Synteny mode**: `r params$synteny_mode` &emsp;
**Assemblies**: `r n_asm` &emsp;
**Date**: `r Sys.Date()`
</p>

# 1. Assembly Overview {.tabset}

```{r compleasm_setup, include=FALSE}
# Check if compleasm data is available in comparison TSV
has_compleasm <- "compleasm_chrs_S" %in% names(comp_df) &&
  any(!is.na(comp_df$compleasm_chrs_S) & comp_df$compleasm_chrs_S != "", na.rm = TRUE)
```

## Classification

```{r classification_table, eval=has_per_asm, results='asis'}
# Summarise per assembly × classification
clf_wide <- all_summary_df %>%
  group_by(assembly, classification) %>%
  summarise(total_mb = sum(length, na.rm = TRUE) / 1e6,
            n_contigs = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = classification,
    values_from = c(total_mb, n_contigs),
    values_fill = list(total_mb = 0, n_contigs = 0L)
  )

# Ensure all classification columns exist (even if no assembly has them)
for (cl in classification_levels) {
  mb_col <- paste0("total_mb_", cl)
  if (!mb_col %in% names(clf_wide)) clf_wide[[mb_col]] <- 0
}

# Build stacked bar SVG
clf_colors <- unname(classification_colors[classification_levels])

make_clf_bar <- function(...) {
  vals <- as.numeric(c(...))
  total <- sum(vals, na.rm = TRUE)
  if (is.na(total) || total == 0) return("")
  bar_w <- 200; bar_h <- 14; x <- 0; rects <- ""
  for (j in seq_along(vals)) {
    w <- vals[j] / total * bar_w
    if (!is.na(w) && w > 0.1) {
      rects <- paste0(rects, sprintf(
        '<rect x="%.1f" y="0" width="%.1f" height="%d" fill="%s"/>', x, w, bar_h, clf_colors[j]))
      x <- x + w
    }
  }
  sprintf('<svg width="%d" height="%d" style="vertical-align:middle" xmlns="http://www.w3.org/2000/svg">%s</svg>',
          bar_w, bar_h, rects)
}

mb_cols <- paste0("total_mb_", classification_levels)

clf_table <- clf_wide %>%
  mutate(assembly = factor(assembly, levels = asm_names)) %>%
  arrange(assembly) %>%
  rowwise() %>%
  mutate(bar = make_clf_bar(!!!syms(mb_cols))) %>%
  ungroup()

# Select and rename columns for gt
clf_gt_df <- clf_table %>%
  select(Assembly = assembly, bar, all_of(mb_cols))

# Short column labels
clf_short <- c(
  "chrom_assigned"     = "Chr",
  "chrom_unassigned"   = "Unasgn",
  "organelle_complete" = "Org",
  "organelle_debris"   = "OrgD",
  "rDNA"               = "rDNA",
  "contaminant"        = "Cont",
  "chrom_debris"       = "ChrD",
  "debris"             = "Debris",
  "unclassified"       = "Unclf"
)

col_labels <- setNames(
  as.list(c("", clf_short[classification_levels])),
  c("bar", mb_cols)
)

# Drop columns that are all zero
nonzero_mb <- mb_cols[sapply(mb_cols, function(col) any(clf_gt_df[[col]] > 0))]
zero_mb <- setdiff(mb_cols, nonzero_mb)

clf_gt_df %>%
  { if (length(zero_mb) > 0) select(., -all_of(zero_mb)) else . } %>%
  gt() %>%
  fmt(columns = bar, fns = function(x) lapply(x, html)) %>%
  fmt_number(columns = any_of(nonzero_mb), decimals = 1) %>%
  cols_label(.list = col_labels[c("bar", nonzero_mb)]) %>%
  cols_width(bar ~ px(220)) %>%
  cols_width(any_of(nonzero_mb) ~ px(52)) %>%
  cols_align(align = "right", columns = any_of(nonzero_mb)) %>%
  tab_header(
    title = md("**Classification**"),
    subtitle = "Mb per category"
  ) %>%
  tab_source_note(md(paste0(
    paste(sapply(classification_levels[classification_levels %in%
      sub("total_mb_", "", nonzero_mb)], function(cl) {
      sprintf('<span style="display:inline-block;width:12px;height:12px;background:%s;margin-right:3px;vertical-align:middle;"></span>%s',
              classification_colors[cl], classification_labels[cl])
    }), collapse = " &nbsp; ")
  ))) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()
```

`r if (has_compleasm) "## BUSCO Completeness" else ""`

```{r compleasm_table, eval=has_compleasm, results='asis'}
busco_colors <- c(
  "Single-copy"  = "#33a2c4",
  "Duplicated"   = "#225ea8",
  "Fragmented"   = "#fdae61",
  "Interspersed" = "#fee08b",
  "Missing"      = "#d73027"
)

make_busco_bar <- function(s, d, f, i, m) {
  vals <- c(s, d, f, i, m)
  cols <- unname(busco_colors)
  total <- sum(vals, na.rm = TRUE)
  if (is.na(total) || total == 0) return("")
  bar_w <- 140; bar_h <- 14; x <- 0; rects <- ""
  for (j in seq_along(vals)) {
    w <- vals[j] / total * bar_w
    if (!is.na(w) && w > 0) {
      rects <- paste0(rects, sprintf(
        '<rect x="%.1f" y="0" width="%.1f" height="%d" fill="%s"/>', x, w, bar_h, cols[j]))
      x <- x + w
    }
  }
  sprintf('<svg width="%d" height="%d" style="vertical-align:middle" xmlns="http://www.w3.org/2000/svg">%s</svg>',
          bar_w, bar_h, rects)
}

# Build wide-format table: one row per assembly, paired columns for chrs / non-chrs
busco_df <- comp_df %>%
  transmute(
    Assembly = assembly,
    chrs_S = as.numeric(compleasm_chrs_S),
    chrs_D = as.numeric(compleasm_chrs_D),
    chrs_F = as.numeric(compleasm_chrs_F),
    chrs_I = as.numeric(compleasm_chrs_I),
    chrs_M = as.numeric(compleasm_chrs_M),
    non_S  = as.numeric(compleasm_non_chrs_S),
    non_D  = as.numeric(compleasm_non_chrs_D),
    non_F  = as.numeric(compleasm_non_chrs_F),
    non_I  = as.numeric(compleasm_non_chrs_I),
    non_M  = as.numeric(compleasm_non_chrs_M)
  ) %>%
  filter(!is.na(chrs_S)) %>%
  rowwise() %>%
  mutate(
    chrs_bar = make_busco_bar(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M),
    non_bar  = make_busco_bar(non_S, non_D, non_F, non_I, non_M)
  ) %>%
  ungroup() %>%
  mutate(spacer = "") %>%
  select(Assembly,
         chrs_bar, chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
         spacer,
         non_bar, non_S, non_D, non_F, non_I, non_M)

lineage_name <- comp_df$compleasm_lineage[!is.na(comp_df$compleasm_lineage) &
                                           comp_df$compleasm_lineage != ""][1]

busco_df %>%
  gt() %>%
  fmt(columns = c(chrs_bar, non_bar), fns = function(x) lapply(x, html)) %>%
  fmt_number(columns = c(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
                          non_S, non_D, non_F, non_I, non_M),
             decimals = 1) %>%
  cols_label(
    chrs_bar = "", chrs_S = "% S", chrs_D = "D", chrs_F = "F",
    chrs_I = "I", chrs_M = "M",
    spacer = "",
    non_bar = "", non_S = "% S", non_D = "D", non_F = "F",
    non_I = "I", non_M = "M"
  ) %>%
  tab_spanner(label = "Chromosomes",
              columns = c(chrs_bar, chrs_S, chrs_D, chrs_F, chrs_I, chrs_M)) %>%
  tab_spanner(label = "Non-chromosomes",
              columns = c(non_bar, non_S, non_D, non_F, non_I, non_M)) %>%
  cols_width(c(chrs_bar, non_bar) ~ px(160)) %>%
  cols_width(c(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
               non_S, non_D, non_F, non_I, non_M) ~ px(48)) %>%
  cols_width(spacer ~ px(6)) %>%
  cols_align(align = "right",
             columns = c(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
                          non_S, non_D, non_F, non_I, non_M)) %>%
  tab_header(
    title = md("**BUSCO Completeness**"),
    subtitle = md(sprintf("Lineage: **%s**", lineage_name))
  ) %>%
  tab_source_note(md(paste0(
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[1], ';margin-right:3px;vertical-align:middle;"></span>Single-copy &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[2], ';margin-right:3px;vertical-align:middle;"></span>Duplicated &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[3], ';margin-right:3px;vertical-align:middle;"></span>Fragmented &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[4], ';margin-right:3px;vertical-align:middle;"></span>Interspersed &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[5], ';margin-right:3px;vertical-align:middle;"></span>Missing'
  ))) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()
```

## Summary Table

```{r overview_table}
# Detect whether reference has a plastid (chrC/chrPt) sequence
ref_has_plastid <- any(grepl("^chr[C]$|^chrPt$", ref_lengths$ref_id))
any_asm_has_chrC <- any(comp_df$chrC_found == "yes", na.rm = TRUE)
show_chrC <- ref_has_plastid || any_asm_has_chrC

tbl <- comp_df %>%
  transmute(
    Assembly = assembly,
    Contigs  = total_contigs,
    `Size (Mb)` = round(total_bp / 1e6, 1),
    N50 = round(n50 / 1e6, 2),
    L50 = l50,
    `Largest (Mb)` = round(largest_contig / 1e6, 1),
    `Asgn / Unasgn / Chim` = paste(n_chrom_assigned, "/",
                                     n_chrom_unassigned, "/",
                                     n_chimeric),
    `Full-length` = n_full_length,
    `Both telo` = n_both_telomeres,
    `Mean ref cov` = ifelse(is.na(mean_ref_coverage), "\u2014",
                            sprintf("%.1f%%", mean_ref_coverage * 100)),
    `Mean identity` = ifelse(is.na(mean_identity), "\u2014",
                             sprintf("%.2f%%", mean_identity * 100)),
    Contaminants = n_contaminants,
    rDNA = n_rdna,
    chrC = ifelse(chrC_found == "yes", "\u2713", "\u2014"),
    chrM = ifelse(chrM_found == "yes", "\u2713", "\u2014")
  )

gt_obj <- tbl %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Assembly Comparison Summary") %>%
  cols_align(align = "right", columns = where(is.numeric)) %>%
  tab_spanner(label = "Contiguity", columns = c(Contigs, `Size (Mb)`, N50, L50, `Largest (Mb)`)) %>%
  tab_spanner(label = "Chromosomes", columns = c(`Asgn / Unasgn / Chim`, `Full-length`, `Both telo`, `Mean ref cov`, `Mean identity`))

# Conditionally include chrC in "Other" spanner
if (show_chrC) {
  gt_obj <- gt_obj %>%
    tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrC, chrM))
} else {
  gt_obj <- gt_obj %>%
    cols_hide(columns = chrC) %>%
    tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrM))
}
gt_obj
```

## Nx Curves

```{r nx_curves, fig.width=10, fig.height=5}
if (has_per_asm) {
  # Compute Nx curve for each assembly
  nx_list <- list()
  for (asm in asm_names) {
    asm_df <- all_summary_df %>% filter(assembly == asm)
    lens <- sort(asm_df$length, decreasing = TRUE)
    total <- sum(lens)
    if (total == 0) next
    cum_frac <- cumsum(lens) / total
    # Sample at 1% intervals for smooth curves
    nx_pts <- tibble(nx_pct = 0:100)
    nx_pts$contig_len <- sapply(nx_pts$nx_pct, function(pct) {
      if (pct == 0) return(lens[1])
      idx <- which(cum_frac >= pct / 100)[1]
      if (is.na(idx)) return(lens[length(lens)])
      lens[idx]
    })
    nx_pts$assembly <- asm
    # Compute N50 and N90 for annotation
    n50_val <- nx_pts$contig_len[nx_pts$nx_pct == 50]
    n90_val <- nx_pts$contig_len[nx_pts$nx_pct == 90]
    nx_pts$n50 <- n50_val
    nx_pts$n90 <- n90_val
    nx_list[[asm]] <- nx_pts
  }

  if (length(nx_list) > 0) {
    nx_df <- bind_rows(nx_list) %>%
      mutate(assembly = factor(assembly, levels = asm_names),
             contig_mb = contig_len / 1e6)

    p_nx <- ggplot(nx_df, aes(x = nx_pct, y = contig_mb,
                              color = assembly, group = assembly)) +
      geom_step_interactive(
        aes(tooltip = paste0(assembly, "\nNx", nx_pct, " = ",
                             sprintf("%.2f Mb", contig_mb)),
            data_id = assembly),
        linewidth = 0.8
      ) +
      scale_color_manual(values = asm_colors, name = NULL) +
      labs(x = "Nx (%)", y = "Contig length (Mb)") +
      theme_minimal(base_size = base_font_pt, base_family = base_family) +
      axis_theme +
      theme(legend.position = "bottom")
    p_nx <- apply_legend_theme(p_nx)

    girafe_styled(ggobj = p_nx, width_svg = 10, height_svg = 5)
  } else {
    cat("No contig data available for Nx curves.")
  }
} else {
  cat("Per-assembly contig summary data not available.")
}
```


# 2. Chromosome Assignment {.tabset}

## Overview

```{r completeness_heatmap, fig.width=10, fig.height=max(3, n_asm * 0.35 + 2)}
if (has_any_chrom) {
  # Exclude organelle chromosomes from the heatmap (they have their own section)
  organelle_ids <- ref_lengths %>%
    filter(grepl("^chr[CM]$|^chrPt$|^chrMt$", ref_id)) %>% pull(ref_id)

  # Natural sort reference chromosomes: strip "chr" prefix, sort by subgenome + number
  ref_order <- ref_lengths %>%
    filter(!ref_id %in% organelle_ids) %>%
    mutate(
      chrom_num = as.numeric(str_extract(ref_id, "\\d+")),
      sg = if (has_subgenomes) subgenome else "G"
    ) %>%
    arrange(sg, chrom_num) %>%
    pull(ref_id)

  hm_data <- compl_df %>%
    filter(!ref_id %in% organelle_ids) %>%
    mutate(
      ref_id = factor(ref_id, levels = ref_order),
      assembly = factor(assembly, levels = asm_names),
      best_ref_coverage = as.numeric(best_ref_coverage),
      coverage_pct = ifelse(is.na(best_ref_coverage), 0, best_ref_coverage * 100),
      mean_identity = as.numeric(mean_identity),
      identity_pct = ifelse(is.na(mean_identity), 0, mean_identity * 100)
    )

  # Join subgenome info from ref_lengths
  hm_data <- hm_data %>%
    left_join(ref_lengths %>% select(ref_id, subgenome), by = "ref_id") %>%
    mutate(sg = if (has_subgenomes) subgenome else "G")

  # --- Compute per-cell fill color using per-subgenome gradient ---
  # For each subgenome, gradient from white (0%) to dark subgenome color (100%)
  # Fill intensity is based on mean identity (not reference coverage)
  compute_sg_fill <- function(identity, sg_val) {
    sg_key <- if (sg_val %in% names(sg_cols_dark)) sg_val else "NA"
    dark_col <- sg_cols_dark[sg_key]
    if (is.na(identity) || identity <= 0) return("#f0f0f0")
    frac <- min(identity / 100, 1)
    ramp <- colorRampPalette(c("white", dark_col))(101)
    ramp[round(frac * 100) + 1]
  }

  hm_data$cell_fill <- mapply(compute_sg_fill, hm_data$identity_pct, hm_data$sg)

  # --- Off-target coverage computation ---
  # For each (assembly, ref_chromosome), find blocks from contigs NOT assigned to that ref_id
  has_off_target <- FALSE
  if (!is.null(per_asm_macro_df) && has_per_asm) {
    # Build contig-to-assigned mapping
    contig_assigned <- all_summary_df %>%
      filter(classification == "chrom_assigned") %>%
      select(assembly, contig, assigned_ref_id)

    off_target_df <- per_asm_macro_df %>%
      left_join(contig_assigned,
                by = c("assembly", "contig" = "contig")) %>%
      filter(!is.na(assigned_ref_id), ref_id != assigned_ref_id) %>%
      group_by(assembly, ref_id) %>%
      summarise(off_target_bp = sum(union_bp, na.rm = TRUE), .groups = "drop") %>%
      left_join(ref_lengths %>% select(ref_id, ref_len), by = "ref_id") %>%
      mutate(off_target_pct = off_target_bp / ref_len * 100)

    if (nrow(off_target_df) > 0) {
      has_off_target <- TRUE
      hm_data <- hm_data %>%
        left_join(off_target_df %>% select(assembly, ref_id, off_target_pct),
                  by = c("assembly", "ref_id")) %>%
        mutate(off_target_pct = ifelse(is.na(off_target_pct), 0, off_target_pct))
    }
  }
  if (!has_off_target) {
    hm_data$off_target_pct <- 0
  }

  # --- Compute per-cell telomere status from per-assembly summaries ---
  if (has_per_asm && "has_5p_telomere" %in% names(all_summary_df) &&
      "has_3p_telomere" %in% names(all_summary_df)) {
    telo_info <- all_summary_df %>%
      filter(classification == "chrom_assigned") %>%
      group_by(assembly, assigned_ref_id) %>%
      summarise(
        any_5p = any(has_5p_telomere == "yes", na.rm = TRUE),
        any_3p = any(has_3p_telomere == "yes", na.rm = TRUE),
        .groups = "drop"
      )
    hm_data <- hm_data %>%
      left_join(telo_info, by = c("assembly" = "assembly",
                                   "ref_id" = "assigned_ref_id")) %>%
      mutate(any_5p = ifelse(is.na(any_5p), FALSE, any_5p),
             any_3p = ifelse(is.na(any_3p), FALSE, any_3p))
  } else {
    hm_data <- hm_data %>%
      mutate(any_5p = (has_both_telomeres == "yes"),
             any_3p = (has_both_telomeres == "yes"))
  }

  # Shape status for each cell
  hm_data <- hm_data %>%
    mutate(
      shape_status = case_when(
        n_contigs == 0              ~ "unassigned",
        is_full_length != "yes"     ~ "incomplete",
        any_5p & any_3p             ~ "telo_both",
        any_5p & !any_3p            ~ "telo_5p",
        !any_5p & any_3p            ~ "telo_3p",
        TRUE                        ~ "full_length"
      ),
      multi_label = ifelse(n_contigs > 1, as.character(n_contigs), "")
    )

  # --- Border color: dark subgenome color or red gradient for off-target ---
  compute_offtarget_border <- function(ot_pct) {
    if (is.na(ot_pct) || ot_pct <= 0) return(NA_character_)
    frac <- min(ot_pct / 20, 1)
    ramp <- colorRampPalette(c("#e08080", "#e74c3c"))(101)
    ramp[round(frac * 100) + 1]
  }
  hm_data <- hm_data %>%
    mutate(
      sg_dark_col = unname(sg_cols_dark[ifelse(sg %in% names(sg_cols_dark), sg, "NA")]),
      border_color = ifelse(
        off_target_pct > 0,
        sapply(off_target_pct, compute_offtarget_border),
        darken(sg_dark_col, 0.2)
      ),
      border_lw = ifelse(off_target_pct > 0, 0.6, 0.3),
      tooltip = paste0(
        ref_id, " \u2014 ", assembly, "\n",
        "Contigs: ", n_contigs, "\n",
        "Assigned bp: ", scales::comma(total_assigned_bp), "\n",
        "Best ref coverage: ", fmt_pct(best_ref_coverage), "\n",
        "Mean identity: ", fmt_pct(mean_identity), "\n",
        "Full-length: ", is_full_length, "\n",
        "Telomeres: ", case_when(
          shape_status == "telo_both" ~ "both",
          shape_status == "telo_5p"   ~ "5\u2032 only",
          shape_status == "telo_3p"   ~ "3\u2032 only",
          TRUE ~ "none"
        ),
        ifelse(off_target_pct > 0,
               paste0("\nOff-target: ", sprintf("%.1f%%", off_target_pct)), "")
      )
    )

  # =====================================================================
  # X-axis layout: manual positioning with subgenome gaps
  # =====================================================================
  x_spacing   <- 1.0
  x_group_gap <- 1.8
  ref_layout <- ref_lengths %>%
    filter(!ref_id %in% organelle_ids) %>%
    mutate(
      chrom_num = as.numeric(str_extract(ref_id, "\\d+")),
      sg = if (has_subgenomes) subgenome else "G"
    ) %>%
    arrange(sg, chrom_num)
  sg_groups   <- unique(ref_layout$sg)
  n_sg_groups <- length(sg_groups)

  x_positions <- tibble(ref_id = character(), x_center = numeric(),
                         sg = character(), chrom_label = character())
  x_cursor <- 1.0
  for (sg_i in seq_along(sg_groups)) {
    sg_val  <- sg_groups[sg_i]
    sg_refs <- ref_layout %>% filter(sg == sg_val)
    for (j in seq_len(nrow(sg_refs))) {
      rid <- sg_refs$ref_id[j]
      lbl <- str_replace(rid, "^chr", "")
      if (has_subgenomes) lbl <- str_replace(lbl, "[A-Z]$", "")
      x_positions <- bind_rows(x_positions, tibble(
        ref_id = rid, x_center = x_cursor, sg = sg_val, chrom_label = lbl
      ))
      if (j < nrow(sg_refs)) x_cursor <- x_cursor + x_spacing
    }
    if (sg_i < n_sg_groups) x_cursor <- x_cursor + x_group_gap
  }

  # Y-axis layout
  y_spacing <- 1.0
  y_positions <- tibble(assembly = asm_names,
                         y_center = seq_along(asm_names) * y_spacing)

  # Join positions into hm_data
  hm_data <- hm_data %>%
    mutate(assembly = as.character(assembly),
           ref_id = as.character(ref_id)) %>%
    left_join(x_positions %>% select(ref_id, x_center), by = "ref_id") %>%
    left_join(y_positions, by = "assembly")

  # =====================================================================
  # Polygon construction helpers
  # =====================================================================
  pill_hw     <- 0.30
  rect_hh     <- 0.32   # half-height of core rectangle (straight edges)
  cap_r_val   <- 0.08   # cap extends beyond rect; ratio = 0.32/0.40 = 80%
  jagged_hh   <- 0.34   # slightly taller than rect for zigzag visibility
  n_teeth_val <- 2
  tooth_d_val <- 0.06   # deeper teeth for visibility

  # rect_hh = half-height of core rectangle; caps extend beyond by cr
  make_pill <- function(xc, yc, hw, rh, cr, cap_top, cap_bot, n_arc = 20) {
    rect_top <- yc - rh
    rect_bot <- yc + rh
    x <- c(); y <- c()
    if (cap_top) {
      theta <- seq(0, pi, length.out = n_arc)
      x <- c(x, xc + hw * cos(theta))
      y <- c(y, rect_top - cr * sin(theta))
    } else {
      x <- c(x, xc + hw, xc - hw)
      y <- c(y, rect_top, rect_top)
    }
    x <- c(x, xc - hw); y <- c(y, rect_bot)
    if (cap_bot) {
      theta <- seq(pi, 0, length.out = n_arc)
      x <- c(x, xc + hw * cos(theta))
      y <- c(y, rect_bot + cr * sin(theta))
    } else {
      x <- c(x, xc + hw); y <- c(y, rect_bot)
    }
    tibble(x = x, y = y)
  }

  make_pill_jagged <- function(xc, yc, hw, hh_rect, n_t, tooth_d) {
    top_y <- yc - hh_rect; bot_y <- yc + hh_rect
    tw <- (2 * hw) / n_t
    x <- c(); y <- c()
    for (i in 0:(n_t - 1)) {
      xl <- xc - hw + i * tw; xm <- xl + tw / 2; xr <- xl + tw
      x <- c(x, xl, xm, xr)
      y <- c(y, top_y, top_y + tooth_d, top_y)
    }
    x <- c(x, xc + hw); y <- c(y, bot_y)
    for (i in (n_t - 1):0) {
      xl <- xc - hw + i * tw; xm <- xl + tw / 2; xr <- xl + tw
      x <- c(x, xr, xm, xl)
      y <- c(y, bot_y, bot_y - tooth_d, bot_y)
    }
    tibble(x = x, y = y)
  }

  make_cell_polygon <- function(xc, yc, status) {
    switch(status,
      telo_both   = make_pill(xc, yc, pill_hw, rect_hh, cap_r_val, TRUE, TRUE),
      telo_5p     = make_pill(xc, yc, pill_hw, rect_hh, cap_r_val, TRUE, FALSE),
      telo_3p     = make_pill(xc, yc, pill_hw, rect_hh, cap_r_val, FALSE, TRUE),
      full_length = make_pill(xc, yc, pill_hw, rect_hh, 0, FALSE, FALSE),
      incomplete  = make_pill_jagged(xc, yc, pill_hw, jagged_hh,
                                     n_teeth_val, tooth_d_val),
      unassigned  = make_pill(xc, yc, pill_hw, rect_hh, 0, FALSE, FALSE)
    )
  }

  # =====================================================================
  # Build polygon data frame
  # =====================================================================
  poly_list <- list()
  for (i in seq_len(nrow(hm_data))) {
    row <- hm_data[i, ]
    if (is.na(row$x_center) || is.na(row$y_center)) next
    pts <- make_cell_polygon(row$x_center, row$y_center, row$shape_status)
    pts$cell_id <- i
    pts$fill_color <- if (row$shape_status == "unassigned") "#f0f0f0" else row$cell_fill
    pts$border_color <- if (row$shape_status == "unassigned") NA_character_ else row$border_color
    pts$border_lw <- if (row$shape_status == "unassigned") 0 else row$border_lw
    pts$tooltip <- row$tooltip
    pts$data_id <- paste0(row$ref_id, "_", row$assembly)
    pts$shape_status <- row$shape_status
    poly_list[[i]] <- pts
  }
  all_polys <- bind_rows(poly_list)
  unassigned_polys <- all_polys %>% filter(shape_status == "unassigned")
  assigned_polys   <- all_polys %>% filter(shape_status != "unassigned")
  multi_df <- hm_data %>% filter(multi_label != "")

  # =====================================================================
  # Annotation data
  # =====================================================================
  sg_extent <- x_positions %>%
    group_by(sg) %>%
    summarise(x_min = min(x_center), x_max = max(x_center),
              x_mid = mean(x_center), .groups = "drop")
  y_grid_top   <- min(y_positions$y_center) - 0.5
  chr_label_y1 <- y_grid_top - 0.10   # odd-numbered chromosomes (closer to grid)
  chr_label_y2 <- y_grid_top - 0.35   # even-numbered chromosomes (staggered up)
  x_positions  <- x_positions %>%
    mutate(label_y = ifelse(as.numeric(chrom_label) %% 2 == 0,
                            chr_label_y2, chr_label_y1))
  sep_line_y   <- y_grid_top - 0.65
  sg_label_y   <- y_grid_top - 0.90

  # =====================================================================
  # Compose ggplot
  # =====================================================================
  p_hm <- ggplot()

  # 1. Unassigned placeholders (grey, no border)
  if (nrow(unassigned_polys) > 0) {
    p_hm <- p_hm +
      geom_polygon(data = unassigned_polys,
                   aes(x = x, y = y, group = cell_id),
                   fill = "#f0f0f0", color = NA)
  }

  # 2. Assigned pills (precomputed fill/color via scale_*_identity)
  if (nrow(assigned_polys) > 0) {
    p_hm <- p_hm +
      geom_polygon_interactive(
        data = assigned_polys,
        aes(x = x, y = y, group = cell_id,
            fill = fill_color, color = border_color,
            linewidth = border_lw,
            tooltip = tooltip, data_id = data_id),
        show.legend = FALSE
      ) +
      scale_fill_identity() +
      scale_color_identity() +
      scale_linewidth_identity()
  }

  # 3. Multi-contig count labels (small number inside pill)
  if (nrow(multi_df) > 0) {
    p_hm <- p_hm +
      geom_text(data = multi_df,
                aes(x = x_center, y = y_center, label = multi_label),
                size = base_font_pt / .pt * 0.8, color = "white",
                fontface = "bold", family = base_family)
  }

  # 4. Subgenome separator lines + labels (polyploid only)
  if (has_subgenomes && n_sg_groups >= 2) {
    sg_line_df <- sg_extent %>%
      mutate(y = sep_line_y,
             color = unname(sg_cols_dark[sg]))
    p_hm <- p_hm +
      geom_segment(data = sg_line_df,
                   aes(x = x_min - 0.4, xend = x_max + 0.4,
                       y = y, yend = y),
                   color = sg_line_df$color, linewidth = 0.5) +
      geom_text(data = sg_extent,
                aes(x = x_mid, y = sg_label_y,
                    label = paste0("Subgenome ", sg)),
                size = base_font_pt / .pt, fontface = "bold",
                family = base_family)
  }

  # 5. Chromosome number labels (staggered odd/even to avoid overlap)
  n_chroms <- length(ref_order)
  p_hm <- p_hm +
    geom_text(data = x_positions,
              aes(x = x_center, y = label_y, label = chrom_label),
              size = base_font_pt / .pt, family = base_family,
              hjust = 0.5)

  # 6. Assembly labels (left of grid)
  x_left_edge <- min(x_positions$x_center) - pill_hw - 0.5
  p_hm <- p_hm +
    geom_text(data = y_positions,
              aes(x = x_left_edge, y = y_center, label = assembly),
              size = base_font_pt / .pt, hjust = 1, family = base_family)

  # Scales and theme
  p_hm <- p_hm +
    scale_y_reverse() +
    coord_cartesian(clip = "off") +
    labs(caption = paste0(
      "Pill = telomeres, rect = full-length, zigzag = incomplete, ",
      "grey = unassigned. Fill intensity = mean identity.",
      if (has_off_target) " Red border = off-target synteny." else ""
    )) +
    theme_void(base_size = base_font_pt, base_family = base_family) +
    theme(
      plot.margin = margin(8, 2, 5,
                           max(25, max(nchar(as.character(asm_names))) * 3)),
      plot.caption = element_text(size = 6, hjust = 0, color = "#666")
    )

  # --- Dynamic sizing ---
  n_chroms <- length(ref_order)
  n_sg_gaps <- max(0, n_sg_groups - 1)
  hm_width  <- max(6, n_chroms * 0.45 + n_sg_gaps * 0.8 + 2.5)
  hm_height <- max(3, n_asm * 0.5 + 2.0)

  save_pdf(pdf_heatmap, p_hm, width = min(7.2, hm_width), height = hm_height)
  girafe_styled(ggobj = p_hm,
                width_svg = min(13, hm_width * 1.4),
                height_svg = hm_height)
} else {
  cat("No chromosome-assigned contigs found in any assembly.")
}
```

<div class="note-box">
<b>Legend</b>:&ensp;
<svg width="14" height="24" style="vertical-align:middle"><rect x="1" y="1" width="12" height="22" rx="6" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&ensp;Both telomeres&emsp;
<svg width="14" height="24" style="vertical-align:middle"><path d="M1,7 C1,1 13,1 13,7 L13,23 L1,23 Z" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&ensp;5&#x2032; telomere&emsp;
<svg width="14" height="24" style="vertical-align:middle"><path d="M1,1 L13,1 L13,17 C13,23 1,23 1,17 Z" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&ensp;3&#x2032; telomere&emsp;
<svg width="14" height="20" style="vertical-align:middle"><rect x="1" y="1" width="12" height="18" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&ensp;Full-length&emsp;
<svg width="14" height="20" style="vertical-align:middle"><path d="M1,4 L4,1 L7,4 L10,1 L13,4 L13,16 L10,19 L7,16 L4,19 L1,16 Z" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&ensp;Incomplete&emsp;
<svg width="14" height="20" style="vertical-align:middle"><rect x="1" y="1" width="12" height="18" fill="#f0f0f0" stroke="none"/></svg>&ensp;Unassigned<br>
Fill intensity = reference coverage. Red border = off-target synteny. Small numbers = multi-contig.
</div>

```{r scatter_setup, eval=(has_any_chrom && has_per_asm), include=FALSE}
scatter_df <- all_summary_df %>%
  filter(classification == "chrom_assigned") %>%
  mutate(
    length_mb = length / 1e6,
    assembly = factor(assembly, levels = asm_names)
  )
if (has_subgenomes && "assigned_subgenome" %in% names(scatter_df)) {
  scatter_df <- scatter_df %>%
    mutate(sg = ifelse(is.na(assigned_subgenome) | assigned_subgenome == "NA",
                       "Other", assigned_subgenome))
} else {
  scatter_df$sg <- "All"
}
has_scatter_ident <- "seq_identity_vs_ref" %in% names(scatter_df) &&
                     !all(is.na(scatter_df$seq_identity_vs_ref))
has_scatter_cov   <- "ref_coverage" %in% names(scatter_df) &&
                     !all(is.na(scatter_df$ref_coverage))
```

```{r scatter_identity_header, eval=(has_any_chrom && has_per_asm && exists("has_scatter_ident") && has_scatter_ident), results='asis'}
cat("## Length vs Identity\n\n")
```

```{r scatter_identity, eval=(has_any_chrom && has_per_asm && exists("has_scatter_ident") && has_scatter_ident), fig.width=10, fig.height=6}
si_df <- scatter_df %>% filter(!is.na(seq_identity_vs_ref))

p_si <- ggplot(si_df, aes(x = length_mb, y = seq_identity_vs_ref * 100,
                           color = assembly)) +
  geom_point_interactive(
    aes(tooltip = paste0(assembly, "\n", contig, "\n",
                         ifelse(!is.na(assigned_ref_id),
                                paste0("Assigned: ", assigned_ref_id, "\n"), ""),
                         sprintf("%.1f Mb, %.2f%% identity", length_mb,
                                 seq_identity_vs_ref * 100))),
    alpha = 0.6, size = 2
  ) +
  { if (has_subgenomes && length(unique(si_df$sg)) > 1) {
    facet_wrap(~ sg)
  }} +
  scale_color_manual(values = asm_colors, name = NULL) +
  labs(x = "Contig length (Mb)", y = "Sequence identity vs reference (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(legend.position = "bottom")
p_si <- apply_legend_theme(p_si)

girafe_styled(ggobj = p_si, width_svg = 10, height_svg = 6)
```

```{r scatter_coverage_header, eval=(has_any_chrom && has_per_asm && exists("has_scatter_cov") && has_scatter_cov), results='asis'}
cat("## Length vs Ref Coverage\n\n")
```

```{r scatter_coverage, eval=(has_any_chrom && has_per_asm && exists("has_scatter_cov") && has_scatter_cov), fig.width=10, fig.height=6}
sc_df <- scatter_df %>% filter(!is.na(ref_coverage))

p_sc <- ggplot(sc_df, aes(x = length_mb, y = ref_coverage * 100, color = assembly)) +
  geom_point_interactive(
    aes(tooltip = paste0(assembly, "\n", contig, "\n",
                         ifelse(!is.na(assigned_ref_id),
                                paste0("Assigned: ", assigned_ref_id, "\n"), ""),
                         sprintf("%.1f Mb, %.1f%% ref coverage", length_mb,
                                 ref_coverage * 100))),
    alpha = 0.6, size = 2
  ) +
  { if (has_subgenomes && length(unique(sc_df$sg)) > 1) {
    facet_wrap(~ sg)
  }} +
  scale_color_manual(values = asm_colors, name = NULL) +
  labs(x = "Contig length (Mb)", y = "Reference coverage (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(legend.position = "bottom")
p_sc <- apply_legend_theme(p_sc)

girafe_styled(ggobj = p_sc, width_svg = 10, height_svg = 6)
```

```{r completeness_table_setup, include=FALSE}
# Prepare completeness table data with subgenome info and natural sort
if (has_any_chrom) {
  compl_tbl_base <- compl_df %>%
    mutate(best_ref_coverage = as.numeric(best_ref_coverage),
           mean_identity = as.numeric(mean_identity)) %>%
    left_join(ref_lengths %>% select(ref_id, subgenome), by = "ref_id") %>%
    filter(!grepl("^chr[CM]$|^chrPt$|^chrMt$", ref_id)) %>%
    mutate(
      sg = if (has_subgenomes) subgenome else "G",
      chrom_num = as.numeric(str_extract(ref_id, "\\d+")),
      `Assigned (Mb)` = round(total_assigned_bp / 1e6, 1),
      `Ref cov (%)` = ifelse(is.na(best_ref_coverage), "\u2014",
                             sprintf("%.1f", best_ref_coverage * 100)),
      `Identity (%)` = ifelse(is.na(mean_identity), "\u2014",
                              sprintf("%.2f", mean_identity * 100)),
      `Full-length` = ifelse(is_full_length == "yes", "\u2713", ""),
      `Both telo` = ifelse(has_both_telomeres == "yes", "\u2713", "")
    ) %>%
    arrange(sg, chrom_num) %>%
    mutate(ref_id = factor(ref_id, levels = unique(ref_id)))

  compl_sg_levels <- if (has_subgenomes) sort(unique(compl_tbl_base$sg)) else "G"

  make_compl_gt <- function(df, title) {
    df %>%
      select(ref_id, assembly, n_contigs, `Assigned (Mb)`, `Ref cov (%)`,
             `Identity (%)`, `Full-length`, `Both telo`) %>%
      gt(groupname_col = "ref_id") %>%
      gt_theme() %>%
      tab_header(title = title) %>%
      cols_label(
        assembly = "Assembly", n_contigs = "Contigs",
        `Assigned (Mb)` = "Assigned (Mb)",
        `Ref cov (%)` = "Ref cov (%)",
        `Identity (%)` = "Identity (%)",
        `Full-length` = "Full-length",
        `Both telo` = "Both telo"
      ) %>%
      cols_align(align = "right", columns = c(n_contigs, `Assigned (Mb)`,
                                               `Ref cov (%)`, `Identity (%)`)) %>%
      cols_align(align = "center", columns = c(`Full-length`, `Both telo`))
  }
}
```

`r if (has_any_chrom && has_subgenomes && length(compl_sg_levels) >= 2) "## Completeness Table {.tabset}" else "## Completeness Table"`

```{r completeness_table_single, eval=(has_any_chrom && !(has_subgenomes && length(compl_sg_levels) >= 2))}
make_compl_gt(compl_tbl_base, "Chromosome Completeness by Reference")
```

```{r completeness_table_tabset, eval=(has_any_chrom && has_subgenomes && length(compl_sg_levels) >= 2), results='asis'}
for (sg_val in compl_sg_levels) {
  cat(sprintf("\n### Subgenome %s\n\n", sg_val))
  sg_df <- compl_tbl_base %>% filter(sg == sg_val)
  gt_out <- make_compl_gt(sg_df,
    sprintf("Chromosome Completeness \u2014 Subgenome %s", sg_val))
  # Render gt to HTML and emit inline
  cat(as.character(as_raw_html(gt_out)))
  cat("\n\n")
}
```

```{r completeness_table_none, eval=(!has_any_chrom)}
cat("No chromosome-assigned contigs found in any assembly.")
```


```{r synteny_pre, eval=has_synteny_data, include=FALSE}
# Pre-compute subgenome count before the section header so we can
# conditionally add {.tabset} for polyploid references.
has_named_subgenomes <- any(
  !is.na(ref_lengths$subgenome) & ref_lengths$subgenome != "NA"
)
if (has_named_subgenomes) {
  n_subgenomes_pre <- ref_lengths %>%
    filter(!is.na(subgenome) & subgenome != "NA") %>%
    pull(subgenome) %>% unique() %>% length()
} else {
  n_subgenomes_pre <- 1
}
```

`r if (has_synteny_data && exists("n_subgenomes_pre") && n_subgenomes_pre > 1) "# 3. Chromosome Synteny {.tabset}" else if (has_synteny_data) "# 3. Chromosome Synteny" else ""`

```{r synteny_setup, eval=has_synteny_data, include=FALSE}
# =========================================================================
# Section 3: Chromosome Synteny Riparian Plot (transposed layout)
#
# X-axis: chromosomes (chr1..chrN), naturally sorted left to right
# Y-axis: rows — Reference at top, then assemblies in FOFN order
# Ribbons flow vertically between adjacent rows.
# For polyploid references, each subgenome gets its own tab and PDF.
# =========================================================================

# --- Assembly contig lookup (incl. original_name for name mapping) ---
asm_contig_lookup <- list()
for (i in seq_along(asm_name_list)) {
  nm <- asm_name_list[i]
  df <- per_asm_summaries[[i]]
  if (!is.null(df)) {
    lookup <- df %>%
      filter(classification == "chrom_assigned") %>%
      select(contig, original_name, assigned_ref_id, reversed, length) %>%
      mutate(assembly = nm, reversed = (reversed == "yes"))
    asm_contig_lookup[[nm]] <- lookup
  }
}
asm_lookup_df <- bind_rows(asm_contig_lookup)

# --- Fix per-assembly macro_blocks: map original_name -> new_name ---
if (!is.null(per_asm_macro_df) && nrow(asm_lookup_df) > 0) {
  name_map <- asm_lookup_df %>%
    distinct(assembly, original_name, .keep_all = TRUE) %>%
    select(assembly, original_name, new_name = contig)
  per_asm_macro_df <- per_asm_macro_df %>%
    left_join(name_map, by = c("assembly", "contig" = "original_name")) %>%
    mutate(contig_new = coalesce(new_name, contig))
} else if (!is.null(per_asm_macro_df)) {
  per_asm_macro_df <- per_asm_macro_df %>% mutate(contig_new = contig)
}

# --- Fix pairwise macro_blocks: map original contig names -> renamed names ---
if (!is.null(pairwise_macro_df) && nrow(asm_lookup_df) > 0) {
  left_map <- asm_lookup_df %>%
    distinct(assembly, original_name, .keep_all = TRUE) %>%
    select(left_assembly = assembly, original_name, new_ref_id = contig)
  right_map <- asm_lookup_df %>%
    distinct(assembly, original_name, .keep_all = TRUE) %>%
    select(right_assembly = assembly, original_name, new_contig = contig)
  pairwise_macro_df <- pairwise_macro_df %>%
    left_join(left_map, by = c("left_assembly", "ref_id" = "original_name")) %>%
    mutate(ref_id = coalesce(new_ref_id, ref_id)) %>%
    select(-new_ref_id) %>%
    left_join(right_map, by = c("right_assembly", "contig" = "original_name")) %>%
    mutate(contig = coalesce(new_contig, contig)) %>%
    select(-new_contig)
}

# --- Nuclear chromosomes sorted by (subgenome, chrom_num) ---

if (has_named_subgenomes) {
  nuc_ref <- ref_lengths %>%
    filter(!is.na(subgenome) & subgenome != "NA") %>%
    mutate(chrom_num = as.numeric(str_extract(chrom_id, "\\d+"))) %>%
    arrange(subgenome, chrom_num)
} else {
  macro_rids <- unique(per_asm_macro_df$ref_id)
  nuc_ref <- ref_lengths %>%
    filter(ref_id %in% macro_rids) %>%
    mutate(
      subgenome = "all",
      chrom_num = as.numeric(str_extract(ref_id, "\\d+"))
    ) %>%
    arrange(chrom_num)
}

subgenome_order <- unique(nuc_ref$subgenome)
n_subgenomes <- length(subgenome_order)

# --- Determine assemblies per subgenome ---
sg_assemblies <- list()
for (sg in subgenome_order) {
  sg_ref_ids <- nuc_ref %>% filter(subgenome == sg) %>% pull(ref_id)
  asms_with_sg <- asm_lookup_df %>%
    filter(assigned_ref_id %in% sg_ref_ids) %>%
    pull(assembly) %>% unique()
  sg_assemblies[[sg]] <- asm_name_list[asm_name_list %in% asms_with_sg]
}

# --- Chromosome color palette (by chrom_id, shared across subgenomes) ---
unique_chroms <- unique(nuc_ref$chrom_id)
unique_chroms <- unique_chroms[order(as.numeric(str_extract(unique_chroms, "\\d+")))]
n_unique_chroms <- length(unique_chroms)
if (n_unique_chroms <= 9) {
  chrom_pal <- setNames(oi[seq_len(n_unique_chroms)], unique_chroms)
} else {
  chrom_pal <- setNames(
    colorRampPalette(oi[1:9])(n_unique_chroms),
    unique_chroms
  )
}

# Fixed min_span (50kb) — proportional scaling is too aggressive for large genomes
min_span <- 50000

# --- Rescue macro_blocks ---
rescue_macro_df <- if (!is.null(pairwise_macro_df) && "is_rescue" %in% names(pairwise_macro_df)) {
  pairwise_macro_df %>% filter(is_rescue)
} else { NULL }

# =========================================================================
# Per-subgenome plot builder
#
# Returns a list(plot, width, height, has_ribbons) for one subgenome.
# Called once per tab (or once for diploids).
# =========================================================================
build_synteny_plot <- function(sg) {
  sg_chr <- nuc_ref %>% filter(subgenome == sg)
  sg_asms <- sg_assemblies[[sg]]
  sg_chrom_nums <- sort(unique(sg_chr$chrom_num))
  n_chroms <- length(sg_chrom_nums)

  # --- Row names: Reference + assemblies ---
  row_names <- c("Reference", sg_asms)
  n_rows <- length(row_names)
  row_spacing <- 1
  row_centers <- setNames(seq_len(n_rows) * row_spacing, row_names)

  # --- X-axis layout: one column per chromosome ---
  # Pill width proportional to contig length.  Column width is the max total
  # pill width across all assemblies (handles multi-copy chromosomes like
  # chr2A_c1 + chr2A_c2 which need more horizontal space).
  max_ref_len <- max(sg_chr$ref_len)
  x_scale <- max_ref_len / 0.8  # ref pill up to 0.8 units for the largest chrom

  # Pre-compute per-chromosome column width: max of ref_len and the widest
  # assembly (sum of contig lengths + inter-pill gaps for all contigs assigned
  # to that ref_id)
  pill_gap <- 0.40  # gap between adjacent pills within the same column
  col_widths <- numeric(length(sg_chrom_nums))
  for (cn_i in seq_along(sg_chrom_nums)) {
    cn <- sg_chrom_nums[cn_i]
    ref_row <- sg_chr %>% filter(chrom_num == cn)
    rid <- ref_row$ref_id[1]
    ref_w <- ref_row$ref_len[1] / x_scale
    max_asm_w <- ref_w
    for (asm in sg_asms) {
      asm_contigs <- asm_lookup_df %>%
        filter(assembly == asm, assigned_ref_id == rid)
      n_ctg <- nrow(asm_contigs)
      total_w <- sum(asm_contigs$length) / x_scale +
                 max(0, n_ctg - 1) * pill_gap
      max_asm_w <- max(max_asm_w, total_w)
    }
    col_widths[cn_i] <- max_asm_w
  }
  # Add padding between columns (0.2 units minimum gap)
  col_gap <- 0.2
  chrom_x <- numeric(length(sg_chrom_nums))
  chrom_x[1] <- col_widths[1] / 2
  for (cn_i in seq_along(sg_chrom_nums)[-1]) {
    chrom_x[cn_i] <- chrom_x[cn_i - 1] + col_widths[cn_i - 1] / 2 +
                      col_gap + col_widths[cn_i] / 2
  }
  names(chrom_x) <- as.character(sg_chrom_nums)

  chrom_labels <- sg_chr %>%
    arrange(chrom_num) %>%
    transmute(chrom_num, label = str_replace(ref_id, "^[Cc]hr", "")) %>%
    distinct()

  # --- Build pill data ---
  pills_list <- list()
  pill_lw <- 1.6

  for (cn_i in seq_along(sg_chrom_nums)) {
    cn <- sg_chrom_nums[cn_i]
    x_pos <- chrom_x[as.character(cn)]
    ref_row <- sg_chr %>% filter(chrom_num == cn)
    if (nrow(ref_row) == 0) next
    rid <- ref_row$ref_id[1]
    ref_len <- ref_row$ref_len[1]
    cid <- ref_row$chrom_id[1]
    pill_w <- ref_len / x_scale

    # Reference pill — centered on column
    y_ref <- row_centers[["Reference"]]
    pills_list[[length(pills_list) + 1]] <- tibble(
      row_name = "Reference", is_ref = TRUE,
      ref_id = rid, chrom_id = cid, chrom_num = cn, subgenome = sg,
      contig = rid, contig_len = ref_len, reversed = FALSE,
      x_center = x_pos, x_left = x_pos - pill_w/2, x_right = x_pos + pill_w/2,
      y_center = y_ref, pill_width = pill_w,
      tooltip = sprintf("Reference\n%s\n%.2f Mb", rid, ref_len / 1e6)
    )

    # Assembly pills — stack left-to-right, centered on column
    for (asm in sg_asms) {
      y_asm <- row_centers[[asm]]
      lk <- asm_lookup_df %>%
        filter(assembly == asm, assigned_ref_id == rid) %>%
        arrange(desc(length))
      if (nrow(lk) == 0) next

      # Total width of all contigs + gaps for centering
      n_ctg <- nrow(lk)
      total_w <- sum(lk$length) / x_scale + max(0, n_ctg - 1) * pill_gap
      x_cursor <- x_pos - total_w / 2  # center the composite

      for (ci in seq_len(n_ctg)) {
        ctg <- lk[ci, ]
        ctg_w <- ctg$length / x_scale
        pills_list[[length(pills_list) + 1]] <- tibble(
          row_name = asm, is_ref = FALSE,
          ref_id = rid, chrom_id = cid, chrom_num = cn, subgenome = sg,
          contig = ctg$contig, contig_len = ctg$length,
          reversed = ctg$reversed,
          x_center = x_cursor + ctg_w / 2,
          x_left = x_cursor, x_right = x_cursor + ctg_w,
          y_center = y_asm, pill_width = ctg_w,
          tooltip = sprintf("%s\n%s\n%.2f Mb", asm, ctg$contig, ctg$length / 1e6)
        )
        x_cursor <- x_cursor + ctg_w + pill_gap
      }
    }
  }

  all_pills <- bind_rows(pills_list)
  if (nrow(all_pills) == 0) return(NULL)

  # --- Helpers ---
  pos_to_x <- function(pos, seq_len, pill_x_left, pill_w) {
    pill_x_left + (pos / seq_len) * pill_w
  }

  # Transposed ribbon: flows vertically (top→bottom) between rows
  make_ribbon_v <- function(gap_top, gap_bot,
                            top_x_left, top_x_right,
                            bot_x_left, bot_x_right,
                            chrom_id_val, alpha_val, tooltip_text,
                            subgenome_val, data_id_val, n_pts = 20) {
    t <- seq(0, 1, length.out = n_pts)
    s <- 0.5 - 0.5 * cos(pi * t)
    y_fwd <- gap_top + t * (gap_bot - gap_top)
    left_fwd  <- top_x_left  + s * (bot_x_left  - top_x_left)
    right_fwd <- top_x_right + s * (bot_x_right - top_x_right)
    tibble(
      x = c(left_fwd, rev(right_fwd)),
      y = c(y_fwd, rev(y_fwd)),
      chrom_id = chrom_id_val,
      alpha = alpha_val,
      tooltip = tooltip_text,
      subgenome = subgenome_val,
      data_id = data_id_val
    )
  }

  # --- Build gap zones between adjacent rows ---
  gap_zones <- list()
  gap_meta  <- list()
  for (ri in seq_len(n_rows - 1)) {
    top_row <- row_names[ri]
    bot_row <- row_names[ri + 1]
    gap_zones[[ri]] <- c(row_centers[[top_row]] + 0.03,
                         row_centers[[bot_row]] - 0.03)
    gap_meta[[ri]] <- list(
      top_row = top_row, bot_row = bot_row,
      top_is_ref = (top_row == "Reference"),
      top_asm = if (top_row == "Reference") NA_character_ else top_row,
      bot_asm = bot_row
    )
  }

  # --- Pill lookup for ribbon building ---
  pill_lookup <- all_pills %>%
    select(row_name, ref_id, chrom_id, chrom_num, contig, contig_len, reversed,
           x_center, x_left, x_right, y_center, pill_width, subgenome)

  # --- Build ribbon polygons ---
  ribbon_polys <- list()
  ribbon_idx <- 0

  for (gap_i in seq_along(gap_zones)) {
    gz <- gap_zones[[gap_i]]
    gm <- gap_meta[[gap_i]]

    # Gather macro_blocks for this row pair
    if (gm$top_is_ref) {
      bot_asm <- gm$bot_asm
      if (!is.null(per_asm_macro_df)) {
        macro_src <- per_asm_macro_df %>%
          filter(assembly == bot_asm)
        if (n_subgenomes > 1) {
          sg_ref_ids <- sg_chr$ref_id
          macro_src <- macro_src %>% filter(ref_id %in% sg_ref_ids)
        }
      } else { macro_src <- NULL }
    } else {
      top_asm <- gm$top_asm
      bot_asm <- gm$bot_asm
      if (!is.null(pairwise_macro_df)) {
        macro_src <- pairwise_macro_df %>%
          filter(left_assembly == top_asm, right_assembly == bot_asm)
        if (n_subgenomes > 1) macro_src <- macro_src %>% filter(subgenome == sg)
      } else { macro_src <- NULL }
    }

    if (is.null(macro_src) || nrow(macro_src) == 0) next
    macro_src <- macro_src %>% filter(qspan_bp >= min_span)
    if (nrow(macro_src) == 0) next

    for (b in seq_len(nrow(macro_src))) {
      blk <- macro_src[b, ]

      if (gm$top_is_ref) {
        tp <- pill_lookup %>% filter(row_name == gm$top_row, contig == blk$ref_id)
        bp <- pill_lookup %>% filter(row_name == gm$bot_row, contig == blk$contig_new)
      } else {
        tp <- pill_lookup %>% filter(row_name == gm$top_row, contig == blk$ref_id)
        bp <- pill_lookup %>% filter(row_name == gm$bot_row, contig == blk$contig)
      }
      if (nrow(tp) == 0 || nrow(bp) == 0) next

      top_len <- tp$contig_len[1]; top_xl <- tp$x_left[1]; top_w <- tp$pill_width[1]
      bot_len <- bp$contig_len[1]; bot_xl <- bp$x_left[1]; bot_w <- bp$pill_width[1]
      cid <- as.character(tp$chrom_id[1])

      l_start <- as.numeric(blk$ref_start); l_end <- as.numeric(blk$ref_end)
      r_start <- as.numeric(blk$qstart);    r_end <- as.numeric(blk$qend)
      if (gm$top_is_ref && isTRUE(bp$reversed[1])) {
        r_tmp <- bot_len - r_end; r_end <- bot_len - r_start; r_start <- r_tmp
      }
      l_start <- max(0, l_start); l_end <- min(top_len, l_end)
      r_start <- max(0, r_start); r_end <- min(bot_len, r_end)
      if (l_end <= l_start || r_end <= r_start) next

      tx_left  <- pos_to_x(l_start, top_len, top_xl, top_w)
      tx_right <- pos_to_x(l_end,   top_len, top_xl, top_w)
      bx_left  <- pos_to_x(r_start, bot_len, bot_xl, bot_w)
      bx_right <- pos_to_x(r_end,   bot_len, bot_xl, bot_w)

      strand_val <- blk$strand
      if (gm$top_is_ref && isTRUE(bp$reversed[1])) {
        strand_val <- if (strand_val == "+") "-" else "+"
      }
      if (strand_val == "-") {
        tmp <- bx_left; bx_left <- bx_right; bx_right <- tmp
      }

      ident_val <- as.numeric(blk$identity)
      alpha_ribbon <- 0.8 + 0.15 * min(1, ident_val)

      tooltip <- sprintf(
        "%s \u2192 %s\n%.2f Mb\nIdentity: %.1f%%\nStrand: %s",
        tp$contig[1], bp$contig[1],
        as.numeric(blk$union_bp) / 1e6,
        ident_val * 100, strand_val
      )

      ribbon_idx <- ribbon_idx + 1
      ribbon_polys[[ribbon_idx]] <- make_ribbon_v(
        gz[1], gz[2], tx_left, tx_right, bx_left, bx_right,
        chrom_id_val = cid, alpha_val = alpha_ribbon,
        tooltip_text = tooltip, subgenome_val = sg,
        data_id_val = cid
      ) %>% mutate(group_id = ribbon_idx)
    }
  }

  # --- Rescue ribbons ---
  if (!is.null(rescue_macro_df) && nrow(rescue_macro_df) > 0) {
    rescue_pairs <- rescue_macro_df %>%
      distinct(left_assembly, right_assembly)
    for (rp_i in seq_len(nrow(rescue_pairs))) {
      r_top_asm <- rescue_pairs$left_assembly[rp_i]
      r_bot_asm <- rescue_pairs$right_assembly[rp_i]
      if (!(r_top_asm %in% row_names) || !(r_bot_asm %in% row_names)) next
      pair_df <- rescue_macro_df %>%
        filter(left_assembly == r_top_asm, right_assembly == r_bot_asm,
               qspan_bp >= min_span)
      if (n_subgenomes > 1) pair_df <- pair_df %>% filter(subgenome == sg)
      if (nrow(pair_df) == 0) next
      gz_rescue <- c(row_centers[[r_top_asm]] + 0.03,
                     row_centers[[r_bot_asm]] - 0.03)
      for (b in seq_len(nrow(pair_df))) {
        blk <- pair_df[b, ]
        tp <- pill_lookup %>% filter(row_name == r_top_asm, contig == blk$ref_id)
        bp <- pill_lookup %>% filter(row_name == r_bot_asm, contig == blk$contig)
        if (nrow(tp) == 0 || nrow(bp) == 0) next
        top_len <- tp$contig_len[1]; top_xl <- tp$x_left[1]; top_w <- tp$pill_width[1]
        bot_len <- bp$contig_len[1]; bot_xl <- bp$x_left[1]; bot_w <- bp$pill_width[1]
        cid <- as.character(tp$chrom_id[1])
        l_start <- max(0, as.numeric(blk$ref_start))
        l_end   <- min(top_len, as.numeric(blk$ref_end))
        r_start <- max(0, as.numeric(blk$qstart))
        r_end   <- min(bot_len, as.numeric(blk$qend))
        if (l_end <= l_start || r_end <= r_start) next
        tx_left  <- pos_to_x(l_start, top_len, top_xl, top_w)
        tx_right <- pos_to_x(l_end,   top_len, top_xl, top_w)
        bx_left  <- pos_to_x(r_start, bot_len, bot_xl, bot_w)
        bx_right <- pos_to_x(r_end,   bot_len, bot_xl, bot_w)
        strand_val <- blk$strand
        if (strand_val == "-") { tmp <- bx_left; bx_left <- bx_right; bx_right <- tmp }
        ident_val <- as.numeric(blk$identity)
        alpha_ribbon <- 0.8 + 0.15 * min(1, ident_val)
        tooltip <- sprintf(
          "%s \u2192 %s\n%.2f Mb\nIdentity: %.1f%%\nStrand: %s",
          tp$contig[1], bp$contig[1],
          as.numeric(blk$union_bp) / 1e6, ident_val * 100, strand_val
        )
        ribbon_idx <- ribbon_idx + 1
        ribbon_polys[[ribbon_idx]] <- make_ribbon_v(
          gz_rescue[1], gz_rescue[2], tx_left, tx_right, bx_left, bx_right,
          chrom_id_val = cid, alpha_val = alpha_ribbon,
          tooltip_text = tooltip, subgenome_val = sg,
          data_id_val = cid
        ) %>% mutate(group_id = ribbon_idx, is_rescue = TRUE)
      }
    }
  }

  has_ribbons <- length(ribbon_polys) > 0
  if (has_ribbons) {
    all_ribbons <- bind_rows(ribbon_polys)
    if (!"is_rescue" %in% names(all_ribbons)) all_ribbons$is_rescue <- FALSE
    all_ribbons$is_rescue[is.na(all_ribbons$is_rescue)] <- FALSE
    all_ribbons <- all_ribbons %>% arrange(desc(is_rescue), group_id)
  } else {
    all_ribbons <- NULL
  }

  # --- Plot sizing ---
  plot_w <- 7.2
  plot_h <- max(4, n_rows * 0.6 + 1)

  # --- Compose the plot ---
  p <- ggplot()

  # Ribbons (behind pills)
  if (has_ribbons) {
    p <- p +
      geom_polygon_interactive(
        data = all_ribbons,
        aes(x = x, y = y, group = group_id, fill = chrom_id,
            alpha = alpha, tooltip = tooltip, data_id = data_id),
        show.legend = FALSE
      ) +
      scale_fill_manual(values = chrom_pal, guide = "none") +
      scale_alpha_identity()
  }

  # Reference pills (darker)
  ref_pills <- all_pills %>% filter(is_ref)
  p <- p +
    geom_segment(
      data = ref_pills,
      aes(x = x_left, xend = x_right, y = y_center, yend = y_center),
      linewidth = pill_lw, lineend = "round", color = "grey55"
    )

  # Assembly pills (lighter, interactive)
  asm_pills <- all_pills %>% filter(!is_ref)
  if (nrow(asm_pills) > 0) {
    p <- p +
      geom_segment_interactive(
        data = asm_pills,
        aes(x = x_left, xend = x_right, y = y_center, yend = y_center,
            tooltip = tooltip, data_id = chrom_id),
        linewidth = pill_lw, lineend = "round", color = "grey78"
      )
  }

  # X-axis: chromosome labels
  chrom_breaks <- chrom_x[as.character(chrom_labels$chrom_num)]
  p <- p +
    scale_x_continuous(
      breaks = chrom_breaks,
      labels = chrom_labels$label,
      sec.axis = dup_axis(),
      expand = expansion(add = c(0.5, 0.5))
    )

  # Y-axis: row labels (Reference + assemblies)
  p <- p +
    scale_y_reverse(
      breaks = as.numeric(row_centers),
      labels = names(row_centers),
      expand = expansion(add = c(0.2, 0.2))
    )

  # Theme
  p <- p +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    theme(
      axis.text.x.bottom = element_text(size = base_font_pt, face = "bold",
                                       family = base_family, margin = margin(t = 2)),
      axis.text.x.top = element_text(size = base_font_pt, face = "bold",
                                      family = base_family, margin = margin(b = 2)),
      axis.text.y = element_text(size = base_font_pt, family = base_family, hjust = 1),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(5, 5, 2, 5)
    )

  list(plot = p, width = plot_w, height = plot_h, has_ribbons = has_ribbons)
}
```

```{r synteny_description, eval=has_synteny_data, results='asis'}
cat('<p class="section-intro">\n')
cat(sprintf("Chromosome synteny between the reference and each assembly (%s mode). ",
            params$synteny_mode))
if (n_subgenomes > 1) {
  cat("Each subgenome is shown in a separate tab. ")
}
cat("Ribbons connect syntenic macro-blocks between adjacent rows. ")
cat("Color indicates the reference chromosome; opacity reflects alignment identity. ")
cat("Crossed ribbons indicate inversions.\n")
cat("</p>\n")
```

```{r synteny_riparian, eval=has_synteny_data, results='asis'}
for (sg_i in seq_along(subgenome_order)) {
  sg <- subgenome_order[sg_i]

  # Emit tab header for polyploid
  if (n_subgenomes > 1) {
    cat(sprintf("\n## Subgenome %s\n\n", sg))
  }

  result <- build_synteny_plot(sg)
  if (is.null(result) || !result$has_ribbons) {
    cat("No ribbon connections could be built for this subgenome.\n\n")
    next
  }

  # Save PDF (one per subgenome)
  pdf_suffix <- if (n_subgenomes > 1) {
    paste0(".chromosome_synteny.", sg, ".pdf")
  } else {
    ".chromosome_synteny.pdf"
  }
  pdf_path <- paste0(params$out_prefix, pdf_suffix)
  save_pdf(pdf_path, result$plot, width = result$width, height = result$height)

  # Render interactive HTML widget
  g <- ggiraph::girafe(ggobj = result$plot,
                       width_svg = min(13, result$width + 2),
                       height_svg = result$height)
  g <- ggiraph::girafe_options(g,
    ggiraph::opts_tooltip(css = paste0(
      "font-family: ", base_family, "; font-size: 9pt; ",
      "background: white; padding: 6px; border: 1px solid #666;"
    )),
    ggiraph::opts_hover(css = "opacity: 0.8;"),
    ggiraph::opts_hover_inv(css = "opacity: 0.15;"),
    ggiraph::opts_toolbar(saveaspng = FALSE)
  )
  cat(as.character(htmltools::as.tags(g)))
  cat("\n\n")
}
```

```{r synteny_empty, eval=(has_synteny_data && length(subgenome_order) == 0)}
cat("Synteny data available but no ribbon connections could be built. ",
    "This may indicate that assemblies have very different chromosome structures.")
```


`r if (has_any_chrom && has_per_asm) "# 4. Quality Comparison {.tabset}" else ""`

```{r quality_section, eval=(has_any_chrom && has_per_asm)}
chrom_df <- all_summary_df %>%
  filter(classification == "chrom_assigned")

# Join subgenome info
if (has_subgenomes && "assigned_subgenome" %in% names(chrom_df)) {
  chrom_df <- chrom_df %>%
    mutate(sg = ifelse(is.na(assigned_subgenome) | assigned_subgenome == "NA",
                       "Other", assigned_subgenome))
} else {
  chrom_df$sg <- "All"
}

has_identity <- "seq_identity_vs_ref" %in% names(chrom_df) &&
                !all(is.na(chrom_df$seq_identity_vs_ref))
has_collin   <- "collinearity_score" %in% names(chrom_df) &&
                !all(is.na(chrom_df$collinearity_score))
has_gc_dev   <- "gc_deviation" %in% names(chrom_df) &&
                !all(is.na(chrom_df$gc_deviation))

# Compute composite sort key for assembly ordering
# Primary: median identity across first subgenome, secondary: others
asm_sort_key <- chrom_df %>%
  filter(!is.na(seq_identity_vs_ref)) %>%
  group_by(assembly) %>%
  summarise(median_ident = median(seq_identity_vs_ref, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(desc(median_ident))
asm_quality_order <- if (nrow(asm_sort_key) > 0) asm_sort_key$assembly else asm_names

n_sg_facets <- length(unique(chrom_df$sg))
```

```{r quality_identity, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), results='asis'}
cat("## Sequence Identity\n\n")
```

```{r quality_identity_plot, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
ident_df <- chrom_df %>%
  filter(!is.na(seq_identity_vs_ref)) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_quality_order)),
         value = seq_identity_vs_ref * 100)

p_ident <- ggplot(ident_df, aes(x = value, y = assembly, fill = sg, color = sg)) +
  geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                      rel_min_height = 0.01,
                      quantile_lines = TRUE, quantiles = 2) +
  { if (has_subgenomes && n_sg_facets > 1) {
    list(
      facet_wrap(~ sg, ncol = n_sg_facets),
      scale_fill_manual(values = sg_cols_light, guide = "none"),
      scale_color_manual(values = sg_cols_dark, guide = "none")
    )
  } else {
    list(
      scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
      scale_color_manual(values = c("All" = oi[5]), guide = "none")
    )
  }} +
  labs(x = "Sequence identity vs reference (%)", y = NULL) +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_ident, height_svg = max(3, n_asm * 0.3 + 1.5))
```

```{r quality_collinearity, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), results='asis'}
cat("## Collinearity\n\n")
```

```{r quality_collinearity_plot, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
collin_df <- chrom_df %>%
  filter(!is.na(collinearity_score)) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_quality_order)),
         value = collinearity_score * 100)

p_collin <- ggplot(collin_df, aes(x = value, y = assembly, fill = sg, color = sg)) +
  geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                      rel_min_height = 0.01,
                      quantile_lines = TRUE, quantiles = 2) +
  { if (has_subgenomes && n_sg_facets > 1) {
    list(
      facet_wrap(~ sg, ncol = n_sg_facets),
      scale_fill_manual(values = sg_cols_light, guide = "none"),
      scale_color_manual(values = sg_cols_dark, guide = "none")
    )
  } else {
    list(
      scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
      scale_color_manual(values = c("All" = oi[5]), guide = "none")
    )
  }} +
  labs(x = "Collinearity (%)", y = NULL) +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_collin, height_svg = max(3, n_asm * 0.3 + 1.5))
```

```{r quality_gc, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), results='asis'}
cat("## GC Deviation\n\n")
```

```{r quality_gc_plot, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
gc_df <- chrom_df %>%
  filter(!is.na(gc_deviation)) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_quality_order)),
         value = abs(gc_deviation))

p_gc <- ggplot(gc_df, aes(x = value, y = assembly, fill = sg, color = sg)) +
  geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                      rel_min_height = 0.01,
                      quantile_lines = TRUE, quantiles = 2) +
  { if (has_subgenomes && n_sg_facets > 1) {
    list(
      facet_wrap(~ sg, ncol = n_sg_facets),
      scale_fill_manual(values = sg_cols_light, guide = "none"),
      scale_color_manual(values = sg_cols_dark, guide = "none")
    )
  } else {
    list(
      scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
      scale_color_manual(values = c("All" = oi[5]), guide = "none")
    )
  }} +
  labs(x = "|GC deviation| (std devs from reference)", y = NULL) +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_gc, height_svg = max(3, n_asm * 0.3 + 1.5))
```


`r if (has_any_depth) "# 5. Read Depth" else ""`

```{r depth_section, eval=(has_any_depth && has_per_asm), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
depth_df <- all_summary_df %>%
  filter(classification == "chrom_assigned", !is.na(depth_mean))

if (nrow(depth_df) > 0) {
  # Add subgenome info
  if (has_subgenomes && "assigned_subgenome" %in% names(depth_df)) {
    depth_df <- depth_df %>%
      mutate(sg = ifelse(is.na(assigned_subgenome) | assigned_subgenome == "NA",
                         "Other", assigned_subgenome))
  } else {
    depth_df$sg <- "All"
  }

  depth_df <- depth_df %>%
    mutate(assembly = factor(assembly, levels = rev(asm_names)))

  n_depth_sg <- length(unique(depth_df$sg))

  p_depth <- ggplot(depth_df, aes(x = depth_mean, y = assembly, fill = sg, color = sg)) +
    geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                        rel_min_height = 0.01,
                        quantile_lines = TRUE, quantiles = 2) +
    { if (has_subgenomes && n_depth_sg > 1) {
      list(
        facet_wrap(~ sg, ncol = n_depth_sg),
        scale_fill_manual(values = sg_cols_light, guide = "none"),
        scale_color_manual(values = sg_cols_dark, guide = "none")
      )
    } else {
      list(
        scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
        scale_color_manual(values = c("All" = oi[5]), guide = "none")
      )
    }} +
    labs(x = "Mean read depth (chromosome contigs)", y = NULL) +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(panel.grid.major.y = element_blank())

  girafe_styled(ggobj = p_depth, height_svg = max(3, n_asm * 0.3 + 1.5))
} else {
  cat("No depth data available for chromosome-assigned contigs.")
}
```


`r if (has_per_asm && any(comp_df$n_rdna > 0, na.rm = TRUE)) "# 6. rDNA Comparison {.tabset}" else ""`

```{r rdna_setup, eval=(has_per_asm && any(comp_df$n_rdna > 0, na.rm = TRUE)), include=FALSE}
has_any_rdna <- any(comp_df$n_rdna > 0, na.rm = TRUE)
```

```{r rdna_summary_header, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna), results='asis'}
cat("## Summary\n\n")
```

```{r rdna_summary, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna)}
rdna_tbl <- comp_df %>%
  transmute(
    Assembly = assembly,
    rdna_mb = round(rdna_bp / 1e6, 2),
    `Contigs` = n_rdna,
    `Arrays` = n_rdna_arrays
  )

max_rdna_mb <- max(rdna_tbl$rdna_mb, na.rm = TRUE)

rdna_tbl %>%
  rowwise() %>%
  mutate(
    `Total (Mb)` = {
      pct <- if (max_rdna_mb > 0) min(100, rdna_mb / max_rdna_mb * 100) else 0
      bar_cell(pct, oi[1], sprintf("%.2f", rdna_mb), alpha = 0.3)
    }
  ) %>%
  ungroup() %>%
  select(Assembly, `Total (Mb)`, Contigs, Arrays) %>%
  gt() %>%
  fmt(columns = `Total (Mb)`, fns = function(x) lapply(x, html)) %>%
  cols_align(align = "right", columns = c(Contigs, Arrays)) %>%
  cols_width(`Total (Mb)` ~ px(140)) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  tab_header(title = md("**rDNA Summary**")) %>%
  gt_theme()
```

```{r rdna_arrays_header, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna && has_rdna_arrays), results='asis'}
cat("## Array Detail\n\n")
```

```{r rdna_arrays_table, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna && has_rdna_arrays)}
arr_tbl <- all_rdna_arrays_df %>%
  transmute(
    Assembly = assembly,
    `Array ID` = array_id,
    Contig = contig,
    `N copies` = n_total,
    Strand = strand,
    `Span (kb)` = span_kb
  )

arr_tbl %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = md("**rDNA Array Details**")) %>%
  cols_align(align = "right", columns = c(`N copies`, `Span (kb)`))
```


`r if (has_per_asm && any(comp_df$chrC_found == "yes" | comp_df$chrM_found == "yes", na.rm = TRUE)) "# 7. Organelle Comparison" else ""`

```{r organelle_section, eval=(has_per_asm && any(comp_df$chrC_found == "yes" | comp_df$chrM_found == "yes", na.rm = TRUE))}
# Extract organelle contigs from per-assembly data
org_df <- all_summary_df %>%
  filter(classification == "organelle_complete") %>%
  mutate(
    org_type = case_when(
      grepl("chrC|Pt|chloro|plast", assigned_ref_id, ignore.case = TRUE) ~ "chrC",
      grepl("chrM|Mt|mito", assigned_ref_id, ignore.case = TRUE) ~ "chrM",
      TRUE ~ "Other"
    )
  ) %>%
  group_by(assembly, org_type) %>%
  summarise(
    n_contigs = n(),
    total_bp = sum(length, na.rm = TRUE),
    contigs = paste(contig, collapse = ", "),
    .groups = "drop"
  ) %>%
  mutate(
    assembly = as.character(assembly),
    total_mb = total_bp / 1e6
  )

# Build wide table with bar cells for chrC and chrM
org_wide <- comp_df %>%
  transmute(Assembly = assembly) %>%
  left_join(
    org_df %>% filter(org_type == "chrC") %>%
      transmute(Assembly = assembly, chrC_bp = total_bp, chrC_contigs = contigs),
    by = "Assembly"
  ) %>%
  left_join(
    org_df %>% filter(org_type == "chrM") %>%
      transmute(Assembly = assembly, chrM_bp = total_bp, chrM_contigs = contigs),
    by = "Assembly"
  ) %>%
  mutate(
    chrC_bp = replace_na(chrC_bp, 0),
    chrM_bp = replace_na(chrM_bp, 0),
    chrC_contigs = replace_na(chrC_contigs, "\u2014"),
    chrM_contigs = replace_na(chrM_contigs, "\u2014")
  )

max_chrC_bp <- max(org_wide$chrC_bp, na.rm = TRUE)
max_chrM_bp <- max(org_wide$chrM_bp, na.rm = TRUE)

# chrM unit: Kb if max >= 1 Mb, else plain bp with commas
chrM_use_kb <- max_chrM_bp >= 1e6
chrM_col_label <- if (chrM_use_kb) "Size (Kb)" else "Size (bp)"

# Chloroplast color: green-tinted, Mitochondria: red-tinted

org_gt_df <- org_wide %>%
  rowwise() %>%
  mutate(
    chrC_size = if (chrC_bp > 0) {
      pct <- min(100, chrC_bp / max_chrC_bp * 100)
      bar_cell(pct, oi[3], formatC(chrC_bp, format = "d", big.mark = ","))
    } else "\u2014",
    chrM_size = if (chrM_bp > 0) {
      pct <- min(100, chrM_bp / max_chrM_bp * 100)
      val_text <- if (chrM_use_kb) {
        formatC(round(chrM_bp / 1e3), format = "d", big.mark = ",")
      } else {
        formatC(chrM_bp, format = "d", big.mark = ",")
      }
      bar_cell(pct, oi[4], val_text)
    } else "\u2014",
    chrC_contig = chrC_contigs,
    chrM_contig = chrM_contigs
  ) %>%
  ungroup() %>%
  select(Assembly, chrC_contig, chrC_size, chrM_contig, chrM_size)

# Dynamic column labels
chrC_size_label <- "Size (bp)"

# Detect plastid visibility (reuse from overview table if available)
show_chrC_org <- exists("show_chrC") && show_chrC

if (show_chrC_org) {
  org_gt_df %>%
    gt() %>%
    fmt(columns = c(chrC_size, chrM_size), fns = function(x) lapply(x, html)) %>%
    cols_label(
      chrC_contig = "Contig", chrC_size = chrC_size_label,
      chrM_contig = "Contig", chrM_size = chrM_col_label
    ) %>%
    cols_width(chrC_size ~ px(140), chrM_size ~ px(140)) %>%
    tab_spanner(label = "Chloroplast", columns = c(chrC_contig, chrC_size)) %>%
    tab_spanner(label = "Mitochondria", columns = c(chrM_contig, chrM_size)) %>%
    tab_style(
      style = css(vertical_align = "middle"),
      locations = cells_body()
    ) %>%
    tab_header(title = md("**Organelle Summary**")) %>%
    gt_theme()
} else {
  org_gt_df %>%
    select(Assembly, chrM_contig, chrM_size) %>%
    gt() %>%
    fmt(columns = chrM_size, fns = function(x) lapply(x, html)) %>%
    cols_label(chrM_contig = "Contig", chrM_size = chrM_col_label) %>%
    cols_width(chrM_size ~ px(140)) %>%
    tab_spanner(label = "Mitochondria", columns = c(chrM_contig, chrM_size)) %>%
    tab_style(
      style = css(vertical_align = "middle"),
      locations = cells_body()
    ) %>%
    tab_header(title = md("**Organelle Summary**")) %>%
    gt_theme()
}
```


`r if (has_any_contam) "# 8. Contamination Comparison {.tabset}" else ""`

`r if (has_any_contam) "## Summary" else ""`

```{r contamination, eval=has_any_contam}
contam_tbl <- comp_df %>%
  filter(n_contaminants > 0) %>%
  transmute(
    Assembly = assembly,
    contam_mb = round(contaminant_bp / 1e6, 2),
    Contigs = n_contaminants,
    Species = n_contaminant_species
  )

max_contam_mb <- max(contam_tbl$contam_mb, na.rm = TRUE)

contam_tbl %>%
  rowwise() %>%
  mutate(
    `Total (Mb)` = {
      pct <- if (max_contam_mb > 0) min(100, contam_mb / max_contam_mb * 100) else 0
      bar_cell(pct, oi[7], sprintf("%.2f", contam_mb))
    }
  ) %>%
  ungroup() %>%
  select(Assembly, `Total (Mb)`, Contigs, Species) %>%
  gt() %>%
  fmt(columns = `Total (Mb)`, fns = function(x) lapply(x, html)) %>%
  cols_align(align = "right", columns = c(Contigs, Species)) %>%
  cols_width(`Total (Mb)` ~ px(140)) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  tab_header(title = md("**Contamination Summary**")) %>%
  gt_theme()
```

`r if (has_any_contam && has_contam_detail) "## Top Taxa" else ""`

```{r top_taxa, eval=(has_any_contam && has_contam_detail), results='asis'}
# --- Top genera by prevalence across assemblies ---
genus_prevalence <- all_contam_detail_df %>%
  filter(nzchar(genus) & !is.na(genus)) %>%
  distinct(assembly, genus) %>%
  group_by(genus) %>%
  summarise(n_assemblies = n(),
            asm_list = paste(sort(assembly), collapse = "\n"),
            .groups = "drop") %>%
  mutate(pct = n_assemblies / n_asm * 100) %>%
  arrange(desc(n_assemblies), genus) %>%
  head(10)

genus_asm_tips <- genus_prevalence$asm_list

genus_gt <- genus_prevalence %>%
  rowwise() %>%
  mutate(bar = bar_cell(pct, oi[7], sprintf("%.0f%%", pct), alpha = 0.3)) %>%
  ungroup() %>%
  transmute(Genus = genus, `% assemblies` = bar, n = n_assemblies) %>%
  gt() %>%
  fmt(columns = `% assemblies`, fns = function(x) lapply(x, html)) %>%
  text_transform(
    locations = cells_body(columns = Genus),
    fn = function(x) sprintf(
      '<span title="%s" style="white-space:nowrap;overflow:hidden;text-overflow:ellipsis;display:block;max-width:120px"><i>%s</i> spp.</span>',
      x, x)
  ) %>%
  text_transform(
    locations = cells_body(columns = n),
    fn = function(x) mapply(
      function(v, tip) sprintf('<span title="%s">%s</span>', gsub('"', '&quot;', tip), v),
      x, genus_asm_tips[seq_along(x)], SIMPLIFY = TRUE)
  ) %>%
  cols_align(align = "right", columns = n) %>%
  cols_width(Genus ~ px(120), `% assemblies` ~ px(150), n ~ px(40)) %>%
  tab_header(title = md("**Top Genera**")) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()

# --- Top species by prevalence across assemblies ---
species_prevalence <- all_contam_detail_df %>%
  filter(nzchar(species) & !is.na(species)) %>%
  mutate(binomial = paste(genus, species)) %>%
  distinct(assembly, binomial) %>%
  group_by(binomial) %>%
  summarise(n_assemblies = n(),
            asm_list = paste(sort(assembly), collapse = "\n"),
            .groups = "drop") %>%
  mutate(pct = n_assemblies / n_asm * 100) %>%
  arrange(desc(n_assemblies), binomial) %>%
  head(10)

species_asm_tips <- species_prevalence$asm_list

species_gt <- species_prevalence %>%
  rowwise() %>%
  mutate(bar = bar_cell(pct, oi[7], sprintf("%.0f%%", pct), alpha = 0.3)) %>%
  ungroup() %>%
  transmute(Species = binomial, `% assemblies` = bar, n = n_assemblies) %>%
  gt() %>%
  fmt(columns = `% assemblies`, fns = function(x) lapply(x, html)) %>%
  text_transform(
    locations = cells_body(columns = Species),
    fn = function(x) sprintf(
      '<span title="%s" style="white-space:nowrap;overflow:hidden;text-overflow:ellipsis;display:block;max-width:200px"><i>%s</i></span>',
      x, x)
  ) %>%
  text_transform(
    locations = cells_body(columns = n),
    fn = function(x) mapply(
      function(v, tip) sprintf('<span title="%s">%s</span>', gsub('"', '&quot;', tip), v),
      x, species_asm_tips[seq_along(x)], SIMPLIFY = TRUE)
  ) %>%
  cols_align(align = "right", columns = n) %>%
  cols_width(Species ~ px(200), `% assemblies` ~ px(150), n ~ px(40)) %>%
  tab_header(title = md("**Top Species**")) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()

# Render side by side using a two-column HTML layout
cat('<div style="display: flex; gap: 12px; align-items: flex-start; flex-wrap: wrap;">\n')
cat('<div style="flex: 1; min-width: 300px;">\n')
cat(as.character(as_raw_html(genus_gt)))
cat('\n</div>\n')
cat('<div style="flex: 1; min-width: 300px;">\n')
cat(as.character(as_raw_html(species_gt)))
cat('\n</div>\n')
cat('</div>\n')
```


---

<div class="note-box">
Generated by **final_finalizer** multi-assembly comparison &mdash; `r Sys.Date()`
</div>
