---
title: "Assembly Finalization Report"
params:
  summary_tsv: "__SUMMARY__"
  ref_lengths_tsv: "__REF__"
  segments_tsv: "__SEGMENTS__"
  evidence_tsv: "__EVIDENCE__"
  macro_blocks_tsv: "__MACRO__"
  rdna_annotations_tsv: "__RDNA_ANNOTATIONS__"
  rdna_arrays_tsv: "__RDNA_ARRAYS__"
  contaminants_tsv: "__CONTAMINANTS_TSV__"
  agp_tsv: "__AGP__"
  compleasm_chrs_summary: "__COMPLEASM_CHRS__"
  compleasm_non_chrs_summary: "__COMPLEASM_NONCHRS__"
  rearrangements_tsv: "__REARRANGEMENTS__"
  assembly_name: "__ASMNAME__"
  reference_name: "__REFNAME__"
  synteny_mode: "__SYNTENY_MODE__"
  chr_like_minlen: "__CHRLIKE__"
  plot_suffix: "__SUFFIX__"
  out_prefix: "__OUTPREFIX__"
  top_n_contaminants: "__TOP_N__"
output:
  html_document:
    self_contained: __SELF_CONTAINED__
    css: "__COMMON_CSS__"
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, dpi = 150)

source("__COMMON_R__")

# ---------------------------------------------------------------------------
# Subgenome detection and color setup
# ---------------------------------------------------------------------------
ref_lengths <- read_tsv(params$ref_lengths_tsv, show_col_types = FALSE,
                        col_types = cols(subgenome = col_character()))
list2env(setup_subgenomes(ref_lengths), envir = environment())

# For depth overview
sg_colors <- if (n_sg > 0) setNames(pal_sg_dark[seq_len(n_sg)], sg_levels) else c("G" = oi[5])
sg_colors <- c(sg_colors, "NA" = pal_sg_dark[1])

# ---------------------------------------------------------------------------
# Unified-report-specific helpers
# ---------------------------------------------------------------------------
gt_inline_bar <- function(value, max_val, color = oi[5], height = "14px") {
  if (is.na(value) || is.na(max_val) || max_val == 0) return("")
  pct <- min(100, round(value / max_val * 100, 1))
  paste0(
    '<div style="position:relative;width:60px;height:', height,
    ';border-radius:3px;display:inline-block;background:#f0f0f0;">',
    sprintf('<div style="position:absolute;inset:0;width:%.1f%%;background-color:%s;border-radius:3px;"></div>', pct, oi_rgba(color, 0.5)),
    '</div>')
}

conf_badge <- function(level) {
  cols <- c(high = "#27ae60", medium = "#f39c12", low = "#e74c3c")
  bg <- cols[level]; if (is.na(bg)) bg <- "#999"
  paste0('<span style="background:', bg, '; color:white; padding:1px 6px; ',
         'border-radius:3px; font-size:10px;">', level, '</span>')
}

version_sort_key <- function(x) {
  num <- as.integer(str_extract(x, "\\d+"))
  num[is.na(num)] <- 99999L
  num
}

# ---------------------------------------------------------------------------
# Load main data
# ---------------------------------------------------------------------------
summary_df <- read_tsv(params$summary_tsv, show_col_types = FALSE,
                       col_types = cols(assigned_subgenome = col_character()))

has_evidence <- nchar(params$evidence_tsv) > 0 && file.exists(params$evidence_tsv)
evidence_df <- if (has_evidence) read_tsv(params$evidence_tsv, show_col_types = FALSE) else NULL

has_segments <- nchar(params$segments_tsv) > 0 && file.exists(params$segments_tsv)
has_macro    <- nchar(params$macro_blocks_tsv) > 0 && file.exists(params$macro_blocks_tsv)

has_rdna_ann <- nchar(params$rdna_annotations_tsv) > 0 && file.exists(params$rdna_annotations_tsv)
rdna_ann_df  <- if (has_rdna_ann) read_tsv(params$rdna_annotations_tsv, show_col_types = FALSE) else NULL

has_rdna_arrays <- nchar(params$rdna_arrays_tsv) > 0 && file.exists(params$rdna_arrays_tsv)
rdna_arrays_df  <- if (has_rdna_arrays) read_tsv(params$rdna_arrays_tsv, show_col_types = FALSE) else NULL
has_rdna_arrays <- !is.null(rdna_arrays_df) && nrow(rdna_arrays_df) > 0

has_contaminants <- nchar(params$contaminants_tsv) > 0 && file.exists(params$contaminants_tsv)
contam_df <- if (has_contaminants) {
  tmp <- read_tsv(params$contaminants_tsv, show_col_types = FALSE)
  if (nrow(tmp) > 0) tmp else NULL
} else NULL
has_contaminants <- !is.null(contam_df)

has_rearrangements <- nchar(params$rearrangements_tsv) > 0 && file.exists(params$rearrangements_tsv)
rearr_df <- if (has_rearrangements) {
  tmp <- read_tsv(params$rearrangements_tsv, show_col_types = FALSE)
  if (nrow(tmp) > 0) tmp else NULL
} else NULL
has_rearrangements <- !is.null(rearr_df)

has_depth <- "depth_mean" %in% names(summary_df) && !all(is.na(summary_df$depth_mean))
has_collinearity <- "collinearity_score" %in% names(summary_df)
has_chrom_assigned <- sum(summary_df$classification == "chrom_assigned", na.rm = TRUE) > 0

chr_like_minlen <- as.numeric(params$chr_like_minlen)
top_n_contam    <- as.integer(params$top_n_contaminants)

# PDF output paths (derived from out_prefix)
pdf_chrom_overview <- paste0(params$out_prefix, ".chromosome_overview.pdf")
pdf_classbar       <- paste0(params$out_prefix, ".classification_summary_bar.pdf")
pdf_depth          <- paste0(params$out_prefix, ".depth_overview.pdf")
```

---

<p class="section-intro">
**Assembly**: `r if (nchar(params$assembly_name) > 0) params$assembly_name else "\u2014"` &emsp;
**Reference**: `r if (nchar(params$reference_name) > 0) params$reference_name else "\u2014"` &emsp;
**Synteny mode**: `r params$synteny_mode` &emsp;
**Date**: `r Sys.Date()`
</p>

```{r compleasm_early_setup, include=FALSE}
parse_compleasm <- function(path) {
  if (is.null(path) || !nzchar(path) || !file.exists(path)) return(NULL)
  txt <- readLines(path, warn = FALSE)
  lineage <- sub("## lineage:\\s*", "", grep("^## lineage:", txt, value = TRUE)[1])
  vals <- list()
  for (code in c("S", "D", "F", "I", "M")) {
    m <- regmatches(txt, regexpr(paste0(code, ":([0-9.]+)%,\\s*(\\d+)"), txt))
    m <- m[nzchar(m)]
    if (length(m) > 0) {
      parts <- regmatches(m[1], regexec(paste0(code, ":([0-9.]+)%,\\s*(\\d+)"), m[1]))[[1]]
      vals[[code]] <- list(pct = as.numeric(parts[2]), n = as.integer(parts[3]))
    } else {
      vals[[code]] <- list(pct = 0, n = 0L)
    }
  }
  n_match <- regmatches(txt, regexpr("N:\\d+", txt))
  n_match <- n_match[nzchar(n_match)]
  n_total <- if (length(n_match) > 0) as.integer(sub("N:", "", n_match[1])) else 0L
  tibble(
    lineage = lineage, n_total = n_total,
    `Single-copy` = vals$S$pct, Duplicated = vals$D$pct,
    Fragmented = vals$F$pct, Interspersed = vals$I$pct, Missing = vals$M$pct,
    n_S = vals$S$n, n_D = vals$D$n, n_F = vals$F$n, n_I = vals$I$n, n_M = vals$M$n
  )
}
compleasm_chrs <- parse_compleasm(params$compleasm_chrs_summary)
compleasm_non <- parse_compleasm(params$compleasm_non_chrs_summary)
has_compleasm <- !is.null(compleasm_chrs) || !is.null(compleasm_non)
```

# 1. Assembly Overview {.tabset}

```{r overview_data}
clf_agg <- summary_df %>%
  mutate(length_mb = length / 1e6) %>%
  group_by(classification) %>%
  summarise(n_contigs = n(), total_mb = sum(length_mb, na.rm = TRUE), .groups = "drop") %>%
  mutate(pct_assembly = total_mb / sum(total_mb) * 100,
         label = classification_labels[classification]) %>%
  arrange(desc(total_mb))

total_mb <- sum(clf_agg$total_mb)
total_contigs <- sum(clf_agg$n_contigs)
```

## Classification

```{r classbar_build, include=FALSE}
# ---------------------------------------------------------------------------
# Classification summary bar (migrated from classification_summary_bar.tmpl.R)
# Waterfall: top bar = full assembly, bottom bar = zoomed non-chromosome
# ---------------------------------------------------------------------------
bar_label_font_pt <- 8
cb_label_font_pt  <- 6.5
cb_annot_font_pt  <- 7

cb_agg <- summary_df %>%
  mutate(classification = factor(classification, levels = classification_levels),
         length_mb = length / 1e6) %>%
  group_by(classification) %>%
  summarise(n_contigs = n(), total_mb = sum(length_mb, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(classification))

cb_total_contigs <- sum(cb_agg$n_contigs)
cb_total_mb      <- sum(cb_agg$total_mb)

cb_agg <- cb_agg %>%
  mutate(pct_contigs = n_contigs / cb_total_contigs * 100,
         pct_mb      = total_mb / cb_total_mb * 100) %>%
  arrange(classification != "chrom_assigned", desc(total_mb)) %>%
  mutate(xmax = cumsum(pct_mb), xmin = xmax - pct_mb, x_center = (xmin + xmax) / 2)

# Subtitle
stats_line <- paste0(comma(cb_total_mb, accuracy = 0.1), " Mb  n=",
                     comma(cb_total_contigs), " contigs")
cb_subtitle <- if (nzchar(params$assembly_name)) {
  paste0(params$assembly_name, "\n", stats_line)
} else stats_line

# Top bar
bar1_ymin <- 0.60; bar1_ymax <- 0.90

cb_bar1 <- cb_agg %>%
  filter(pct_mb > 0) %>%
  mutate(tooltip = paste0(
    clf_labels_short[as.character(classification)],
    "\n", comma(total_mb, accuracy = 0.1), " Mb (", sprintf("%.1f%%", pct_mb), ")",
    "\nn=", n_contigs, " (", sprintf("%.1f%%", pct_contigs), ")"
  ))

cb_chr <- cb_agg %>% filter(classification == "chrom_assigned" & pct_mb > 0)
cb_other <- cb_agg %>% filter(classification != "chrom_assigned" & pct_mb > 0)
chr_xmax <- if (nrow(cb_chr) > 0) cb_chr$xmax[1] else 0

cb_chr_label <- cb_chr %>%
  mutate(label_text = paste0(
    clf_labels_short[as.character(classification)], "\n",
    comma(total_mb, accuracy = 0.1), " Mb (", sprintf("%.1f%%", pct_mb), ")\n",
    "n=", comma(n_contigs), " (", sprintf("%.1f%%", pct_contigs), ")"
  ))

# Bottom bar
bar2_ymin <- 0.1925; bar2_ymax <- 0.2675

if (nrow(cb_other) > 0) {
  cb_other_total <- sum(cb_other$total_mb)
  cb_other_pct   <- sum(cb_other$pct_mb)

  cb_bar2 <- cb_other %>%
    mutate(pct_within = total_mb / cb_other_total * 100) %>%
    arrange(desc(total_mb)) %>%
    mutate(xmax2 = cumsum(pct_within), xmin2 = xmax2 - pct_within,
           x_center2 = (xmin2 + xmax2) / 2,
           tooltip = paste0(
             clf_labels_short[as.character(classification)],
             "\n", comma(total_mb, accuracy = 0.1), " Mb (", sprintf("%.1f%%", pct_mb), ") of assembly",
             "\nn=", n_contigs, " (", sprintf("%.1f%%", pct_contigs), " of assembly)"
           ),
           label_text = paste0(
             clf_labels_short[as.character(classification)], "\n",
             comma(total_mb, accuracy = 0.1), " Mb (", sprintf("%.1f%%", pct_mb), ")\n",
             "n=", comma(n_contigs), " (", sprintf("%.1f%%", pct_contigs), ")"
           ))

  # S-curve connector ribbons
  n_curve <- 20; t_vals <- seq(0, 1, length.out = n_curve)
  cb_connectors <- bind_rows(lapply(seq_len(nrow(cb_bar2)), function(i) {
    row <- cb_bar2[i, ]
    s <- 0.5 - 0.5 * cos(pi * t_vals)
    left_x  <- row$xmin  + s * (row$xmin2 - row$xmin)
    right_x <- row$xmax + s * (row$xmax2 - row$xmax)
    y_seq   <- bar1_ymin + t_vals * (bar2_ymax - bar1_ymin)
    tibble(x = c(left_x, rev(right_x)), y = c(y_seq, rev(y_seq)),
           classification = as.character(row$classification))
  }))
} else {
  cb_bar2 <- tibble(); cb_other_pct <- 0; cb_connectors <- tibble()
}

# Shared ggrepel params for bottom bar labels
cb_repel <- list(
  size = cb_label_font_pt / .pt, family = base_family, lineheight = 0.95,
  direction = "x", nudge_y = -0.5, segment.color = "grey70", segment.size = 0.25,
  segment.curvature = -1e-20, segment.shape = 1, segment.ncp = 1,
  segment.angle = 90, segment.inflect = TRUE, min.segment.length = 0,
  box.padding = 0.25, point.padding = 0.75, point.size = 2,
  force = 1.5, force_pull = 0, max.overlaps = Inf,
  ylim = c(-Inf, bar2_ymin - 0.02), seed = 42
)

# Subtitle-to-bar ribbon
cb_y_min <- -0.35; cb_y_max <- 1.10
rib_n <- 20; rib_t <- seq(0, 1, length.out = rib_n)
rib_s <- 0.5 - 0.5 * cos(pi * rib_t)
rib_left  <- (50 - 2) + rib_s * (0   - (50 - 2))
rib_right <- (50 + 2) + rib_s * (100 - (50 + 2))
rib_y     <- (cb_y_max - 0.01) + rib_t * (bar1_ymax - (cb_y_max - 0.01))
cb_ribbon <- data.frame(x = c(rib_left, rev(rib_right)), y = c(rib_y, rev(rib_y)))

# Build the plot (interactive version -- works for both PDF and girafe)
p_classbar <- ggplot() +
  geom_polygon(data = cb_ribbon, aes(x = x, y = y), fill = "grey95", color = NA) +
  { if (nrow(cb_bar2) > 0)
      geom_polygon(data = cb_connectors,
                   aes(x = x, y = y, fill = classification, group = classification),
                   alpha = 0.25, color = NA) } +
  ggiraph::geom_rect_interactive(
    data = cb_bar1,
    aes(xmin = xmin, xmax = xmax, ymin = bar1_ymin, ymax = bar1_ymax,
        fill = classification, tooltip = tooltip),
    color = "white", linewidth = 0.3
  ) +
  { if (nrow(cb_chr_label) > 0)
      geom_text(data = cb_chr_label,
                aes(x = x_center, y = (bar1_ymin + bar1_ymax) / 2, label = label_text),
                size = bar_label_font_pt / .pt, family = base_family,
                lineheight = 0.95, color = "white") } +
  { if (nrow(cb_bar2) > 0)
      annotate("text", x = 0, y = bar2_ymax + 0.08,
               label = paste0("Remaining ", sprintf("%.1f%%", cb_other_pct), " of assembly:"),
               size = cb_annot_font_pt / .pt, family = base_family, fontface = "italic",
               color = "grey30", hjust = 0) } +
  { if (nrow(cb_bar2) > 0)
      ggiraph::geom_rect_interactive(
        data = cb_bar2,
        aes(xmin = xmin2, xmax = xmax2, ymin = bar2_ymin, ymax = bar2_ymax,
            fill = classification, tooltip = tooltip),
        color = NA) } +
  { if (nrow(cb_bar2) > 0)
      do.call(ggrepel::geom_text_repel, c(
        list(data = cb_bar2,
             mapping = aes(x = x_center2, y = bar2_ymin, label = label_text)),
        cb_repel)) } +
  scale_fill_manual(values = classification_colors, labels = clf_labels_short, drop = FALSE) +
  scale_x_continuous(expand = c(0.01, 0.01), limits = c(0, 100)) +
  scale_y_continuous(limits = c(cb_y_min, cb_y_max), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  theme_void(base_family = base_family) +
  theme(plot.title = element_text(hjust = 0.5, size = 11, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 9, color = "grey20"),
        legend.position = "none",
        plot.margin = margin(t = 10, r = 15, b = 15, l = 15)) +
  labs(title = "Contig classification", subtitle = cb_subtitle)

save_pdf(pdf_classbar, p_classbar, width = 7.2, height = 3.0)
```

```{r classbar_widget}
girafe_styled(p_classbar, width_svg = 10, height_svg = 4.2)
```

```{r classification_table}
max_mb <- max(clf_agg$total_mb, na.rm = TRUE)

clf_agg %>%
  mutate(
    bar = mapply(function(v) gt_inline_bar(v, max_mb), total_mb),
    color_swatch = mapply(function(c) {
      col <- classification_colors[c]; if (is.na(col)) col <- "#999"
      paste0('<span style="display:inline-block;width:10px;height:10px;background:', col,
             ';border-radius:2px;margin-right:4px;"></span>')
    }, classification),
    label = paste0(color_swatch, label)
  ) %>%
  select(label, n_contigs, total_mb, pct_assembly, bar) %>%
  gt() %>%
  cols_label(label = "Category", n_contigs = "Contigs", total_mb = "Total Mb",
             pct_assembly = "% Assembly", bar = "") %>%
  fmt_number(columns = total_mb, decimals = 1) %>%
  fmt_number(columns = pct_assembly, decimals = 1) %>%
  fmt(columns = label, fns = function(x) x) %>%
  fmt(columns = bar, fns = function(x) x) %>%
  tab_header(title = md("**Classification Breakdown**"),
             subtitle = md(paste0(total_contigs, " contigs | ", sprintf("%.1f", total_mb), " Mb total"))) %>%
  gt_theme()
```

## Assembly Stats

```{r assembly_stats}
lengths_sorted <- sort(summary_df$length, decreasing = TRUE)
total_bp <- sum(lengths_sorted)
cumsum_bp <- cumsum(lengths_sorted)
n50_idx <- which(cumsum_bp >= total_bp / 2)[1]
n50 <- lengths_sorted[n50_idx]; l50 <- n50_idx

chr_df_stats <- summary_df %>% filter(classification == "chrom_assigned")
n_assigned <- nrow(chr_df_stats)
n_unassigned <- sum(summary_df$classification == "chrom_unassigned")
n_chimeric <- sum(chr_df_stats$chimeric == "yes", na.rm = TRUE)
n_full_length <- sum(chr_df_stats$is_full_length == "yes", na.rm = TRUE)
n_both_telo <- sum(chr_df_stats$has_5p_telomere == "yes" & chr_df_stats$has_3p_telomere == "yes", na.rm = TRUE)

mean_ref_cov <- if ("ref_coverage" %in% names(chr_df_stats) && any(!is.na(chr_df_stats$ref_coverage)))
  mean(as.numeric(chr_df_stats$ref_coverage), na.rm = TRUE) else NA
mean_ident <- if ("seq_identity_vs_ref" %in% names(chr_df_stats) && any(!is.na(chr_df_stats$seq_identity_vs_ref)))
  mean(as.numeric(chr_df_stats$seq_identity_vs_ref), na.rm = TRUE) else NA

n_contaminants <- sum(summary_df$classification == "contaminant")
n_rdna <- sum(summary_df$classification == "rDNA")
has_chrC <- any(grepl("chrC|Pt|chloro|plast", summary_df$assigned_ref_id[summary_df$classification == "organelle_complete"], ignore.case = TRUE))
has_chrM <- any(grepl("chrM|Mt|mito", summary_df$assigned_ref_id[summary_df$classification == "organelle_complete"], ignore.case = TRUE))

tibble(
  Contigs = as.integer(length(lengths_sorted)),
  `Size (Mb)` = round(total_bp / 1e6, 1),
  N50 = round(n50 / 1e6, 2),
  L50 = as.integer(l50),
  `Largest (Mb)` = round(max(lengths_sorted) / 1e6, 1),
  `Asgn / Unasgn / Chim` = paste(n_assigned, "/", n_unassigned, "/", n_chimeric),
  `Full-length` = as.integer(n_full_length),
  `Both telo` = as.integer(n_both_telo),
  `Mean ref cov` = if (is.na(mean_ref_cov)) "\u2014" else sprintf("%.1f%%", mean_ref_cov * 100),
  `Mean identity` = if (is.na(mean_ident)) "\u2014" else sprintf("%.2f%%", mean_ident * 100),
  Contaminants = as.integer(n_contaminants),
  rDNA = as.integer(n_rdna),
  chrC = if (has_chrC) "\u2713" else "\u2014",
  chrM = if (has_chrM) "\u2713" else "\u2014"
) %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Assembly Summary") %>%
  cols_align(align = "right", columns = where(is.numeric)) %>%
  tab_spanner(label = "Contiguity", columns = c(Contigs, `Size (Mb)`, N50, L50, `Largest (Mb)`)) %>%
  tab_spanner(label = "Chromosomes", columns = c(`Asgn / Unasgn / Chim`, `Full-length`, `Both telo`, `Mean ref cov`, `Mean identity`)) %>%
  tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrC, chrM))
```

`r if (has_compleasm) "## BUSCO Completeness" else ""`

```{r compleasm_table_overview, eval=has_compleasm, results='asis'}
busco_colors <- c(
  "Single-copy"  = "#33a2c4",
  "Duplicated"   = "#225ea8",
  "Fragmented"   = "#fdae61",
  "Interspersed" = "#fee08b",
  "Missing"      = "#d73027"
)

rows <- list()
if (!is.null(compleasm_chrs)) {
  rows[[length(rows) + 1]] <- tibble(
    Dataset = "Chromosomes",
    S = compleasm_chrs$`Single-copy`, D = compleasm_chrs$Duplicated,
    Frag = compleasm_chrs$Fragmented, Inter = compleasm_chrs$Interspersed,
    M = compleasm_chrs$Missing,
    N = compleasm_chrs$n_total
  )
}
if (!is.null(compleasm_non)) {
  rows[[length(rows) + 1]] <- tibble(
    Dataset = "Non-chromosomes",
    S = compleasm_non$`Single-copy`, D = compleasm_non$Duplicated,
    Frag = compleasm_non$Fragmented, Inter = compleasm_non$Interspersed,
    M = compleasm_non$Missing,
    N = compleasm_non$n_total
  )
}
busco_df <- bind_rows(rows)

make_busco_bar <- function(s, d, f, i, m) {
  vals <- c(s, d, f, i, m)
  cols <- unname(busco_colors)
  total <- sum(vals)
  if (total == 0) return("")
  bar_w <- 220; bar_h <- 16; x <- 0; rects <- ""
  for (j in seq_along(vals)) {
    w <- vals[j] / total * bar_w
    if (w > 0) {
      rects <- paste0(rects, sprintf(
        '<rect x="%.1f" y="0" width="%.1f" height="%d" fill="%s"/>',
        x, w, bar_h, cols[j]))
      x <- x + w
    }
  }
  sprintf('<svg width="%d" height="%d" xmlns="http://www.w3.org/2000/svg">%s</svg>',
    bar_w, bar_h, rects)
}

busco_df <- busco_df %>%
  rowwise() %>%
  mutate(bar = make_busco_bar(S, D, Frag, Inter, M)) %>%
  ungroup()

lineage_name <- if (!is.null(compleasm_chrs)) compleasm_chrs$lineage else compleasm_non$lineage

busco_df %>%
  select(Dataset, bar, S, D, Frag, Inter, M, N) %>%
  gt() %>%
  fmt(columns = bar, fns = function(x) lapply(x, html)) %>%
  fmt_number(columns = c(S, D, Frag, Inter, M), decimals = 1, pattern = "{x}%") %>%
  cols_label(
    Dataset = "Dataset",
    bar = "BUSCO Composition",
    S = "Single %", D = "Dup %", Frag = "Frag %", Inter = "Inter %", M = "Miss %",
    N = "Total"
  ) %>%
  cols_width(bar ~ px(240)) %>%
  cols_align(align = "center", columns = c(S, D, Frag, Inter, M, N)) %>%
  tab_header(
    title = md("**BUSCO Completeness**"),
    subtitle = md(sprintf("Lineage: **%s** | %d total BUSCOs",
                          lineage_name, busco_df$N[1]))
  ) %>%
  tab_source_note(md(paste0(
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[1], ';margin-right:3px;vertical-align:middle;"></span>Single-copy &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[2], ';margin-right:3px;vertical-align:middle;"></span>Duplicated &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[3], ';margin-right:3px;vertical-align:middle;"></span>Fragmented &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[4], ';margin-right:3px;vertical-align:middle;"></span>Interspersed &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[5], ';margin-right:3px;vertical-align:middle;"></span>Missing'
  ))) %>%
  gt_theme()
```

# 2. Chromosome Assignments {.tabset}

```{r chr_data}
chr_df <- summary_df %>%
  filter(classification == "chrom_assigned") %>%
  arrange(version_sort_key(assigned_ref_id), assigned_ref_id)

has_telomeres  <- "has_5p_telomere" %in% names(chr_df)
has_full_length <- "is_full_length" %in% names(chr_df)
```

```{r chrom_assign_stub, eval=!has_chrom_assigned, results='asis'}
cat('<div class="note-box">No contigs were assigned to reference chromosomes. This may occur when the query assembly is too divergent from the reference, or when synteny gate thresholds are too strict.</div>')
```

```{r chrom_overview_build, include=FALSE, eval=has_chrom_assigned && has_segments}
# ---------------------------------------------------------------------------
# Chromosome overview
# ---------------------------------------------------------------------------

# Load segment/macro/evidence data
seg <- readr::read_tsv(params$segments_tsv, show_col_types = FALSE,
  col_types = cols(contig = col_character(), contig_len = col_double(),
    target_chrom_id = col_character(), target_subgenome = col_character(),
    strand = col_character(), chain_id = col_double(),
    qstart = col_double(), qend = col_double()))

macro <- readr::read_tsv(params$macro_blocks_tsv, show_col_types = FALSE,
  col_types = cols(contig = col_character(), contig_len = col_double(),
    ref_id = col_character(), chrom_id = col_character(),
    subgenome = col_character(), strand = col_character(),
    chain_id = col_double(), qstart = col_double(), qend = col_double(),
    qspan_bp = col_double(), union_bp = col_double(),
    matches = col_double(), aln_len = col_double(),
    identity = col_double(), score = col_double(),
    n_segments = col_double(), gene_count_chain = col_double())) %>%
  mutate(target_chrom_id = chrom_id, target_subgenome = subgenome)

ev <- readr::read_tsv(params$evidence_tsv, show_col_types = FALSE,
  col_types = cols(contig = col_character(), ref_id = col_character(),
    chrom_id = col_character(), subgenome = col_character(),
    union_bp = col_double(), identity = col_double(), score_all = col_double()))

# rDNA annotations (optional)
rdna_annot <- if (nzchar(params$rdna_annotations_tsv) && file.exists(params$rdna_annotations_tsv)) {
  readr::read_tsv(params$rdna_annotations_tsv, show_col_types = FALSE,
    col_types = cols(contig = col_character(), start = col_double(), end = col_double(),
      strand = col_character(), identity = col_double(),
      consensus_coverage = col_double(), copy_type = col_character(),
      sub_features = col_character(), array_id = col_character(),
      contig_classification = col_character())) %>%
    filter(copy_type == "full")
} else tibble()

# Scaffold AGP (optional)
scaffold_brackets <- if (nzchar(params$agp_tsv) && file.exists(params$agp_tsv)) {
  agp_raw <- readr::read_tsv(params$agp_tsv, comment = "#", col_names = FALSE,
    show_col_types = FALSE, col_types = cols(.default = col_character()))
  agp_raw %>% filter(X5 == "W") %>% select(scaffold = X1, component = X6) %>%
    group_by(scaffold) %>% filter(n() > 1) %>% ungroup()
} else tibble()

# Reference chromosome data
# Show reference chromosomes that either have an assigned contig or are large
# enough to be a chromosome (≥ chr_like_minlen).  Large unassigned chromosomes
# are retained to reveal aneuploidy at a glance; small unassigned scaffolds
# are excluded to avoid bloating the y-axis.
assigned_chrom_ids <- summary_df %>%
  filter(classification == "chrom_assigned") %>%
  pull(assigned_chrom_id) %>% unique()
show_chrom_ids <- ref_lengths %>%
  filter(chrom_id %in% assigned_chrom_ids | ref_len >= chr_like_minlen) %>%
  pull(chrom_id) %>% unique()

if (has_subgenomes) {
  ref_sg <- ref_lengths %>%
    filter(subgenome %in% sg_levels, chrom_id %in% show_chrom_ids) %>%
    mutate(chrom_label = str_replace(chrom_id, "^[Cc]hr", ""),
           chrom_index = suppressWarnings(as.integer(chrom_label)),
           chrom_index = if_else(is.na(chrom_index), 9999L, chrom_index))
  chrom_levels <- ref_sg %>% distinct(chrom_id, chrom_index) %>%
    arrange(chrom_index, chrom_id) %>% pull(chrom_id)
} else {
  ref_sg <- ref_lengths %>%
    filter(!(ref_id %in% c("chrC", "chrM", "ChrC", "ChrM")),
           chrom_id %in% show_chrom_ids) %>%
    mutate(chrom_label = str_replace(chrom_id, "^[Cc]hr", ""),
           chrom_index = suppressWarnings(as.integer(chrom_label)),
           chrom_index = if_else(is.na(chrom_index), 9999L, chrom_index),
           subgenome = "G")
  chrom_levels <- ref_sg %>% distinct(chrom_id, chrom_index) %>%
    arrange(chrom_index, chrom_id) %>% pull(chrom_id)
  seg <- seg %>% mutate(target_subgenome = "G")
  macro <- macro %>% mutate(target_subgenome = "G")
}

# Vertical orientation: chr1 at top
chrom_spacing <- 1.2
y_levels <- rev(chrom_levels)
y_tbl <- tibble(chrom_id = factor(y_levels, levels = y_levels),
                y_index = seq_along(y_levels) * chrom_spacing)

ref_lines_base <- ref_sg %>%
  mutate(chrom_id = factor(chrom_id, levels = y_levels)) %>%
  left_join(y_tbl, by = "chrom_id") %>%
  filter(!is.na(y_index)) %>%
  mutate(ref_len_mb = ref_len / 1e6)

df_plot <- summary_df %>%
  mutate(
    contig_len = if_else(is.na(length), 0, as.numeric(length)),
    assigned_subgenome = case_when(
      !is.na(assigned_subgenome) & assigned_subgenome %in% plot_levels ~ as.character(assigned_subgenome),
      TRUE ~ "NA"),
    assigned_chrom_id = if_else(assigned_chrom_id %in% chrom_levels, assigned_chrom_id, "Un"),
    chrom_id = factor(assigned_chrom_id, levels = y_levels),
    best_identity = as.numeric(best_identity),
    is_reversed = if ("reversed" %in% names(.)) reversed == "yes" else FALSE,
    raw_5p_telo = if ("has_5p_telomere" %in% names(.)) has_5p_telomere == "yes" else FALSE,
    raw_3p_telo = if ("has_3p_telomere" %in% names(.)) has_3p_telomere == "yes" else FALSE,
    has_5p_telo = if_else(is_reversed, raw_3p_telo, raw_5p_telo),
    has_3p_telo = if_else(is_reversed, raw_5p_telo, raw_3p_telo)
  ) %>%
  filter(contig_len >= chr_like_minlen)

ev <- ev %>%
  mutate(subgenome = case_when(
    !is.na(subgenome) & subgenome %in% sg_levels ~ as.character(subgenome), TRUE ~ "NA"
  )) %>% filter(subgenome %in% sg_levels)

# Pill sizing
n_per_chrom <- df_plot %>%
  filter(assigned_chrom_id != "Un", chrom_id %in% y_tbl$chrom_id) %>%
  count(chrom_id, name = "n_contigs")
max_n <- max(n_per_chrom$n_contigs, na.rm = TRUE)
if (!is.finite(max_n) || max_n <= 0) max_n <- 1

n_chroms <- nrow(y_tbl)
fig_height_in <- max(4.0, min(12.0, (n_chroms * 8.0 / 25.4) + 0.8))
plot_area_mm <- (fig_height_in - 0.8) * 25.4
y_range <- n_chroms * chrom_spacing * 1.02
mm_per_unit <- plot_area_mm / y_range

pill_yw_fit <- 0.40 / (1 + log(max_n))
pill_cap <- 1.8 / mm_per_unit; pill_min <- 0.8 / mm_per_unit
pill_yw <- max(pill_min, min(pill_cap, pill_yw_fit))
telo_size <- max(0.8, pill_yw * mm_per_unit * 1.375)
pill_dy <- pill_yw * 1.15

# Slot positions
df_slots <- df_plot %>%
  mutate(contig_len_mb = contig_len / 1e6) %>%
  filter(assigned_chrom_id != "Un") %>%
  left_join(y_tbl, by = "chrom_id") %>%
  filter(!is.na(y_index)) %>%
  group_by(chrom_id) %>%
  mutate(contig = factor(contig, levels = str_sort(unique(contig), numeric = TRUE))) %>%
  arrange(assigned_subgenome, desc(contig_len), contig, .by_group = TRUE) %>%
  mutate(n_in_bin = n(), slot = row_number(),
         y_offset = (slot - (n_in_bin + 1) / 2) * pill_dy,
         y_plot = y_index + y_offset,
         ref_subgenome = factor(assigned_subgenome, levels = names(col_dark))) %>%
  ungroup()

# Scaffold brackets
has_brackets <- nrow(scaffold_brackets) > 0
if (has_brackets) {
  has_conf_cols <- all(c("scaffold_location_confidence",
                         "scaffold_orientation_confidence") %in% names(summary_df))
  bracket_with_conf <- scaffold_brackets %>%
    inner_join(df_slots %>% select(original_name, y_plot), by = c("component" = "original_name"))
  if (has_conf_cols) {
    bracket_with_conf <- bracket_with_conf %>%
      left_join(summary_df %>% select(original_name, scaffold_location_confidence,
                                       scaffold_orientation_confidence),
                by = c("component" = "original_name"))
  }
  bracket_data <- bracket_with_conf %>%
    group_by(scaffold) %>%
    summarise(y_min = min(y_plot), y_max = max(y_plot), y_mid = (min(y_plot) + max(y_plot)) / 2,
              n_components = n(),
              tooltip = {
                hdr <- paste0("Scaffold: ", first(scaffold), " (", n(), " contigs)")
                if (has_conf_cols) {
                  lines <- paste0(component, ": loc=",
                    ifelse(is.na(scaffold_location_confidence), "?", sprintf("%.3f", scaffold_location_confidence)),
                    ", orient=",
                    ifelse(is.na(scaffold_orientation_confidence), "?", sprintf("%.3f", scaffold_orientation_confidence)))
                } else lines <- component
                paste(c(hdr, lines), collapse = "\n")
              }, .groups = "drop") %>%
    filter(n_components > 1)
  has_brackets <- nrow(bracket_data) > 0
}

if (has_brackets) {
  max_len <- max(df_slots$contig_len_mb, na.rm = TRUE)
  bracket_x <- -max_len * 0.035; foot_xend <- -max_len * 0.015
  bracket_segments <- bind_rows(
    bracket_data %>% transmute(x = bracket_x, xend = bracket_x,
      y = y_min - pill_yw/6, yend = y_max + pill_yw/6, scaffold, tooltip),
    bracket_data %>% transmute(x = bracket_x, xend = foot_xend,
      y = y_max + pill_yw/6, yend = y_max + pill_yw/6, scaffold, tooltip),
    bracket_data %>% transmute(x = bracket_x, xend = foot_xend,
      y = y_min - pill_yw/6, yend = y_min - pill_yw/6, scaffold, tooltip))
  bracket_labels <- bracket_data %>%
    transmute(x = bracket_x - max_len * 0.008, y = y_mid, label = "S", scaffold, tooltip)
}

x_expand_left <- if (has_brackets) 0.025 else 0.02

# Rearrangement labels
extract_short_chrom_ids <- function(candidates) {
  if (is.na(candidates) || candidates == "") return(NA_character_)
  paste(str_replace(str_split(candidates, ",")[[1]], "^[Cc]hr", ""), collapse = ",")
}

has_rearr_col <- "rearrangement_candidates" %in% names(df_plot)
if (has_rearr_col) {
  df_rearr <- df_plot %>%
    filter(!is.na(rearrangement_candidates) & rearrangement_candidates != "") %>%
    mutate(
      short_ids = sapply(rearrangement_candidates, extract_short_chrom_ids, USE.NAMES = FALSE),
      rearr_label = paste0("<-> ", short_ids),
      first_candidate = str_split_fixed(rearrangement_candidates, ",", 2)[, 1],
      offtarget_subgenome = str_extract(first_candidate, "[A-Z]$"),
      offtarget_chrom_id = str_replace(first_candidate, "[A-Z]$", ""),
      is_homeologous = (offtarget_chrom_id == assigned_chrom_id),
      label_color = if_else(
        is_homeologous & !is.na(offtarget_subgenome) & offtarget_subgenome %in% names(sg_cols_dark),
        sg_cols_dark[offtarget_subgenome], "red"))

  rearr_labels <- df_rearr %>%
    inner_join(df_slots %>% select(original_name, y_plot, contig_len_mb), by = "original_name") %>%
    mutate(label_x = contig_len_mb)
  max_contig_length <- max(df_slots$contig_len_mb, na.rm = TRUE)
  rearr_label_xmin <- max_contig_length + 0.5
  x_expand_right <- 0.12
} else {
  rearr_labels <- tibble(); rearr_label_xmin <- Inf; x_expand_right <- 0.02
}

# Reference lines
ref_lines_data <- df_slots %>%
  group_by(chrom_id) %>%
  summarise(y_top = max(y_plot), subgenomes_present = list(unique(assigned_subgenome)),
            .groups = "drop") %>%
  tidyr::unnest(subgenomes_present) %>%
  rename(assigned_subgenome = subgenomes_present) %>%
  left_join(ref_lines_base %>% select(chrom_id, subgenome, ref_len_mb),
            by = c("chrom_id" = "chrom_id", "assigned_subgenome" = "subgenome")) %>%
  filter(!is.na(ref_len_mb)) %>%
  group_by(chrom_id) %>% arrange(assigned_subgenome, .by_group = TRUE) %>%
  mutate(n_sg = n(), sg_idx = row_number()) %>% ungroup()

ref_tick_offset <- pill_yw * 0.5
ref_lines <- ref_lines_data %>%
  mutate(y_ref = y_top + pill_yw/2 + ref_tick_offset + (sg_idx - 1) * 0.04 * 2)

# Segment/macro plot data (with coordinate flipping for reversed contigs)
prep_blocks <- function(blocks, slot_data) {
  blocks %>%
    filter(contig %in% slot_data$original_name) %>%
    mutate(qstart_mb = qstart / 1e6, qend_mb = qend / 1e6, contig_len_mb = contig_len / 1e6) %>%
    left_join(slot_data %>% select(original_name, assigned_chrom_id = chrom_id, y_plot,
                                    contig_len_mb_slot = contig_len_mb, is_reversed),
              by = c("contig" = "original_name")) %>%
    filter(!is.na(y_plot)) %>%
    mutate(
      assigned_chrom_id = as.character(assigned_chrom_id),
      off_target = (target_chrom_id != assigned_chrom_id),
      max_len_mb = if_else(is.na(contig_len_mb_slot), contig_len_mb, contig_len_mb_slot),
      qstart_flip = if_else(is_reversed, max_len_mb - qend_mb, qstart_mb),
      qend_flip = if_else(is_reversed, max_len_mb - qstart_mb, qend_mb),
      qstart_mb = pmax(0, pmin(qstart_flip, max_len_mb)),
      qend_mb = pmax(0, pmin(qend_flip, max_len_mb))
    ) %>% filter(qend_mb > qstart_mb) %>%
    mutate(ymin = y_plot - pill_yw/2, ymax = y_plot + pill_yw/2, xmin = qstart_mb, xmax = qend_mb)
}

seg_plot <- prep_blocks(seg, df_slots) %>%
  mutate(tooltip = paste0("contig: ", contig, "\nassigned: ", assigned_chrom_id,
    "\nsegment: ", target_chrom_id, " ", target_subgenome,
    "\nqstart: ", sprintf("%.3f", qstart_mb), " Mbp",
    "\nqend: ", sprintf("%.3f", qend_mb), " Mbp", "\nchain: ", chain_id))

macro_plot <- prep_blocks(macro, df_slots) %>%
  mutate(tooltip = paste0("contig: ", contig, "\nassigned: ", assigned_chrom_id,
    "\nmacro: ", target_chrom_id, " ", target_subgenome,
    "\nqstart: ", sprintf("%.3f", qstart_mb), " Mbp",
    "\nqend: ", sprintf("%.3f", qend_mb), " Mbp",
    "\nidentity: ", sprintf("%.4f", identity), "\nscore: ", sprintf("%.3f", score),
    "\nchain: ", chain_id))

# rDNA annotations
rdna_plot <- if (nrow(rdna_annot) > 0) {
  rdna_annot %>%
    filter(contig %in% df_slots$original_name) %>%
    mutate(start_mb = start / 1e6, end_mb = end / 1e6) %>%
    left_join(df_slots %>% select(original_name, y_plot, contig_len_mb, is_reversed),
              by = c("contig" = "original_name")) %>%
    filter(!is.na(y_plot)) %>%
    mutate(
      start_flip = if_else(is_reversed, contig_len_mb - end_mb, start_mb),
      end_flip = if_else(is_reversed, contig_len_mb - start_mb, end_mb),
      start_mb = pmax(0, pmin(start_flip, contig_len_mb)),
      end_mb = pmax(0, pmin(end_flip, contig_len_mb))
    ) %>% filter(end_mb > start_mb) %>%
    mutate(ymin = y_plot - pill_yw/2, ymax = y_plot + pill_yw/2,
           xmin = start_mb, xmax = end_mb,
           tooltip = paste0("rDNA full copy\ncontig: ", contig,
             "\nstart: ", sprintf("%.3f", start_mb), " Mbp",
             "\nend: ", sprintf("%.3f", end_mb), " Mbp",
             "\nidentity: ", sprintf("%.4f", identity),
             "\ncoverage: ", sprintf("%.4f", consensus_coverage)))
} else tibble()

# Build composition panel (interactive version for both PDF + girafe)
p_comp <- ggplot() +
  geom_segment(data = ref_lines,
    aes(x = 0, xend = ref_len_mb, y = y_ref, yend = y_ref, color = assigned_subgenome),
    linewidth = 0.4, alpha = 0.9, show.legend = FALSE) +
  scale_color_manual(values = col_light, guide = "none", drop = FALSE) +
  ggnewscale::new_scale_color() +
  # Telomere indicators (short horizontal segments with round lineend)
  ggiraph::geom_segment_interactive(
    data = df_slots[df_slots$has_5p_telo, ],
    aes(x = -0.02, xend = 0.02,
        y = y_plot, yend = y_plot,
        tooltip = paste0(original_name, "\n5' telomere detected")),
    lineend = "round", linewidth = telo_size,
    color = "grey75", show.legend = FALSE) +
  ggiraph::geom_segment_interactive(
    data = df_slots[df_slots$has_3p_telo, ],
    aes(x = contig_len_mb - 0.02, xend = contig_len_mb + 0.02,
        y = y_plot, yend = y_plot,
        tooltip = paste0(original_name, "\n3' telomere detected")),
    lineend = "round", linewidth = telo_size,
    color = "grey75", show.legend = FALSE) +
  # Pill backgrounds
  geom_rect(data = df_slots,
    aes(xmin = 0, xmax = contig_len_mb, ymin = y_plot - pill_yw/2, ymax = y_plot + pill_yw/2),
    fill = "grey88", color = NA, show.legend = FALSE) +
  ggnewscale::new_scale_fill() +
  # Segments (on-target)
  ggiraph::geom_rect_interactive(
    data = seg_plot %>% filter(!off_target),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
        fill = target_subgenome, tooltip = tooltip),
    color = NA, alpha = 1.0) +
  # Segments (off-target = red)
  ggiraph::geom_rect_interactive(
    data = seg_plot %>% filter(off_target),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, tooltip = tooltip),
    fill = "red", color = NA, alpha = 1.0) +
  scale_fill_manual(values = col_light, guide = "none", drop = FALSE) +
  ggnewscale::new_scale_fill() +
  # Macro blocks (on-target)
  ggiraph::geom_rect_interactive(
    data = macro_plot %>% filter(!off_target),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
        fill = target_subgenome, tooltip = tooltip),
    color = NA, alpha = 1.0, show.legend = FALSE) +
  # Macro blocks (off-target)
  ggiraph::geom_rect_interactive(
    data = macro_plot %>% filter(off_target),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, tooltip = tooltip),
    fill = "red", color = NA, alpha = 1.0, show.legend = FALSE) +
  # rDNA annotations
  { if (nrow(rdna_plot) > 0)
      ggiraph::geom_rect_interactive(data = rdna_plot,
        aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, tooltip = tooltip),
        fill = "#7B2D8E", color = NA, alpha = 1.0, show.legend = FALSE) } +
  # Scaffold brackets
  { if (has_brackets) list(
      ggiraph::geom_segment_interactive(data = bracket_segments,
        aes(x = x, xend = xend, y = y, yend = yend, tooltip = tooltip, data_id = scaffold),
        color = "red", linewidth = 0.2),
      ggiraph::geom_text_interactive(data = bracket_labels,
        aes(x = x, y = y, label = label, tooltip = tooltip, data_id = scaffold),
        color = "red", size = (base_font_pt + 2) / .pt * 0.5, family = base_family, hjust = 1)
    ) } +
  # Rearrangement labels
  { if (nrow(rearr_labels) > 0)
      lapply(unique(rearr_labels$label_color), function(col) {
        ggrepel::geom_text_repel(
          data = rearr_labels %>% filter(label_color == col),
          aes(x = label_x, y = y_plot, label = rearr_label),
          color = col, size = (base_font_pt + 2) / .pt * 0.5, family = base_family,
          direction = "x", xlim = c(rearr_label_xmin, NA),
          box.padding = 0.3, point.padding = 0.5,
          segment.color = col, segment.size = 0.2,
          max.overlaps = 20, min.segment.length = 0, show.legend = FALSE)
      }) } +
  scale_fill_manual(values = col_dark, guide = "none", drop = FALSE) +
  guides(fill = guide_legend(title.position = "top", title.hjust = 0.5, nrow = 1, byrow = TRUE,
                             override.aes = list(alpha = 1))) +
  scale_x_continuous(expand = expansion(mult = c(x_expand_left, x_expand_right)), position = "top") +
  scale_y_continuous(breaks = y_tbl$y_index,
                     labels = str_replace(as.character(y_tbl$chrom_id), "^[Cc]hr", ""),
                     expand = expansion(mult = c(0.01, 0.01))) +
  theme_classic(base_family = base_family, base_size = base_font_pt) +
  axis_theme +
  theme(axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5, size = base_font_pt, family = base_family),
        panel.grid = element_blank(),
        legend.position = c(0.985, 0.985), legend.justification = c(1, 1)) +
  labs(x = "Mbp", y = "Assigned chromosome",
       title = paste0("Contig composition (", params$plot_suffix, ")"))
p_comp <- apply_legend_theme(p_comp, text_pt = 6, key_pt = 6, tight = TRUE) +
  theme(legend.margin = margin(1, 1, 1, 1))

# Radar panel (polyploid only)
if (has_subgenomes) {
  radar_support <- ev %>%
    mutate(support = pmax(0, union_bp) * pmax(0, pmin(1, identity))) %>%
    group_by(contig, subgenome) %>%
    summarise(support = sum(support, na.rm = TRUE), .groups = "drop") %>%
    tidyr::complete(contig, subgenome = plot_levels, fill = list(support = 0)) %>%
    group_by(contig) %>%
    mutate(wsum = sum(support), w = if_else(wsum > 0, support / wsum, NA_real_)) %>%
    ungroup()

  get_vertices <- function(pl) {
    n <- length(pl)
    if (n == 2) tibble(subgenome = pl, vx = c(-0.5, 0.5), vy = c(0, 0))
    else if (n == 3) tibble(subgenome = pl, vx = c(-0.5, 0.5, 0.0), vy = c(0.0, 0.0, sqrt(3)/2))
    else if (n == 4) tibble(subgenome = pl, vx = c(-0.5, 0.5, 0.5, -0.5), vy = c(-0.5, -0.5, 0.5, 0.5))
    else stop("get_vertices expects 2..4 subgenomes")
  }
  verts <- get_vertices(plot_levels)

  label_verts <- if (n_sg == 2) {
    verts %>% mutate(lx = if_else(vx < 0, -0.58, 0.58), ly = 0,
                     hjust = if_else(vx < 0, 1, 0), vjust = 0.5)
  } else if (n_sg == 3) {
    verts %>% mutate(centroid_y = sqrt(3)/6,
      dir_x = vx, dir_y = vy - centroid_y,
      dir_len = sqrt(dir_x^2 + dir_y^2),
      lx = vx + 0.12 * dir_x / dir_len, ly = vy + 0.12 * dir_y / dir_len,
      hjust = 0.5, vjust = 0.5) %>%
      select(-centroid_y, -dir_x, -dir_y, -dir_len)
  } else {
    verts %>% mutate(lx = vx * 1.2, ly = vy * 1.2, hjust = 0.5, vjust = 0.5)
  }

  radar_xy <- radar_support %>%
    left_join(verts, by = "subgenome") %>%
    group_by(contig) %>%
    summarise(rx = sum(w * vx, na.rm = TRUE), ry = sum(w * vy, na.rm = TRUE), .groups = "drop") %>%
    left_join(df_plot %>% select(original_name, assigned_subgenome, classification),
              by = c("contig" = "original_name")) %>%
    mutate(assigned_subgenome = if_else(is.na(assigned_subgenome), "NA", as.character(assigned_subgenome)),
           classification = if_else(is.na(classification), "NA", as.character(classification))) %>%
    filter(assigned_subgenome %in% plot_levels) %>%
    mutate(assigned_subgenome = factor(assigned_subgenome, levels = c(plot_levels, "NA")))

  frame <- if (n_sg == 2) tibble(x = c(-0.5, 0.5), y = c(0, 0))
           else verts %>% transmute(x = vx, y = vy) %>% bind_rows(.[1, ])

  radar_xlim <- c(-0.6, 0.6)
  radar_ylim <- if (n_sg == 2) c(-0.6, 0.6) else c(-0.15, max(verts$vy) + 0.06)
  axis_pos <- (0 - radar_ylim[1]) / (radar_ylim[2] - radar_ylim[1])
  legend_pos <- c(0.5, pmax(0.02, axis_pos - 0.08))

  p_radar <- ggplot() +
    geom_path(data = frame, aes(x = x, y = y), linewidth = 0.4, color = "grey40") +
    geom_text(data = label_verts,
      aes(x = lx, y = ly, label = subgenome, hjust = hjust, vjust = vjust),
      color = "grey20", size = base_font_pt / .pt, family = base_family) +
    geom_point(data = radar_xy %>% filter(!is.na(rx), !is.na(ry)),
      aes(x = rx, y = ry, color = assigned_subgenome),
      shape = 16, alpha = 0.80, size = 1.6, show.legend = FALSE) +
    scale_color_manual(values = col_dark, drop = FALSE) +
    coord_fixed(ratio = 1, xlim = radar_xlim, ylim = radar_ylim, clip = "off") +
    theme_classic(base_family = base_family, base_size = base_font_pt) +
    theme(axis.title.x = element_text(hjust = 0.5, size = base_font_pt, family = base_family,
                                       margin = margin(t = -2)),
          axis.title.y = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.line = element_blank(),
          panel.grid = element_blank(), plot.margin = margin(2, 2, 2, 2)) +
    labs(x = "Chromosome-set support", y = NULL)
  p_radar <- apply_legend_theme(p_radar, text_pt = 6, key_pt = 6, tight = TRUE) +
    theme(legend.position = legend_pos, legend.justification = c(0.5, 1))
} else {
  p_radar <- ggplot() + theme_void(base_family = base_family, base_size = base_font_pt) +
    labs(title = "Subgenome support (n/a)")
}

# Identity panel
if (has_subgenomes) {
  df_id_assigned <- df_plot %>%
    filter(!is.na(best_identity), assigned_subgenome %in% plot_levels) %>%
    mutate(assigned_sg = factor(assigned_subgenome, levels = plot_levels))
  id_stats <- df_id_assigned %>%
    group_by(assigned_sg) %>%
    summarise(mean_id = mean(best_identity, na.rm = TRUE),
              sd_id = sd(best_identity, na.rm = TRUE), .groups = "drop") %>%
    mutate(ymin = pmax(0, mean_id - sd_id), ymax = pmin(1, mean_id + sd_id))
  p_id <- ggplot(df_id_assigned, aes(x = assigned_sg, y = pmax(0, pmin(1, best_identity)),
                                      color = assigned_sg)) +
    geom_jitter(width = 0.18, height = 0, alpha = 0.55, size = 1.2, show.legend = FALSE) +
    geom_errorbar(data = id_stats, aes(x = assigned_sg, y = mean_id, ymin = ymin, ymax = ymax,
                                        color = assigned_sg),
                  width = 0.1, linewidth = 0.4, show.legend = FALSE, position = position_nudge(x = 0.35)) +
    geom_point(data = id_stats, aes(x = assigned_sg, y = mean_id, color = assigned_sg),
               shape = 16, size = 1.5, show.legend = FALSE, position = position_nudge(x = 0.35)) +
    scale_color_manual(values = col_dark, drop = FALSE) +
    scale_y_continuous(limits = c(0.4, 1), breaks = seq(0.4, 1, 0.1)) +
    theme_classic(base_family = base_family, base_size = base_font_pt) +
    axis_theme + theme(axis.ticks = element_blank(), panel.grid = element_blank(),
                       plot.margin = margin(5.5, 5.5, 5.5, 5.5)) +
    labs(x = "Assigned subgenome", y = "Alignment identity")
} else {
  df_id_assigned <- df_plot %>%
    filter(!is.na(best_identity), assigned_ref_id != "NA") %>%
    mutate(assigned_bin = factor("Genome", levels = "Genome"))
  p_id <- ggplot(df_id_assigned, aes(x = assigned_bin, y = pmax(0, pmin(1, best_identity)))) +
    geom_jitter(width = 0.18, height = 0, alpha = 0.55, size = 1.2, color = "grey35", show.legend = FALSE) +
    stat_summary(fun = median, geom = "point", shape = 95, size = 7, color = "grey15", show.legend = FALSE) +
    scale_y_continuous(limits = c(0.4, 1), breaks = seq(0.4, 1, 0.1)) +
    theme_classic(base_family = base_family, base_size = base_font_pt) +
    axis_theme + theme(axis.ticks = element_blank(), panel.grid = element_blank(),
                       plot.margin = margin(5.5, 5.5, 5.5, 5.5)) +
    labs(x = NULL, y = "Alignment identity")
}

# Patchwork layout
p_placeholder <- ggplot() +
  theme_void(base_family = base_family, base_size = base_font_pt) +
  theme(plot.background = element_rect(fill = "white", color = NA),
        plot.margin = margin(5.5, 5.5, 5.5, 5.5))

usable_in <- fig_height_in - 0.8
id_h <- 1.3 / usable_in; gap_h <- 0.15 / usable_in
radar_h <- 1.7 / usable_in; spacer_h <- max(0.01, 1 - id_h - gap_h - radar_h)

right_col <- if (has_subgenomes) {
  p_id / plot_spacer() / p_radar / p_placeholder +
    plot_layout(heights = c(id_h, gap_h, radar_h, spacer_h))
} else {
  p_id / p_placeholder + plot_layout(heights = c(id_h, 1 - id_h))
}

full_plot_chrom <- (p_comp | right_col) + plot_layout(widths = c(7, 3))
save_pdf(pdf_chrom_overview, full_plot_chrom, width = 7.2, height = fig_height_in)
```

```{r chrom_overview_widget, eval=has_chrom_assigned && has_segments}
g_chrom <- ggiraph::girafe(ggobj = full_plot_chrom,
                           width_svg = 7.2,
                           height_svg = fig_height_in)
ggiraph::girafe_options(g_chrom,
  ggiraph::opts_sizing(rescale = TRUE, width = 1),
  ggiraph::opts_tooltip(css = paste0(
    "font-family: ", base_family, "; font-size: 9pt; ",
    "background: white; padding: 6px; border: 1px solid #666;"
  )),
  ggiraph::opts_hover(css = "opacity: 0.8;"),
  ggiraph::opts_toolbar(saveaspng = FALSE)
)
```

```{r subgenome_tab_heading, eval=has_subgenomes && has_evidence && !is.null(evidence_df), results='asis'}
cat('\n## Subgenome Composition\n')
```

```{r subgenome_table, eval=has_subgenomes && has_evidence && !is.null(evidence_df)}
subg_df <- chr_df %>%
  filter(!is.na(assigned_subgenome) & assigned_subgenome != "NA") %>%
  group_by(query_subgenome = assigned_subgenome) %>%
  summarise(n_contigs = n(), total_mb = sum(length / 1e6, na.rm = TRUE),
    mean_identity = if ("seq_identity_vs_ref" %in% names(.)) mean(seq_identity_vs_ref, na.rm = TRUE) * 100
                    else if ("best_identity" %in% names(.)) mean(best_identity, na.rm = TRUE) * 100
                    else NA_real_,
    mean_collinearity = if (has_collinearity) mean(collinearity_score, na.rm = TRUE) else NA_real_,
    .groups = "drop")

if (nrow(subg_df) > 1) {
  sel <- c("query_subgenome", "n_contigs", "total_mb", "mean_identity")
  labels <- list(query_subgenome = "Subgenome", n_contigs = "Contigs",
                 total_mb = "Total Mb", mean_identity = "Mean Identity %")
  if (has_collinearity) { sel <- c(sel, "mean_collinearity"); labels$mean_collinearity <- "Mean Collinearity" }
  subg_df %>% select(all_of(sel)) %>% gt() %>% cols_label(.list = labels) %>%
    fmt_number(columns = total_mb, decimals = 1) %>%
    fmt_number(columns = mean_identity, decimals = 1) %>%
    fmt_number(columns = any_of("mean_collinearity"), decimals = 2) %>%
    tab_header(title = md("**Subgenome Composition**")) %>% gt_theme()
}
```

```{r chr_tab_heading, eval=has_chrom_assigned, results='asis'}
cat('\n## Per-Chromosome Summary\n')
```

```{r chr_assignment_table, eval=nrow(chr_df)>0}
chr_tbl <- chr_df %>%
  mutate(
    length_mb = length / 1e6,
    ref_cov_pct = if ("ref_coverage" %in% names(.)) ref_coverage * 100 else NA_real_,
    identity_pct = if ("seq_identity_vs_ref" %in% names(.)) seq_identity_vs_ref * 100
                   else if ("best_identity" %in% names(.)) best_identity * 100 else NA_real_,
    gene_count = if ("ref_gene_proportion" %in% names(.)) round(ref_gene_proportion * 100, 1) else NA_real_,
    conf_html = sapply(classification_confidence, conf_badge),
    collinearity = if (has_collinearity) collinearity_score else NA_real_,
    tel_5p = if (has_telomeres) has_5p_telomere else NA_character_,
    tel_3p = if (has_telomeres) has_3p_telomere else NA_character_,
    full_len = if (has_full_length) is_full_length else NA_character_,
    subgenome_col = if (has_subgenomes) assigned_subgenome else NA_character_
  )

sel_cols <- c("contig", "original_name", "assigned_ref_id", "length_mb")
col_labels <- list(contig = "Contig", original_name = "Original Name", assigned_ref_id = "Ref Chromosome", length_mb = "Length (Mb)")

if (has_subgenomes) { sel_cols <- c(sel_cols, "subgenome_col"); col_labels$subgenome_col <- "Subgenome" }
sel_cols <- c(sel_cols, "ref_cov_pct"); col_labels$ref_cov_pct <- "Ref Cov %"
if (has_full_length) { sel_cols <- c(sel_cols, "full_len"); col_labels$full_len <- "Full-length" }
if (has_telomeres) {
  sel_cols <- c(sel_cols, "tel_5p", "tel_3p")
  col_labels$tel_5p <- "Tel 5'"; col_labels$tel_3p <- "Tel 3'"
}
sel_cols <- c(sel_cols, "gene_count"); col_labels$gene_count <- "Gene %"
sel_cols <- c(sel_cols, "identity_pct"); col_labels$identity_pct <- "Identity %"
if (has_collinearity) { sel_cols <- c(sel_cols, "collinearity"); col_labels$collinearity <- "Collinearity" }
sel_cols <- c(sel_cols, "conf_html"); col_labels$conf_html <- "Confidence"

chr_tbl %>%
  select(all_of(sel_cols)) %>% gt() %>%
  cols_label(.list = col_labels) %>%
  fmt_number(columns = any_of("length_mb"), decimals = 1) %>%
  fmt_number(columns = any_of(c("ref_cov_pct", "identity_pct", "gene_count")), decimals = 1) %>%
  fmt_number(columns = any_of("collinearity"), decimals = 2) %>%
  fmt(columns = any_of("conf_html"), fns = function(x) x) %>%
  tab_header(title = md("**Per-Chromosome Summary**"),
             subtitle = md(paste0(nrow(chr_df), " contigs assigned to ",
                                  n_distinct(chr_df$assigned_ref_id), " reference chromosomes"))) %>%
  gt_theme()
```


```{r ref_cov_tab_heading, eval=has_chrom_assigned, results='asis'}
cat('\n## Reference Coverage\n')
```

```{r ref_coverage_table, eval=nrow(chr_df)>0}
ref_cov <- chr_df %>%
  group_by(assigned_ref_id) %>%
  summarise(n_contigs = n(), total_query_mb = sum(length / 1e6, na.rm = TRUE), .groups = "drop") %>%
  left_join(
    ref_lengths %>%
      filter(!str_detect(ref_id, "^chr[CM]$|_random$")) %>%
      mutate(ref_length_mb = ref_len / 1e6) %>%
      select(ref_id, ref_length_mb),
    by = c("assigned_ref_id" = "ref_id")
  ) %>%
  mutate(coverage_pct = if_else(!is.na(ref_length_mb) & ref_length_mb > 0,
                                 total_query_mb / ref_length_mb * 100, NA_real_)) %>%
  arrange(version_sort_key(assigned_ref_id), assigned_ref_id)

max_cov <- max(ref_cov$coverage_pct, na.rm = TRUE)
ref_cov %>%
  mutate(bar = mapply(function(v) gt_inline_bar(v, max(max_cov, 100), color = oi[6]), coverage_pct)) %>%
  select(assigned_ref_id, ref_length_mb, total_query_mb, coverage_pct, bar, n_contigs) %>%
  gt() %>%
  cols_label(assigned_ref_id = "Ref Chromosome", ref_length_mb = "Ref (Mb)",
             total_query_mb = "Query (Mb)", coverage_pct = "Coverage %", bar = "", n_contigs = "Contigs") %>%
  fmt_number(columns = c(ref_length_mb, total_query_mb), decimals = 1) %>%
  fmt_number(columns = coverage_pct, decimals = 1) %>%
  fmt(columns = bar, fns = function(x) x) %>%
  tab_header(title = md("**Reference Coverage Summary**")) %>%
  gt_theme()
```


# 3. Evidence Quality

```{r evidence_placeholder, eval=!has_evidence || is.null(evidence_df), results='asis'}
cat('<div class="note-box">No synteny evidence data available.</div>')
```

```{r evidence_table, eval=has_evidence && !is.null(evidence_df)}
ev_df <- summary_df %>%
  filter(classification %in% c("chrom_assigned", "chrom_unassigned")) %>%
  mutate(
    length_mb = length / 1e6,
    conf_html = sapply(classification_confidence, conf_badge),
    span_frac = if ("best_ref_union_frac" %in% names(.)) best_ref_union_frac else NA_real_,
    gene_pct = if ("ref_gene_proportion" %in% names(.)) ref_gene_proportion * 100 else NA_real_,
    syn_score = if ("synteny_score" %in% names(.)) synteny_score else NA_real_,
    collin = if (has_collinearity) collinearity_score else NA_real_,
    gc_dev = if ("gc_deviation" %in% names(.)) gc_deviation else NA_real_,
    # Cap sentinel value (1e9) to display as "max"
    s_ratio = if ("score_ratio" %in% names(.)) {
      ifelse(score_ratio >= 1e8, Inf, score_ratio)
    } else NA_real_
  ) %>%
  arrange(version_sort_key(assigned_ref_id), assigned_ref_id, desc(length_mb))

sel <- c("contig", "original_name", "length_mb", "assigned_ref_id", "syn_score", "span_frac", "gene_pct")
labels <- list(contig = "Contig", original_name = "Original Name", length_mb = "Length (Mb)", assigned_ref_id = "Best Ref",
               syn_score = "Synteny Score", span_frac = "Span Frac", gene_pct = "Gene %")
if (has_collinearity) { sel <- c(sel, "collin"); labels$collin <- "Collinearity" }
sel <- c(sel, "s_ratio", "gc_dev", "conf_html")
labels$s_ratio <- "Score Ratio"; labels$gc_dev <- "GC Dev"; labels$conf_html <- "Confidence"

ev_df %>%
  select(all_of(sel)) %>% gt() %>%
  cols_label(.list = labels) %>%
  fmt_number(columns = any_of(c("length_mb", "span_frac", "gene_pct", "gc_dev")), decimals = 1) %>%
  fmt_number(columns = any_of(c("syn_score", "collin")), decimals = 2) %>%
  fmt(columns = any_of("s_ratio"), fns = function(x) {
    ifelse(is.infinite(x), "\u221e", sprintf("%.1f", x))
  }) %>%
  fmt(columns = any_of("conf_html"), fns = function(x) x) %>%
  tab_header(title = md("**Per-Contig Evidence Summary**"),
             subtitle = md("Sorted by chromosome assignment")) %>%
  gt_theme()
```

<div class="legend-box">
<b>Synteny Score</b> = total aligned bp from best chain to assigned reference.
<b>Span Frac</b> = fraction of contig covered by synteny blocks.
<b>Gene %</b> = proportion of reference genes found on this contig (protein mode only).
<b>Collinearity</b> = fraction of synteny blocks in reference order (1.0 = perfectly collinear).
<b>Score Ratio</b> = best score / second-best score (&infin; = no second hit; higher = more confident assignment).
<b>GC Dev</b> = GC content deviation from reference mean (in standard deviations).
</div>


# 4. Rearrangements / Misassemblies

```{r rearr_placeholder, eval=!has_rearrangements, results='asis'}
cat('<div class="note-box">No structural rearrangements or potential misassemblies were detected in this assembly.</div>')
```

```{r rearr_filter, eval=has_rearrangements, results='asis'}
n_high <- sum(rearr_df$confidence == "high")
n_med  <- sum(rearr_df$confidence == "medium")
n_low  <- sum(rearr_df$confidence == "low")
cat(sprintf('
<div class="legend-box" style="border-left-color: #2a6496;">
<b>Min. confidence</b>:&ensp;
<label><input type="radio" name="rearr_conf" value="high" checked
  onchange="filterRearr(this.value)"> High (%d)</label>&emsp;
<label><input type="radio" name="rearr_conf" value="medium"
  onchange="filterRearr(this.value)"> Medium (%d)</label>&emsp;
<label><input type="radio" name="rearr_conf" value="low"
  onchange="filterRearr(this.value)"> Low (%d)</label>
</div>
<script>
function filterRearr(minConf) {
  var levels = {high: 3, medium: 2, low: 1};
  var minLevel = levels[minConf] || 1;
  var tbl = document.getElementById("rearr-gt");
  if (!tbl) return;
  var rows = tbl.querySelectorAll("tbody tr");
  rows.forEach(function(tr) {
    var conf = tr.getAttribute("data-conf") || "low";
    tr.style.display = (levels[conf] >= minLevel) ? "" : "none";
  });
}
</script>
', n_high, n_med, n_low))
```

```{r rearr_table, eval=has_rearrangements, results='asis'}
# rearr_df already has 'contig' (renamed) and 'original_name' from the TSV
rearr_show <- rearr_df %>%
  mutate(
    type_label = str_replace_all(rearrangement_type, "_", " ") %>% str_to_sentence(),
    span_mb = span_bp / 1e6,
    conf_html = sapply(confidence, conf_badge)
  )

# Build gt table with id for JS filtering
tbl <- rearr_show %>%
  select(type_label, contig, assigned_ref_id, partner_ref_id,
         span_mb, evidence, conf_html) %>%
  gt() %>%
  cols_label(
    type_label = "Type*",
    contig = "Contig",
    assigned_ref_id = "Assigned",
    partner_ref_id = "Partner",
    span_mb = "Span (Mb)",
    evidence = "Evidence",
    conf_html = "Confidence"
  ) %>%
  fmt_number(columns = span_mb, decimals = 2) %>%
  fmt(columns = conf_html, fns = function(x) x) %>%
  sub_missing(missing_text = "\u2014") %>%
  tab_header(title = md("**Rearrangements / Misassemblies**")) %>%
  gt_theme()

# Inject table id for JS filtering
tbl_html <- as.character(as_raw_html(tbl))
tbl_html <- sub("<table", '<table id="rearr-gt"', tbl_html, fixed = TRUE)
cat(tbl_html)
# Tag rows and apply initial filter after table is in the DOM
cat('<script>
(function() {
  var tbl = document.getElementById("rearr-gt");
  if (!tbl) return;
  var rows = tbl.querySelectorAll("tbody tr");
  rows.forEach(function(tr) {
    var cells = tr.querySelectorAll("td");
    var last = cells[cells.length - 1];
    if (last) {
      var txt = last.textContent.trim().toLowerCase();
      if (txt === "high" || txt === "medium" || txt === "low") {
        tr.setAttribute("data-conf", txt);
      }
    }
  });
  filterRearr("high");
})();
</script>\n')

# Legend: misassembly mimics for each rearrangement type
mimic_map <- rearr_show %>%
  distinct(rearrangement_type, misassembly_mimic) %>%
  filter(nzchar(misassembly_mimic)) %>%
  mutate(type_label = str_replace_all(rearrangement_type, "_", " ") %>%
           str_to_sentence())
```

```{r rearr_legend, eval=has_rearrangements, results='asis'}
cat('<div class="legend-box">\n')
cat('<b>*Misassembly mimics</b>: Each rearrangement type has an alternative ')
cat('explanation as an assembly artifact.<br>\n')
for (i in seq_len(nrow(mimic_map))) {
  cat(sprintf('<b>%s</b> \u2014 %s', mimic_map$type_label[i], mimic_map$misassembly_mimic[i]))
  if (i < nrow(mimic_map)) cat('<br>\n')
}
cat('\n</div>\n')
```


# 5. Read Depth Analysis {.tabset}

```{r depth_placeholder, eval=!has_depth, results='asis'}
cat('<div class="note-box">Read depth data is not available. Provide sequencing reads with <code>--reads</code> to enable depth analysis.</div>')
```

```{r depth_build, include=FALSE, eval=has_depth}
# ---------------------------------------------------------------------------
# Depth overview (migrated from depth_overview.tmpl.R)
# ---------------------------------------------------------------------------

# Extended classification colors for depth (includes chrC/chrM splits)
depth_classification_order <- c(
  "chrom_assigned", "chrom_debris", "chrom_unassigned",
  "chrC", "chrC_debris", "chrM", "chrM_debris",
  "organelle_complete", "organelle_debris",
  "rDNA", "contaminant", "debris", "unclassified"
)
depth_classification_colors <- c(
  "chrom_assigned"     = oi[5],
  "chrom_debris"       = lighten(oi[5], amount = 0.5),
  "chrom_unassigned"   = oi[2],
  "chrC"               = oi[3],
  "chrC_debris"        = lighten(oi[3], amount = 0.5),
  "chrM"               = oi[6],
  "chrM_debris"        = lighten(oi[6], amount = 0.5),
  "organelle_complete" = oi[3],
  "organelle_debris"   = oi[4],
  "rDNA"               = oi[1],
  "contaminant"        = oi[7],
  "debris"             = oi[9],
  "unclassified"       = oi[8]
)

df_depth <- summary_df %>%
  filter(!is.na(depth_mean)) %>%
  mutate(
    display_class = case_when(
      contig == "chrC" ~ "chrC",
      contig == "chrM" ~ "chrM",
      classification == "organelle_debris" & assigned_ref_id == "chrC" ~ "chrC_debris",
      classification == "organelle_debris" & assigned_ref_id == "chrM" ~ "chrM_debris",
      TRUE ~ classification
    ),
    display_class = factor(display_class, levels = rev(depth_classification_order)),
    length_mb = as.numeric(length) / 1e6,
    chrom_num = as.integer(str_extract(assigned_chrom_id, "\\d+")),
    chrom_num = if_else(is.na(chrom_num), 999L, chrom_num)
  ) %>%
  arrange(display_class, chrom_num, assigned_chrom_id, desc(length_mb))

# Classification depth violin+box (interactive jitter points)
df_depth_tt <- df_depth %>%
  mutate(tooltip = paste0(
    "Contig: ", contig, "\nClassification: ", display_class,
    "\nLength: ", sprintf("%.2f", length_mb), " Mbp",
    "\nMean depth: ", sprintf("%.1f", depth_mean), "x",
    "\nMedian depth: ", sprintf("%.1f", depth_median), "x",
    "\nBreadth 1x: ", sprintf("%.1f%%", depth_breadth_1x * 100),
    "\nBreadth 10x: ", sprintf("%.1f%%", depth_breadth_10x * 100)))

p_class_depth <- ggplot(df_depth_tt, aes(x = depth_mean, y = display_class, fill = display_class)) +
  geom_violin(alpha = 0.5, scale = "width", width = 0.7) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, width = 0.3) +
  ggiraph::geom_point_interactive(
    aes(color = display_class, tooltip = tooltip),
    position = position_jitter(height = 0.12), alpha = 0.5, size = 1.2, show.legend = FALSE) +
  scale_fill_manual(values = depth_classification_colors, drop = FALSE, guide = "none") +
  scale_color_manual(values = depth_classification_colors, drop = FALSE) +
  scale_y_discrete(labels = function(x) str_replace_all(x, "_", " ")) +
  theme_classic(base_family = base_family, base_size = base_font_pt) +
  axis_theme +
  theme(plot.title = element_text(hjust = 0.5, size = base_font_pt, family = base_family),
        axis.text.y = element_text(size = base_font_pt - 1),
        panel.grid.major.x = element_line(color = "grey90", linewidth = 0.3)) +
  labs(title = paste0("Read depth by classification (", params$plot_suffix, ")"),
       x = "Mean read depth", y = NULL)

# Chromosome depth bars
df_chrom_depth <- df_depth %>%
  filter(classification == "chrom_assigned") %>%
  mutate(
    chrom_id = assigned_chrom_id,
    chrom_label = if_else(
      !is.na(assigned_subgenome) & assigned_subgenome != "" & assigned_subgenome != "NA",
      paste0(chrom_id, assigned_subgenome), chrom_id)
  ) %>%
  arrange(chrom_num, chrom_label, desc(length_mb)) %>%
  mutate(plot_order = row_number())

if (nrow(df_chrom_depth) > 0) {
  chrom_breaks <- df_chrom_depth %>%
    group_by(chrom_label) %>% summarise(mid_pos = mean(plot_order), .groups = "drop")
  df_chrom_depth <- df_chrom_depth %>%
    mutate(fill_category = if_else(
      !is.na(assigned_subgenome) & assigned_subgenome != "NA",
      as.character(assigned_subgenome), "NA"),
      tooltip = paste0("Contig: ", contig, "\nChromosome: ", chrom_label,
        "\nLength: ", sprintf("%.2f", length_mb), " Mbp",
        "\nMean depth: ", sprintf("%.1f", depth_mean), "x"))

  p_chrom_depth <- ggplot(df_chrom_depth, aes(x = depth_mean, y = plot_order)) +
    ggiraph::geom_col_interactive(
      aes(fill = fill_category, tooltip = tooltip),
      width = 0.8, alpha = 0.85, orientation = "y") +
    scale_fill_manual(values = sg_colors, name = "Subgenome",
                      guide = if (has_subgenomes) "legend" else "none") +
    scale_y_reverse(breaks = chrom_breaks$mid_pos,
                    labels = str_replace(chrom_breaks$chrom_label, "^chr", ""),
                    expand = expansion(mult = c(0.01, 0.01))) +
    theme_classic(base_family = base_family, base_size = base_font_pt) +
    axis_theme +
    theme(plot.title = element_text(hjust = 0.5, size = base_font_pt, family = base_family),
          axis.text.y = element_text(size = base_font_pt - 1),
          panel.grid.major.x = element_line(color = "grey90", linewidth = 0.3),
          legend.position = "top") +
    labs(title = "Chromosome contig depth", x = "Mean depth", y = "Chromosome")
  p_chrom_depth <- apply_legend_theme(p_chrom_depth, text_pt = 6, key_pt = 6, tight = TRUE) +
    theme(legend.margin = margin(1, 1, 1, 1))
} else {
  p_chrom_depth <- ggplot() +
    theme_void(base_family = base_family, base_size = base_font_pt) +
    labs(title = "Chromosome depth (no data)")
}

# Breadth coverage boxplots
df_breadth <- df_depth %>%
  select(contig, classification, display_class, depth_breadth_1x, depth_breadth_10x) %>%
  tidyr::pivot_longer(cols = c(depth_breadth_1x, depth_breadth_10x),
                      names_to = "metric", values_to = "breadth") %>%
  mutate(metric = case_when(metric == "depth_breadth_1x" ~ ">=1x",
                            metric == "depth_breadth_10x" ~ ">=10x", TRUE ~ metric),
         metric = factor(metric, levels = c(">=1x", ">=10x")))

p_breadth <- ggplot(df_breadth, aes(x = breadth, y = display_class, fill = metric)) +
  geom_boxplot(alpha = 0.7, position = position_dodge(width = 0.75), width = 0.6, outlier.size = 0.8) +
  scale_fill_manual(values = c(">=1x" = "#66C2A5", ">=10x" = "#1F77B4"), name = "Coverage") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2), labels = scales::percent) +
  scale_y_discrete(labels = function(x) str_replace_all(x, "_", " ")) +
  theme_classic(base_family = base_family, base_size = base_font_pt) +
  axis_theme +
  theme(plot.title = element_text(hjust = 0.5, size = base_font_pt, family = base_family),
        axis.text.y = element_text(size = base_font_pt - 1),
        panel.grid.major.x = element_line(color = "grey90", linewidth = 0.3),
        legend.position = "top") +
  labs(title = "Coverage breadth by classification", x = "Fraction of bases covered", y = NULL)
p_breadth <- apply_legend_theme(p_breadth, text_pt = 6, key_pt = 6, tight = TRUE) +
  theme(legend.margin = margin(1, 1, 1, 1))

# Combine
full_plot_depth <- p_chrom_depth | (p_class_depth / p_breadth) + plot_layout(widths = c(1, 1))
save_pdf(pdf_depth, full_plot_depth, width = 7.2, height = 5)
```

```{r depth_widget, eval=has_depth}
girafe_styled(full_plot_depth, width_svg = 10, height_svg = 6.9)
```

```{r depth_tab_heading, eval=has_depth, results='asis'}
cat('\n## Depth Summary by Category\n')
```

```{r depth_table, eval=has_depth}
depth_by_cat <- summary_df %>%
  filter(!is.na(depth_mean)) %>%
  group_by(classification) %>%
  summarise(
    n = n(), mean_depth = mean(depth_mean, na.rm = TRUE),
    median_depth = median(depth_median, na.rm = TRUE),
    depth_cv = if (mean(depth_mean, na.rm = TRUE) > 0)
                 sd(depth_mean, na.rm = TRUE) / mean(depth_mean, na.rm = TRUE)
               else NA_real_,
    breadth_1x = if ("depth_breadth_1x" %in% names(.)) mean(depth_breadth_1x, na.rm = TRUE) * 100 else NA_real_,
    breadth_10x = if ("depth_breadth_10x" %in% names(.)) mean(depth_breadth_10x, na.rm = TRUE) * 100 else NA_real_,
    .groups = "drop"
  ) %>%
  mutate(
    color_swatch = mapply(function(c) {
      col <- classification_colors[c]; if (is.na(col)) col <- "#999"
      paste0('<span style="display:inline-block;width:10px;height:10px;background:', col,
             ';border-radius:2px;margin-right:4px;"></span>')
    }, classification),
    label = paste0(color_swatch, classification_labels[classification])
  ) %>% arrange(desc(mean_depth))

sel <- c("label", "n", "mean_depth", "median_depth", "depth_cv")
labels <- list(label = "Category", n = "Contigs", mean_depth = "Mean Depth",
               median_depth = "Median Depth", depth_cv = "Depth CV")
has_breadth_col <- "depth_breadth_1x" %in% names(summary_df)
if (has_breadth_col) {
  sel <- c(sel, "breadth_1x", "breadth_10x")
  labels$breadth_1x <- "Breadth >= 1x"; labels$breadth_10x <- "Breadth >= 10x"
}
depth_by_cat %>% select(all_of(sel)) %>% gt() %>% cols_label(.list = labels) %>%
  fmt(columns = label, fns = function(x) x) %>%
  fmt_number(columns = any_of(c("mean_depth", "median_depth")), decimals = 1) %>%
  fmt_number(columns = any_of("depth_cv"), decimals = 2) %>%
  fmt_number(columns = any_of(c("breadth_1x", "breadth_10x")), decimals = 1) %>%
  tab_header(title = md("**Depth Summary by Category**")) %>% gt_theme()
```


```{r organelle_rdna_data}
org_df <- summary_df %>% filter(classification %in% c("organelle_complete", "organelle_debris"))
rdna_df <- summary_df %>% filter(classification == "rDNA")
has_organelles <- nrow(org_df) > 0
has_rdna <- nrow(rdna_df) > 0
```

# 6. rDNA {.tabset}

```{r rdna_stub, eval=!has_rdna, results='asis'}
cat('<div class="note-box">No rDNA contigs were detected in this assembly.</div>')
```

```{r rdna_ann_prep, include=FALSE, eval=has_rdna_ann}
# Map original contig names to new names
name_map <- summary_df %>% select(original_name, contig) %>% distinct()
rdna_ann_show <- rdna_ann_df %>%
  rename(original_name = contig) %>%
  left_join(name_map, by = "original_name") %>%
  mutate(contig = if_else(!is.na(contig), contig, original_name)) %>%
  relocate(contig, original_name)

.render_rdna_ann_gt <- function(df, title, subtitle) {
  df %>%
    select(-array_id) %>%
    rename_with(~ case_when(. == "contig" ~ "contig", . == "original_name" ~ "original_name", TRUE ~ .)) %>%
    gt() %>%
    cols_label(contig = "Contig", original_name = "Original Name") %>%
    tab_header(title = md(paste0("**", title, "**")), subtitle = md(subtitle)) %>%
    tab_options(table.font.size = px(10)) %>% gt_theme()
}

array_df <- rdna_ann_show %>% filter(!is.na(array_id) & nzchar(array_id))
other_df <- rdna_ann_show %>% filter(is.na(array_id) | !nzchar(array_id))
has_array_loci <- nrow(array_df) > 0
has_other_rdna_ann <- nrow(other_df) > 0
```

```{r rdna_arrays_tab, eval=has_rdna_arrays, results='asis'}
cat('\n## 45S rDNA Arrays\n')
```

```{r rdna_arrays_table, eval=has_rdna_arrays}
# Map original contig names to new names for arrays
arrays_show <- rdna_arrays_df %>%
  rename(original_name = contig) %>%
  left_join(name_map, by = "original_name") %>%
  mutate(contig = if_else(!is.na(contig), contig, original_name)) %>%
  relocate(array_id, contig)

# Collect per-array identity values from the per-locus annotations
# for gt_plt_dist boxplots
if (has_rdna_ann) {
  id_by_array <- rdna_ann_df %>%
    filter(!is.na(array_id) & nzchar(array_id)) %>%
    group_by(array_id) %>%
    summarise(id_dist = list(identity), .groups = "drop")
} else {
  id_by_array <- tibble(array_id = character(), id_dist = list())
}

max_span <- max(arrays_show$span_kb, na.rm = TRUE)

arrays_tbl <- arrays_show %>%
  left_join(id_by_array, by = "array_id") %>%
  mutate(
    span_bar = mapply(function(v) gt_inline_bar(v, max_span, color = oi[6], height = "14px"), span_kb),
    span_text = sprintf("%.0f kb", span_kb),
    id_text = sprintf("%.2f%% (%.2f\u2013%.2f%%)",
                       identity_median * 100, identity_min * 100, identity_max * 100),
    id_dist = if (nrow(id_by_array) > 0) id_dist else lapply(seq_len(n()), function(i) {
      c(identity_min[i], identity_median[i], identity_max[i])
    }),
    coords = sprintf("%.2f\u2013%.2f Mb", start / 1e6, end / 1e6),
    copies_text = paste0(n_total, " (", n_full, "F / ", n_partial, "P / ", n_fragment, "Fr)")
  )

arrays_tbl %>%
  select(array_id, contig, original_name, coords, span_bar, span_text,
         copies_text, id_dist, id_text) %>%
  gt() %>%
  cols_label(
    array_id = "Array", contig = "Contig", original_name = "Original Name",
    coords = "Coordinates",
    span_bar = "", span_text = "Span",
    copies_text = "Copies (F/P/Fr)",
    id_dist = "", id_text = "Identity"
  ) %>%
  fmt(columns = span_bar, fns = function(x) x) %>%
  gt_plt_dist(id_dist, type = "boxplot", line_color = "steelblue",
              fill_color = "lightsteelblue") %>%
  cols_move(columns = id_text, after = id_dist) %>%
  tab_header(title = md("**45S rDNA Arrays**"),
    subtitle = md(paste0(nrow(arrays_show), " tandem repeat arrays detected"))) %>%
  gt_theme()
```

<div class="legend-box">
<b>Copies</b>: F&thinsp;=&thinsp;full-length 45S units &emsp; P&thinsp;=&thinsp;partial (truncated at array boundary) &emsp; Fr&thinsp;=&thinsp;fragment (&lt;&thinsp;50% of consensus)
</div>

```{r rdna_summary_tab, eval=has_rdna, results='asis'}
cat('\n## rDNA Debris Summary\n')
```

```{r rdna_table, eval=has_rdna}
# Compute per-contig rDNA coverage from annotations (merge overlapping loci)
rdna_cov <- if (has_rdna_ann) {
  rdna_ann_df %>%
    filter(contig_classification == "rDNA") %>%
    group_by(contig) %>%
    summarise(
      rdna_bp = {
        s <- sort(start); e <- sort(end)
        # Merge overlapping intervals
        merged_bp <- 0; cur_s <- s[1]; cur_e <- e[1]
        if (length(s) > 1) for (i in 2:length(s)) {
          if (s[i] <= cur_e) { cur_e <- max(cur_e, e[i]) }
          else { merged_bp <- merged_bp + cur_e - cur_s; cur_s <- s[i]; cur_e <- e[i] }
        }
        merged_bp + cur_e - cur_s
      }, .groups = "drop")
} else { tibble(contig = character(), rdna_bp = numeric()) }

rdna_show <- rdna_df %>%
  mutate(length_kb = length / 1e3) %>%
  left_join(rdna_cov, by = c("original_name" = "contig")) %>%
  mutate(rdna_pct = ifelse(!is.na(rdna_bp), rdna_bp / length * 100, NA_real_),
         conf_html = sapply(classification_confidence, conf_badge))

rdna_show %>%
  select(contig, original_name, length_kb, rdna_pct, conf_html) %>%
  gt() %>%
  cols_label(contig = "Contig", original_name = "Original Name",
             length_kb = "Length (kb)", rdna_pct = "rDNA %",
             conf_html = "Confidence") %>%
  fmt_number(columns = length_kb, decimals = 1) %>%
  fmt_number(columns = rdna_pct, decimals = 1) %>%
  fmt(columns = conf_html, fns = function(x) x) %>%
  sub_missing(missing_text = "\u2014") %>%
  tab_header(title = md("**rDNA Debris Summary**")) %>% gt_theme()
```

```{r rdna_other_tab, eval=has_rdna_ann && has_other_rdna_ann, results='asis'}
cat('\n## Non-Array rDNA Annotations\n')
```

```{r rdna_other_table, eval=has_rdna_ann && has_other_rdna_ann}
.render_rdna_ann_gt(other_df %>% head(50), "Other rDNA Annotations",
  paste0("Showing up to 50 of ", nrow(other_df), " non-array annotations"))
```

# 7. Organelles

```{r organelle_stub, eval=!has_organelles, results='asis'}
cat('<div class="note-box">No organelle contigs were detected in this assembly.</div>')
```

```{r organelle_prep, include=FALSE, eval=has_organelles}
org_show <- org_df %>%
  mutate(
    length_kb = length / 1e3,
    type = case_when(str_detect(assigned_ref_id, "chrC|chloroplast|plastid") ~ "chrC",
                     str_detect(assigned_ref_id, "chrM|mito") ~ "chrM", TRUE ~ "organelle"),
    cov_pct = if ("best_ref_union_frac" %in% names(.)) best_ref_union_frac * 100 else NA_real_,
    depth_val = if (has_depth) depth_mean else NA_real_
  )

org_complete <- org_show %>% filter(classification == "organelle_complete")
org_debris   <- org_show %>% filter(classification == "organelle_debris")
if (has_depth) {
  org_debris <- org_debris %>% arrange(desc(depth_val))
}
has_org_complete <- nrow(org_complete) > 0
has_org_debris <- nrow(org_debris) > 0

.render_org_gt <- function(df, title) {
  sel <- c("contig", "original_name", "type", "length_kb")
  labels <- list(contig = "Contig", original_name = "Original Name", type = "Type", length_kb = "Length (kb)")
  if (has_depth) { sel <- c(sel, "depth_val"); labels$depth_val <- "Depth" }
  df %>% select(all_of(sel)) %>% gt() %>% cols_label(.list = labels) %>%
    fmt_number(columns = length_kb, decimals = 1) %>%
    fmt_number(columns = any_of("depth_val"), decimals = 1) %>%
    tab_header(title = md(paste0("**", title, "**"))) %>% gt_theme()
}
```

```{r organelle_table, eval=has_organelles && has_org_complete}
.render_org_gt(org_complete, "Complete Organelles")
```

```{r organelle_debris_table, eval=has_organelles && has_org_debris}
.med_range <- function(vals) {
  if (length(vals) == 0 || all(is.na(vals))) return("\u2014")
  vals <- vals[!is.na(vals)]
  if (length(vals) == 1) return(sprintf("%.1f", vals))
  sprintf("%.1f (%.1f\u2013%.1f)", median(vals), min(vals), max(vals))
}

debris_summary <- org_debris %>%
  group_by(type) %>%
  summarise(
    n = n(),
    total_kb = sum(length_kb),
    length_dist = list(length_kb),
    length_text = .med_range(length_kb),
    depth_dist = if (has_depth) list(depth_val[!is.na(depth_val)]) else list(numeric(0)),
    depth_text = if (has_depth) .med_range(depth_val[!is.na(depth_val)]) else "\u2014",
    .groups = "drop"
  )

gt_tbl <- debris_summary %>%
  gt() %>%
  cols_label(type = "Type", n = "Contigs", total_kb = "Total (kb)",
             length_dist = "", length_text = "Length (kb)",
             depth_dist = "", depth_text = "Depth") %>%
  fmt_number(columns = total_kb, decimals = 1) %>%
  gt_plt_dist(length_dist, type = "boxplot", line_color = "steelblue",
              fill_color = "lightsteelblue") %>%
  cols_move(columns = length_text, after = length_dist) %>%
  tab_header(title = md("**Organelle Debris**"),
             subtitle = md(paste0(nrow(org_debris), " debris contigs summarized by type"))) %>%
  gt_theme()

if (has_depth) {
  gt_tbl <- gt_tbl %>%
    gt_plt_dist(depth_dist, type = "boxplot", line_color = "steelblue",
                fill_color = "lightsteelblue") %>%
    cols_move(columns = depth_text, after = depth_dist)
} else {
  gt_tbl <- gt_tbl %>% cols_hide(c(depth_dist, depth_text))
}
gt_tbl
```

# 8. Contaminants

```{r contaminant_table_build, eval=has_contaminants}
# ---------------------------------------------------------------------------
# Contaminant table
# ---------------------------------------------------------------------------
ct_has_depth <- "depth_mean" %in% names(contam_df) && !all(is.na(contam_df$depth_mean))
ct_df <- contam_df
if (ct_has_depth) ct_df <- ct_df %>% filter(!is.na(depth_mean))

if (nrow(ct_df) > 0) {
  ct_df <- ct_df %>%
    mutate(
      is_circular = str_detect(contig, "c$"),
      species_clean = str_replace_all(species, "_", " "),
      species_clean = if_else(!is.na(species_clean) & str_starts(species_clean, paste0(genus, " ")),
                              str_replace(species_clean, paste0("^", genus, " "), ""), species_clean),
      binomial = case_when(
        is.na(species_clean) | species_clean == "" | species_clean == genus | species_clean == "Unknown" ~
          paste0(genus, " sp."),
        TRUE ~ paste0(genus, " ", species_clean)),
      binomial = str_replace_all(binomial, "_", " "),
      abundance = if (ct_has_depth) depth_mean * length else length
    )

  ct_agg <- ct_df %>%
    group_by(binomial, kingdom, family) %>%
    summarise(
      n_contigs = n(), total_bp = sum(length), abundance = sum(abundance),
      depth_mean_agg = if (ct_has_depth) mean(depth_mean) else NA_real_,
      depth_min = if (ct_has_depth) min(depth_mean) else NA_real_,
      depth_max = if (ct_has_depth) max(depth_mean) else NA_real_,
      length_min = min(length), length_max = max(length),
      n_circular = sum(is_circular), mean_coverage = mean(coverage),
      .groups = "drop") %>%
    arrange(desc(abundance)) %>% head(top_n_contam) %>%
    mutate(rank = row_number())

  if (nrow(ct_agg) > 0) {
    # Domain badges
    domain_info <- tibble(
      kingdom = c("Bacteria", "Archaea", "Viruses", "Animalia", "Plantae", "Fungi", "Chromista", "Protozoa"),
      domain_abbr = c("Bac", "Arc", "Vir", "Ani", "Pla", "Fun", "Chr", "Pro"),
      domain_color = c("#E41A1C", "#984EA3", "#4DAF4A", "#377EB8", "#4DAF4A", "#FF7F00", "#A65628", "#999999"))

    ct_agg <- ct_agg %>%
      left_join(domain_info, by = "kingdom") %>%
      mutate(domain_abbr = if_else(is.na(domain_abbr), "Unk", domain_abbr),
             domain_color = if_else(is.na(domain_color), "#999999", domain_color))

    # Family colors
    families <- unique(ct_agg$family); n_fam <- length(families)
    fam_colors <- if (n_fam <= 8) scales::brewer_pal(palette = "Set2")(max(n_fam, 3))[1:n_fam]
                  else if (n_fam <= 12) scales::brewer_pal(palette = "Set3")(max(n_fam, 3))[1:n_fam]
                  else scales::hue_pal()(n_fam)
    names(fam_colors) <- families

    ct_agg <- ct_agg %>%
      mutate(
        family_color = fam_colors[family],
        circ_indicator = case_when(n_circular == 0 ~ "\u25CB",
                                   n_circular == n_contigs ~ "\u25CF", TRUE ~ "\u25D0"),
        total_mb = total_bp / 1e6,
        total_pct = total_mb / max(total_mb) * 100,
        total_text = if_else(n_contigs == 1, sprintf("%.2f Mb", total_mb),
          sprintf("%.2f Mb (%.2f\u2013%.2f)", total_mb, length_min / 1e6, length_max / 1e6)))

    if (ct_has_depth) {
      max_depth <- max(ct_agg$depth_mean_agg)
      ct_agg <- ct_agg %>%
        mutate(depth_pct = depth_mean_agg / max_depth * 100,
               depth_text = if_else(n_contigs == 1, sprintf("%.0f\u00D7", depth_mean_agg),
                 sprintf("%.0f\u00D7 (%.0f\u2013%.0f)", depth_mean_agg, depth_min, depth_max)))
    }

    # Build gt table
    if (ct_has_depth) {
      tbl_data <- ct_agg %>% select(rank, domain_abbr, domain_color, family, family_color, binomial,
        n_contigs, total_text, total_pct, depth_text, depth_pct, mean_coverage, circ_indicator)
    } else {
      tbl_data <- ct_agg %>% select(rank, domain_abbr, domain_color, family, family_color, binomial,
        n_contigs, total_text, total_pct, mean_coverage, circ_indicator)
    }

    gt_tbl <- tbl_data %>% gt() %>%
      fmt_percent(columns = mean_coverage, decimals = 0) %>%
      text_transform(locations = cells_body(columns = domain_abbr), fn = function(x) {
        paste0("<span style='background-color:", tbl_data$domain_color,
               "; color: white; padding: 2px 4px; border-radius: 3px; font-size: 10px;'>", x, "</span>")
      }) %>%
      text_transform(locations = cells_body(columns = family), fn = function(x) {
        paste0("<span style='border-left: 4px solid ", tbl_data$family_color, "; padding-left: 6px;'>", x, "</span>")
      }) %>%
      text_transform(locations = cells_body(columns = binomial), fn = function(x) {
        paste0("<em>", x, "</em>")
      }) %>%
      text_transform(locations = cells_body(columns = total_text), fn = function(x) {
        mapply(function(val, p) bar_cell(p, oi[7], val, alpha = 0.3),
               x, tbl_data$total_pct, SIMPLIFY = TRUE)
      }) %>%
      cols_align(align = "center", columns = c(rank, n_contigs, circ_indicator, mean_coverage)) %>%
      cols_align(align = "left", columns = total_text) %>%
      gt_theme() %>%
      tab_footnote(footnote = "\u25CF = all circularized, \u25D0 = mixed, \u25CB = all linear",
                   locations = cells_column_labels(columns = circ_indicator)) %>%
      tab_source_note(source_note = md(paste0("**Families:** ",
        paste(sapply(names(fam_colors), function(f) {
          paste0("<span style='border-left: 4px solid ", fam_colors[f],
                 "; padding-left: 4px; margin-right: 12px;'>", f, "</span>")
        }), collapse = " "))))

    if (ct_has_depth) {
      gt_tbl <- gt_tbl %>%
        cols_label(rank = "", domain_abbr = "Domain", family = "Family", binomial = "Species",
                   n_contigs = "n", total_text = "Total Mb", depth_text = "Depth",
                   mean_coverage = "Cov%", circ_indicator = "Circ") %>%
        cols_hide(c(domain_color, family_color, total_pct, depth_pct)) %>%
        text_transform(locations = cells_body(columns = depth_text), fn = function(x) {
          mapply(function(val, p) bar_cell(p, oi[8], val, alpha = 0.3),
                 x, tbl_data$depth_pct, SIMPLIFY = TRUE)
        }) %>%
        cols_align(align = "left", columns = depth_text) %>%
        tab_options(table.width = pct(100)) %>%
        tab_header(title = md(paste0("**Top ", nrow(tbl_data), " Contaminants**")),
                   subtitle = md(paste0("Ranked by abundance (depth \u00D7 length) \u2014 ", params$plot_suffix)))
    } else {
      gt_tbl <- gt_tbl %>%
        cols_label(rank = "", domain_abbr = "Domain", family = "Family", binomial = "Species",
                   n_contigs = "n", total_text = "Total Mb",
                   mean_coverage = "Cov%", circ_indicator = "Circ") %>%
        cols_hide(c(domain_color, family_color, total_pct)) %>%
        tab_options(table.width = pct(100)) %>%
        tab_header(title = md(paste0("**Top ", nrow(tbl_data), " Contaminants**")),
                   subtitle = md(paste0("Ranked by total length \u2014 ", params$plot_suffix)))
    }
    gt_tbl
  }
}
```

```{r no_contam_placeholder, eval=!has_contaminants, results='asis'}
cat('<div class="note-box">No contaminant sequences were detected. If contamination screening was not performed, provide a Centrifuger index with <code>--centrifuger-idx</code>.</div>')
```

# 9. Unclassified & Debris

```{r debris_table}
debris_df <- summary_df %>%
  filter(classification %in% c("unclassified", "debris", "chrom_debris",
                                "chrom_unassigned")) %>%
  mutate(
    length_mb = length / 1e6,
    best_ref = if ("assigned_ref_id" %in% names(.)) assigned_ref_id else NA_character_,
    aligned_bp = if ("best_ref_union_bp" %in% names(.)) best_ref_union_bp else NA_integer_,
    reason = case_when(
      classification == "chrom_unassigned" ~ "Below synteny gates",
      classification == "chrom_debris" ~ "Duplicate of assembled chromosome",
      TRUE ~ classification
    )
  ) %>% arrange(desc(length_mb))

if (nrow(debris_df) > 0) {
  debris_df %>% head(50) %>%
    select(contig, original_name, length_mb, classification, best_ref, aligned_bp, reason) %>%
    gt() %>%
    cols_label(contig = "Contig", original_name = "Original Name", length_mb = "Length (Mb)", classification = "Category",
               best_ref = "Best Ref Hit", aligned_bp = "Aligned Bp", reason = "Reason") %>%
    fmt_number(columns = length_mb, decimals = 2) %>%
    fmt_number(columns = any_of("aligned_bp"), decimals = 0) %>%
    sub_missing(missing_text = "\u2014") %>%
    tab_header(title = md("**Unclassified & Debris**"),
               subtitle = md(paste0("Showing up to 50 of ", nrow(debris_df), " contigs"))) %>%
    gt_theme()
} else {
  htmltools::HTML('<div class="note-box">No unclassified or debris contigs.</div>')
}
```

---

<div style="font-size: 11px; color: #888; border-top: 1px solid #ddd; padding-top: 8px; margin-top: 24px;">
Generated by **dnadis** | `r params$synteny_mode` mode | `r Sys.time()` | `r params$plot_suffix`
</div>
