---
title: "Multi-Assembly Comparison Report"
params:
  comparison_tsv: "__COMPARISON_TSV__"
  completeness_tsv: "__COMPLETENESS_TSV__"
  ref_lengths_tsv: "__REF_LENGTHS_TSV__"
  per_assembly_summary_tsvs: "__PER_ASM_TSVS__"
  assembly_names: "__ASM_NAMES__"
  per_assembly_macro_blocks_tsvs: "__PER_ASM_MACRO_TSVS__"
  per_assembly_rdna_arrays_tsvs: "__PER_ASM_RDNA_TSVS__"
  per_assembly_contaminants_tsvs: "__PER_ASM_CONTAM_TSVS__"
  pairwise_macro_blocks_tsvs: "__PAIRWISE_MACRO_TSVS__"
  pairwise_names: "__PAIRWISE_NAMES__"
  reference_name: "__REFNAME__"
  synteny_mode: "__SYNTENY_MODE__"
  chr_like_minlen: "__CHRLIKE__"
  out_prefix: "__OUTPREFIX__"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
---

```{css, echo=FALSE}
body {
  font-family: "Helvetica Neue", Helvetica, "Liberation Sans", Arial, sans-serif;
  font-size: 13px;
  max-width: 1300px;
  margin: auto;
}
h1.title { font-size: 24px; margin-bottom: 4px; }
h1 { font-size: 20px; border-bottom: 2px solid #2c3e50; padding-bottom: 4px; }
h2 { font-size: 16px; }
.gt_table { font-size: 11px !important; }
.gt_col_heading { font-size: 10px !important; font-weight: bold !important; }
.widget-container { margin: 16px 0; }
.note-box {
  background: #f8f9fa; border-left: 4px solid #6c757d;
  padding: 10px 14px; margin: 12px 0; font-size: 12px; color: #495057;
}
.section-intro { font-size: 13px; color: #555; margin-bottom: 12px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, dpi = 150)

if (!requireNamespace("pacman", quietly = TRUE))
  install.packages("pacman", repos = "https://cloud.r-project.org")
library(pacman)
pacman::p_load(readr, dplyr, stringr, tidyr, tibble,
               ggplot2, ggnewscale, patchwork, grid,
               ggiraph, htmlwidgets, ggrepel,
               ggokabeito, colorspace,
               gt, gtExtras, scales,
               showtext, sysfonts, knitr,
               ggridges)

# ---------------------------------------------------------------------------
# Font setup (shared across all plots)
# ---------------------------------------------------------------------------
.resolve_font <- function(family) {
  slug <- tolower(gsub(" ", "", family))
  path <- systemfonts::match_fonts(family)$path
  if (nzchar(path) && grepl(slug, tolower(gsub("[- ]", "", basename(path)))))
    return(path)
  NULL
}
base_family <- "sans"
for (.fam in c("Helvetica Neue", "Helvetica", "Liberation Sans")) {
  .r <- .resolve_font(.fam)
  if (!is.null(.r)) {
    sysfonts::font_add(.fam,
      regular    = .r,
      bold       = systemfonts::match_fonts(.fam, weight = "bold")$path,
      italic     = systemfonts::match_fonts(.fam, italic = TRUE)$path,
      bolditalic = systemfonts::match_fonts(.fam, italic = TRUE, weight = "bold")$path
    )
    base_family <- .fam
    break
  }
}
showtext_auto()

# ---------------------------------------------------------------------------
# Shared theming
# ---------------------------------------------------------------------------
base_font_pt <- 8
axis_title_margin_pt <- 4
axis_text_margin_pt <- 4

axis_theme <- theme(
  axis.title.x = element_text(size = base_font_pt, family = base_family,
                               margin = margin(t = axis_title_margin_pt)),
  axis.title.y = element_text(size = base_font_pt, family = base_family,
                               margin = margin(r = axis_title_margin_pt)),
  axis.text.x = element_text(size = base_font_pt, family = base_family,
                               margin = margin(t = axis_text_margin_pt)),
  axis.text.y = element_text(size = base_font_pt, family = base_family)
)

apply_legend_theme <- function(p, text_pt = base_font_pt, key_pt = base_font_pt,
                               tight = TRUE) {
  base <- theme(
    legend.title = element_text(size = text_pt, family = base_family),
    legend.text  = element_text(size = text_pt, family = base_family),
    legend.key.height = grid::unit(key_pt, "pt"),
    legend.key.width  = grid::unit(key_pt, "pt"),
    legend.direction = "horizontal",
    legend.box = "horizontal"
  )
  if (tight) {
    base <- base + theme(
      legend.box.margin = margin(0, 0, 0, 0),
      legend.margin = margin(0, 0, 0, 0),
      legend.spacing.y = grid::unit(0, "pt"),
      legend.spacing.x = grid::unit(2, "pt")
    )
  }
  p + base
}

# ---------------------------------------------------------------------------
# Okabe-Ito palette and classification colors
# ---------------------------------------------------------------------------
oi <- palette_okabe_ito()
oi_rgba <- function(col, alpha = 0.25) {
  v <- col2rgb(col)
  sprintf("rgba(%d,%d,%d,%.2f)", v[1], v[2], v[3], alpha)
}
bar_cell <- function(pct, col, label, alpha = 0.25) {
  paste0(
    '<div style="position:relative;padding:2px 6px;white-space:nowrap;border-radius:2px;">',
    sprintf('<div style="position:absolute;inset:0;width:%.1f%%;background-color:%s;border-radius:2px;z-index:0;"></div>', pct, oi_rgba(col, alpha)),
    '<span style="position:relative;z-index:1;">', label, '</span></div>'
  )
}

classification_colors <- c(
  "chrom_assigned"     = oi[5],
  "chrom_unassigned"   = oi[2],
  "organelle_complete" = oi[3],
  "organelle_debris"   = oi[4],
  "rDNA"               = oi[1],
  "contaminant"        = oi[7],
  "chrom_debris"       = lighten(oi[5], amount = 0.5),
  "debris"             = oi[9],
  "unclassified"       = oi[8]
)

classification_labels <- c(
  "chrom_assigned"     = "Chromosome (assigned)",
  "chrom_unassigned"   = "Chromosome (unassigned)",
  "organelle_complete" = "Organelle (complete)",
  "organelle_debris"   = "Organelle (debris)",
  "rDNA"               = "rDNA",
  "contaminant"        = "Contaminant",
  "chrom_debris"       = "Chromosome debris",
  "debris"             = "Debris",
  "unclassified"       = "Unclassified"
)

classification_levels <- c(
  "chrom_assigned", "chrom_unassigned", "organelle_complete", "organelle_debris",
  "rDNA", "contaminant", "chrom_debris", "debris", "unclassified"
)

# ---------------------------------------------------------------------------
# Helpers
# ---------------------------------------------------------------------------
save_pdf <- function(path, plot, width = 7.2, height = 5, dpi = 300) {
  dev <- if (capabilities("cairo")) cairo_pdf else "pdf"
  invisible(ggsave(path, plot, width = width, height = height, dpi = dpi, device = dev))
}

girafe_styled <- function(ggobj, width_svg = 10, height_svg = 6.9) {
  g <- ggiraph::girafe(ggobj = ggobj, width_svg = width_svg, height_svg = height_svg)
  ggiraph::girafe_options(g,
    ggiraph::opts_tooltip(css = paste0(
      "font-family: ", base_family, "; font-size: 9pt; ",
      "background: white; padding: 6px; border: 1px solid #666;"
    )),
    ggiraph::opts_hover(css = "opacity: 0.8;"),
    ggiraph::opts_toolbar(saveaspng = FALSE)
  )
}

gt_theme <- function(gt_obj) {
  gt_obj %>%
    tab_options(
      table.font.size = px(11),
      column_labels.font.weight = "bold",
      column_labels.border.top.style = "none",
      column_labels.border.bottom.width = px(2),
      column_labels.border.bottom.color = "#333",
      table.border.top.style = "none",
      table.border.bottom.style = "none",
      table_body.border.bottom.color = "#999",
      data_row.padding = px(4),
      heading.border.bottom.style = "none"
    ) %>%
    opt_row_striping(row_striping = FALSE)
}

fmt_mb <- function(x) sprintf("%.1f", x / 1e6)
fmt_pct <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.1f%%", x * 100))

# Assembly color palette (one color per assembly)
asm_pal <- c(oi[5], oi[1], oi[3], oi[6], oi[7], oi[4], oi[2], oi[8], oi[9])

# ---------------------------------------------------------------------------
# Subgenome detection and color setup (from unified report)
# ---------------------------------------------------------------------------
ref_lengths <- read_tsv(params$ref_lengths_tsv, show_col_types = FALSE,
                        col_types = cols(subgenome = col_character()))

sg_levels <- ref_lengths %>%
  distinct(subgenome) %>% pull(subgenome) %>% as.character()
sg_levels <- sg_levels[grepl("^[A-Z]$", sg_levels)]
sg_levels <- sort(sg_levels)
n_sg <- length(sg_levels)
if (n_sg > 8) { sg_levels <- sg_levels[1:8]; n_sg <- 8 }
has_subgenomes <- (n_sg >= 2)

pal_sg_dark  <- c(oi[5], oi[1], oi[3], oi[6], oi[7], oi[4], oi[2], oi[8])
pal_sg_light <- lighten(pal_sg_dark, amount = 0.5)
plot_levels  <- if (has_subgenomes) sg_levels else c("G")

sg_cols_dark  <- setNames(pal_sg_dark[seq_along(plot_levels)],  plot_levels)
sg_cols_light <- setNames(pal_sg_light[seq_along(plot_levels)], plot_levels)
col_dark  <- c(sg_cols_dark,  "NA" = pal_sg_dark[1])
col_light <- c(sg_cols_light, "NA" = pal_sg_light[1])

# ---------------------------------------------------------------------------
# Load data
# ---------------------------------------------------------------------------
comp_df <- read_tsv(params$comparison_tsv, show_col_types = FALSE)
compl_df <- read_tsv(params$completeness_tsv, show_col_types = FALSE)

n_asm <- nrow(comp_df)
asm_names <- comp_df$assembly
asm_colors <- setNames(rep_len(asm_pal, n_asm), asm_names)

# Load per-assembly summary TSVs
asm_tsv_paths <- strsplit(params$per_assembly_summary_tsvs, ";")[[1]]
asm_name_list <- strsplit(params$assembly_names, ";")[[1]]

per_asm_summaries <- list()
for (i in seq_along(asm_tsv_paths)) {
  p <- asm_tsv_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(assigned_subgenome = col_character()))
    df$assembly <- asm_name_list[i]
    per_asm_summaries[[i]] <- df
  }
}
all_summary_df <- if (length(per_asm_summaries) > 0) bind_rows(per_asm_summaries) else NULL
has_per_asm <- !is.null(all_summary_df) && nrow(all_summary_df) > 0

has_any_chrom <- any(comp_df$n_chrom_assigned > 0, na.rm = TRUE)
has_any_contam <- any(comp_df$n_contaminants > 0, na.rm = TRUE)
has_any_depth <- any(!is.na(comp_df$mean_chrom_depth) & comp_df$mean_chrom_depth > 0)

chr_like_minlen <- as.numeric(params$chr_like_minlen)

# PDF output paths
pdf_classbar <- paste0(params$out_prefix, ".classification_comparison.pdf")
pdf_heatmap  <- paste0(params$out_prefix, ".chromosome_completeness.pdf")
pdf_synteny  <- paste0(params$out_prefix, ".chromosome_synteny.pdf")

# ---------------------------------------------------------------------------
# Load per-assembly macro_blocks and pairwise macro_blocks for synteny plot
# ---------------------------------------------------------------------------
per_asm_macro_paths <- if (nzchar(params$per_assembly_macro_blocks_tsvs))
  strsplit(params$per_assembly_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_macro_paths <- if (nzchar(params$pairwise_macro_blocks_tsvs))
  strsplit(params$pairwise_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_pair_names <- if (nzchar(params$pairwise_names))
  strsplit(params$pairwise_names, ";")[[1]] else character(0)

# Load per-assembly macro_blocks (ref->asm) and tag with assembly name
per_asm_macro_list <- list()
for (i in seq_along(per_asm_macro_paths)) {
  p <- per_asm_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      per_asm_macro_list[[i]] <- df
    }
  }
}
per_asm_macro_df <- if (length(per_asm_macro_list) > 0) bind_rows(per_asm_macro_list) else NULL

# Load pairwise macro_blocks (asm->asm)
pairwise_macro_list <- list()
for (i in seq_along(pairwise_macro_paths)) {
  p <- pairwise_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0 && i <= length(pairwise_pair_names)) {
      df$pair_name <- pairwise_pair_names[i]
      # Detect rescue pairs (prefixed with "rescue:")
      raw_name <- pairwise_pair_names[i]
      df$is_rescue <- startsWith(raw_name, "rescue:")
      clean_name <- sub("^rescue:", "", raw_name)
      # Extract left and right assembly names from pair name
      parts <- strsplit(clean_name, "_vs_", fixed = TRUE)[[1]]
      if (length(parts) >= 2) {
        df$left_assembly <- parts[1]
        df$right_assembly <- paste(parts[-1], collapse = "_vs_")
      } else {
        warning(sprintf("Cannot parse pairwise pair name: %s", pairwise_pair_names[i]))
        next
      }
      pairwise_macro_list[[i]] <- df
    }
  }
}
pairwise_macro_df <- if (length(pairwise_macro_list) > 0) bind_rows(pairwise_macro_list) else NULL

# Synteny data available for both nucleotide and protein modes
has_synteny_data <- !is.null(per_asm_macro_df) && nrow(per_asm_macro_df) > 0

# ---------------------------------------------------------------------------
# Load per-assembly rDNA arrays TSVs
# ---------------------------------------------------------------------------
rdna_arrays_paths <- if (nzchar(params$per_assembly_rdna_arrays_tsvs))
  strsplit(params$per_assembly_rdna_arrays_tsvs, ";")[[1]] else character(0)

rdna_arrays_list <- list()
for (i in seq_along(rdna_arrays_paths)) {
  p <- rdna_arrays_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE)
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      rdna_arrays_list[[i]] <- df
    }
  }
}
all_rdna_arrays_df <- if (length(rdna_arrays_list) > 0) bind_rows(rdna_arrays_list) else NULL
has_rdna_arrays <- !is.null(all_rdna_arrays_df) && nrow(all_rdna_arrays_df) > 0

# ---------------------------------------------------------------------------
# Load per-assembly contaminants TSVs (for Top Taxa section)
# ---------------------------------------------------------------------------
contam_tsv_paths <- if (nzchar(params$per_assembly_contaminants_tsvs))
  strsplit(params$per_assembly_contaminants_tsvs, ";")[[1]] else character(0)

contam_detail_list <- list()
for (i in seq_along(contam_tsv_paths)) {
  p <- contam_tsv_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE)
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      contam_detail_list[[i]] <- df
    }
  }
}
all_contam_detail_df <- if (length(contam_detail_list) > 0) bind_rows(contam_detail_list) else NULL
has_contam_detail <- !is.null(all_contam_detail_df) && nrow(all_contam_detail_df) > 0
```

---

<p class="section-intro">
**Reference**: `r if (nchar(params$reference_name) > 0) params$reference_name else "\u2014"` &emsp;
**Synteny mode**: `r params$synteny_mode` &emsp;
**Assemblies**: `r n_asm` &emsp;
**Date**: `r Sys.Date()`
</p>

# 1. Assembly Overview {.tabset}

```{r compleasm_setup, include=FALSE}
# Check if compleasm data is available in comparison TSV
has_compleasm <- "compleasm_chrs_S" %in% names(comp_df) &&
  any(!is.na(comp_df$compleasm_chrs_S) & comp_df$compleasm_chrs_S != "", na.rm = TRUE)
```

## Classification

```{r classification_table, eval=has_per_asm, results='asis'}
# Summarise per assembly × classification
clf_wide <- all_summary_df %>%
  group_by(assembly, classification) %>%
  summarise(total_mb = sum(length, na.rm = TRUE) / 1e6,
            n_contigs = n(), .groups = "drop") %>%
  tidyr::pivot_wider(
    names_from = classification,
    values_from = c(total_mb, n_contigs),
    values_fill = list(total_mb = 0, n_contigs = 0L)
  )

# Ensure all classification columns exist (even if no assembly has them)
for (cl in classification_levels) {
  mb_col <- paste0("total_mb_", cl)
  if (!mb_col %in% names(clf_wide)) clf_wide[[mb_col]] <- 0
}

# Build stacked bar SVG
clf_colors <- unname(classification_colors[classification_levels])

make_clf_bar <- function(...) {
  vals <- as.numeric(c(...))
  total <- sum(vals, na.rm = TRUE)
  if (is.na(total) || total == 0) return("")
  bar_w <- 200; bar_h <- 14; x <- 0; rects <- ""
  for (j in seq_along(vals)) {
    w <- vals[j] / total * bar_w
    if (!is.na(w) && w > 0.1) {
      rects <- paste0(rects, sprintf(
        '<rect x="%.1f" y="0" width="%.1f" height="%d" fill="%s"/>', x, w, bar_h, clf_colors[j]))
      x <- x + w
    }
  }
  sprintf('<svg width="%d" height="%d" style="vertical-align:middle" xmlns="http://www.w3.org/2000/svg">%s</svg>',
          bar_w, bar_h, rects)
}

mb_cols <- paste0("total_mb_", classification_levels)

clf_table <- clf_wide %>%
  mutate(assembly = factor(assembly, levels = asm_names)) %>%
  arrange(assembly) %>%
  rowwise() %>%
  mutate(bar = make_clf_bar(!!!syms(mb_cols))) %>%
  ungroup()

# Select and rename columns for gt
clf_gt_df <- clf_table %>%
  select(Assembly = assembly, bar, all_of(mb_cols))

# Short column labels
clf_short <- c(
  "chrom_assigned"     = "Chr",
  "chrom_unassigned"   = "Unasgn",
  "organelle_complete" = "Org",
  "organelle_debris"   = "OrgD",
  "rDNA"               = "rDNA",
  "contaminant"        = "Cont",
  "chrom_debris"       = "ChrD",
  "debris"             = "Debris",
  "unclassified"       = "Unclf"
)

col_labels <- setNames(
  as.list(c("", clf_short[classification_levels])),
  c("bar", mb_cols)
)

# Drop columns that are all zero
nonzero_mb <- mb_cols[sapply(mb_cols, function(col) any(clf_gt_df[[col]] > 0))]
zero_mb <- setdiff(mb_cols, nonzero_mb)

clf_gt_df %>%
  { if (length(zero_mb) > 0) select(., -all_of(zero_mb)) else . } %>%
  gt() %>%
  fmt(columns = bar, fns = function(x) lapply(x, html)) %>%
  fmt_number(columns = any_of(nonzero_mb), decimals = 1) %>%
  cols_label(.list = col_labels[c("bar", nonzero_mb)]) %>%
  cols_width(bar ~ px(220)) %>%
  cols_width(any_of(nonzero_mb) ~ px(52)) %>%
  cols_align(align = "right", columns = any_of(nonzero_mb)) %>%
  tab_header(
    title = md("**Classification**"),
    subtitle = "Mb per category"
  ) %>%
  tab_source_note(md(paste0(
    paste(sapply(classification_levels[classification_levels %in%
      sub("total_mb_", "", nonzero_mb)], function(cl) {
      sprintf('<span style="display:inline-block;width:12px;height:12px;background:%s;margin-right:3px;vertical-align:middle;"></span>%s',
              classification_colors[cl], classification_labels[cl])
    }), collapse = " &nbsp; ")
  ))) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()
```

`r if (has_compleasm) "## BUSCO Completeness" else ""`

```{r compleasm_table, eval=has_compleasm, results='asis'}
busco_colors <- c(
  "Single-copy"  = "#33a2c4",
  "Duplicated"   = "#225ea8",
  "Fragmented"   = "#fdae61",
  "Interspersed" = "#fee08b",
  "Missing"      = "#d73027"
)

make_busco_bar <- function(s, d, f, i, m) {
  vals <- c(s, d, f, i, m)
  cols <- unname(busco_colors)
  total <- sum(vals, na.rm = TRUE)
  if (is.na(total) || total == 0) return("")
  bar_w <- 140; bar_h <- 14; x <- 0; rects <- ""
  for (j in seq_along(vals)) {
    w <- vals[j] / total * bar_w
    if (!is.na(w) && w > 0) {
      rects <- paste0(rects, sprintf(
        '<rect x="%.1f" y="0" width="%.1f" height="%d" fill="%s"/>', x, w, bar_h, cols[j]))
      x <- x + w
    }
  }
  sprintf('<svg width="%d" height="%d" style="vertical-align:middle" xmlns="http://www.w3.org/2000/svg">%s</svg>',
          bar_w, bar_h, rects)
}

# Build wide-format table: one row per assembly, paired columns for chrs / non-chrs
busco_df <- comp_df %>%
  transmute(
    Assembly = assembly,
    chrs_S = as.numeric(compleasm_chrs_S),
    chrs_D = as.numeric(compleasm_chrs_D),
    chrs_F = as.numeric(compleasm_chrs_F),
    chrs_I = as.numeric(compleasm_chrs_I),
    chrs_M = as.numeric(compleasm_chrs_M),
    non_S  = as.numeric(compleasm_non_chrs_S),
    non_D  = as.numeric(compleasm_non_chrs_D),
    non_F  = as.numeric(compleasm_non_chrs_F),
    non_I  = as.numeric(compleasm_non_chrs_I),
    non_M  = as.numeric(compleasm_non_chrs_M)
  ) %>%
  filter(!is.na(chrs_S)) %>%
  rowwise() %>%
  mutate(
    chrs_bar = make_busco_bar(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M),
    non_bar  = make_busco_bar(non_S, non_D, non_F, non_I, non_M)
  ) %>%
  ungroup() %>%
  mutate(spacer = "") %>%
  select(Assembly,
         chrs_bar, chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
         spacer,
         non_bar, non_S, non_D, non_F, non_I, non_M)

lineage_name <- comp_df$compleasm_lineage[!is.na(comp_df$compleasm_lineage) &
                                           comp_df$compleasm_lineage != ""][1]

busco_df %>%
  gt() %>%
  fmt(columns = c(chrs_bar, non_bar), fns = function(x) lapply(x, html)) %>%
  fmt_number(columns = c(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
                          non_S, non_D, non_F, non_I, non_M),
             decimals = 1) %>%
  cols_label(
    chrs_bar = "", chrs_S = "% S", chrs_D = "D", chrs_F = "F",
    chrs_I = "I", chrs_M = "M",
    spacer = "",
    non_bar = "", non_S = "% S", non_D = "D", non_F = "F",
    non_I = "I", non_M = "M"
  ) %>%
  tab_spanner(label = "Chromosomes",
              columns = c(chrs_bar, chrs_S, chrs_D, chrs_F, chrs_I, chrs_M)) %>%
  tab_spanner(label = "Non-chromosomes",
              columns = c(non_bar, non_S, non_D, non_F, non_I, non_M)) %>%
  cols_width(c(chrs_bar, non_bar) ~ px(160)) %>%
  cols_width(c(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
               non_S, non_D, non_F, non_I, non_M) ~ px(48)) %>%
  cols_width(spacer ~ px(6)) %>%
  cols_align(align = "right",
             columns = c(chrs_S, chrs_D, chrs_F, chrs_I, chrs_M,
                          non_S, non_D, non_F, non_I, non_M)) %>%
  tab_header(
    title = md("**BUSCO Completeness**"),
    subtitle = md(sprintf("Lineage: **%s**", lineage_name))
  ) %>%
  tab_source_note(md(paste0(
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[1], ';margin-right:3px;vertical-align:middle;"></span>Single-copy &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[2], ';margin-right:3px;vertical-align:middle;"></span>Duplicated &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[3], ';margin-right:3px;vertical-align:middle;"></span>Fragmented &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[4], ';margin-right:3px;vertical-align:middle;"></span>Interspersed &nbsp; ',
    '<span style="display:inline-block;width:12px;height:12px;background:', busco_colors[5], ';margin-right:3px;vertical-align:middle;"></span>Missing'
  ))) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()
```

## Summary Table

```{r overview_table}
# Detect whether reference has a plastid (chrC/chrPt) sequence
ref_has_plastid <- any(grepl("^chr[C]$|^chrPt$", ref_lengths$ref_id))
any_asm_has_chrC <- any(comp_df$chrC_found == "yes", na.rm = TRUE)
show_chrC <- ref_has_plastid || any_asm_has_chrC

tbl <- comp_df %>%
  transmute(
    Assembly = assembly,
    Contigs  = total_contigs,
    `Size (Mb)` = round(total_bp / 1e6, 1),
    N50 = round(n50 / 1e6, 2),
    L50 = l50,
    `Largest (Mb)` = round(largest_contig / 1e6, 1),
    `Asgn / Unasgn / Chim` = paste(n_chrom_assigned, "/",
                                     n_chrom_unassigned, "/",
                                     n_chimeric),
    `Full-length` = n_full_length,
    `Both telo` = n_both_telomeres,
    `Mean ref cov` = ifelse(is.na(mean_ref_coverage), "\u2014",
                            sprintf("%.1f%%", mean_ref_coverage * 100)),
    `Mean identity` = ifelse(is.na(mean_identity), "\u2014",
                             sprintf("%.2f%%", mean_identity * 100)),
    Contaminants = n_contaminants,
    rDNA = n_rdna,
    chrC = ifelse(chrC_found == "yes", "\u2713", "\u2014"),
    chrM = ifelse(chrM_found == "yes", "\u2713", "\u2014")
  )

gt_obj <- tbl %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Assembly Comparison Summary") %>%
  cols_align(align = "right", columns = where(is.numeric)) %>%
  tab_spanner(label = "Contiguity", columns = c(Contigs, `Size (Mb)`, N50, L50, `Largest (Mb)`)) %>%
  tab_spanner(label = "Chromosomes", columns = c(`Asgn / Unasgn / Chim`, `Full-length`, `Both telo`, `Mean ref cov`, `Mean identity`))

# Conditionally include chrC in "Other" spanner
if (show_chrC) {
  gt_obj <- gt_obj %>%
    tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrC, chrM))
} else {
  gt_obj <- gt_obj %>%
    cols_hide(columns = chrC) %>%
    tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrM))
}
gt_obj
```

## Nx Curves

```{r nx_curves, fig.width=10, fig.height=5}
if (has_per_asm) {
  # Compute Nx curve for each assembly
  nx_list <- list()
  for (asm in asm_names) {
    asm_df <- all_summary_df %>% filter(assembly == asm)
    lens <- sort(asm_df$length, decreasing = TRUE)
    total <- sum(lens)
    if (total == 0) next
    cum_frac <- cumsum(lens) / total
    # Sample at 1% intervals for smooth curves
    nx_pts <- tibble(nx_pct = 0:100)
    nx_pts$contig_len <- sapply(nx_pts$nx_pct, function(pct) {
      if (pct == 0) return(lens[1])
      idx <- which(cum_frac >= pct / 100)[1]
      if (is.na(idx)) return(lens[length(lens)])
      lens[idx]
    })
    nx_pts$assembly <- asm
    # Compute N50 and N90 for annotation
    n50_val <- nx_pts$contig_len[nx_pts$nx_pct == 50]
    n90_val <- nx_pts$contig_len[nx_pts$nx_pct == 90]
    nx_pts$n50 <- n50_val
    nx_pts$n90 <- n90_val
    nx_list[[asm]] <- nx_pts
  }

  if (length(nx_list) > 0) {
    nx_df <- bind_rows(nx_list) %>%
      mutate(assembly = factor(assembly, levels = asm_names),
             contig_mb = contig_len / 1e6)

    p_nx <- ggplot(nx_df, aes(x = nx_pct, y = contig_mb,
                              color = assembly, group = assembly)) +
      geom_step_interactive(
        aes(tooltip = paste0(assembly, "\nNx", nx_pct, " = ",
                             sprintf("%.2f Mb", contig_mb)),
            data_id = assembly),
        linewidth = 0.8
      ) +
      scale_color_manual(values = asm_colors, name = NULL) +
      labs(x = "Nx (%)", y = "Contig length (Mb)") +
      theme_minimal(base_size = base_font_pt, base_family = base_family) +
      axis_theme +
      theme(legend.position = "bottom")
    p_nx <- apply_legend_theme(p_nx)

    girafe_styled(ggobj = p_nx, width_svg = 10, height_svg = 5)
  } else {
    cat("No contig data available for Nx curves.")
  }
} else {
  cat("Per-assembly contig summary data not available.")
}
```


# 2. Chromosome Assignment {.tabset}

## Overview

```{r completeness_heatmap, fig.width=10, fig.height=max(3, n_asm * 0.35 + 2)}
if (has_any_chrom) {
  # Exclude organelle chromosomes from the heatmap (they have their own section)
  organelle_ids <- ref_lengths %>%
    filter(grepl("^chr[CM]$|^chrPt$|^chrMt$", ref_id)) %>% pull(ref_id)

  # Natural sort reference chromosomes: strip "chr" prefix, sort by subgenome + number
  ref_order <- ref_lengths %>%
    filter(!ref_id %in% organelle_ids) %>%
    mutate(
      chrom_num = as.numeric(str_extract(ref_id, "\\d+")),
      sg = if (has_subgenomes) subgenome else "G"
    ) %>%
    arrange(sg, chrom_num) %>%
    pull(ref_id)

  hm_data <- compl_df %>%
    filter(!ref_id %in% organelle_ids) %>%
    mutate(
      ref_id = factor(ref_id, levels = ref_order),
      assembly = factor(assembly, levels = asm_names),
      best_ref_coverage = as.numeric(best_ref_coverage),
      coverage_pct = ifelse(is.na(best_ref_coverage), 0, best_ref_coverage * 100),
      mean_identity = as.numeric(mean_identity),
      identity_pct = ifelse(is.na(mean_identity), 0, mean_identity * 100)
    )

  # Join subgenome info from ref_lengths
  hm_data <- hm_data %>%
    left_join(ref_lengths %>% select(ref_id, subgenome), by = "ref_id") %>%
    mutate(sg = if (has_subgenomes) subgenome else "G")

  # --- Compute per-cell fill color using per-subgenome gradient ---
  # For each subgenome, gradient from white (0%) to dark subgenome color (100%)
  # Fill intensity is based on mean identity (not reference coverage)
  compute_sg_fill <- function(identity, sg_val) {
    sg_key <- if (sg_val %in% names(sg_cols_dark)) sg_val else "NA"
    dark_col <- sg_cols_dark[sg_key]
    if (is.na(identity) || identity <= 0) return("#f0f0f0")
    frac <- min(identity / 100, 1)
    ramp <- colorRampPalette(c("white", dark_col))(101)
    ramp[round(frac * 100) + 1]
  }

  hm_data$cell_fill <- mapply(compute_sg_fill, hm_data$identity_pct, hm_data$sg)

  # --- Off-target coverage computation ---
  # For each (assembly, ref_chromosome), find blocks from contigs NOT assigned to that ref_id
  has_off_target <- FALSE
  if (!is.null(per_asm_macro_df) && has_per_asm) {
    # Build contig-to-assigned mapping
    contig_assigned <- all_summary_df %>%
      filter(classification == "chrom_assigned") %>%
      select(assembly, contig, assigned_ref_id)

    off_target_df <- per_asm_macro_df %>%
      left_join(contig_assigned,
                by = c("assembly", "contig" = "contig")) %>%
      filter(!is.na(assigned_ref_id), ref_id != assigned_ref_id) %>%
      group_by(assembly, ref_id) %>%
      summarise(off_target_bp = sum(union_bp, na.rm = TRUE), .groups = "drop") %>%
      left_join(ref_lengths %>% select(ref_id, ref_len), by = "ref_id") %>%
      mutate(off_target_pct = off_target_bp / ref_len * 100)

    if (nrow(off_target_df) > 0) {
      has_off_target <- TRUE
      hm_data <- hm_data %>%
        left_join(off_target_df %>% select(assembly, ref_id, off_target_pct),
                  by = c("assembly", "ref_id")) %>%
        mutate(off_target_pct = ifelse(is.na(off_target_pct), 0, off_target_pct))
    }
  }
  if (!has_off_target) {
    hm_data$off_target_pct <- 0
  }

  # --- Compute per-cell telomere status from per-assembly summaries ---
  if (has_per_asm && "has_5p_telomere" %in% names(all_summary_df) &&
      "has_3p_telomere" %in% names(all_summary_df)) {
    telo_info <- all_summary_df %>%
      filter(classification == "chrom_assigned") %>%
      group_by(assembly, assigned_ref_id) %>%
      summarise(
        any_5p = any(has_5p_telomere == "yes", na.rm = TRUE),
        any_3p = any(has_3p_telomere == "yes", na.rm = TRUE),
        .groups = "drop"
      )
    hm_data <- hm_data %>%
      left_join(telo_info, by = c("assembly" = "assembly",
                                   "ref_id" = "assigned_ref_id")) %>%
      mutate(any_5p = ifelse(is.na(any_5p), FALSE, any_5p),
             any_3p = ifelse(is.na(any_3p), FALSE, any_3p))
  } else {
    hm_data <- hm_data %>%
      mutate(any_5p = (has_both_telomeres == "yes"),
             any_3p = (has_both_telomeres == "yes"))
  }

  # Shape status for each cell
  hm_data <- hm_data %>%
    mutate(
      shape_status = case_when(
        n_contigs == 0              ~ "unassigned",
        is_full_length != "yes"     ~ "incomplete",
        any_5p & any_3p             ~ "telo_both",
        any_5p & !any_3p            ~ "telo_5p",
        !any_5p & any_3p            ~ "telo_3p",
        TRUE                        ~ "full_length"
      ),
      multi_label = ifelse(n_contigs > 1, as.character(n_contigs), "")
    )

  # --- Border color: dark subgenome color or red gradient for off-target ---
  compute_offtarget_border <- function(ot_pct) {
    if (is.na(ot_pct) || ot_pct <= 0) return(NA_character_)
    frac <- min(ot_pct / 20, 1)
    ramp <- colorRampPalette(c("#e08080", "#e74c3c"))(101)
    ramp[round(frac * 100) + 1]
  }
  hm_data <- hm_data %>%
    mutate(
      sg_dark_col = unname(sg_cols_dark[ifelse(sg %in% names(sg_cols_dark), sg, "NA")]),
      border_color = ifelse(
        off_target_pct > 0,
        sapply(off_target_pct, compute_offtarget_border),
        darken(sg_dark_col, 0.2)
      ),
      border_lw = ifelse(off_target_pct > 0, 0.6, 0.3),
      tooltip = paste0(
        ref_id, " \u2014 ", assembly, "\n",
        "Contigs: ", n_contigs, "\n",
        "Assigned bp: ", scales::comma(total_assigned_bp), "\n",
        "Best ref coverage: ", fmt_pct(best_ref_coverage), "\n",
        "Mean identity: ", fmt_pct(mean_identity), "\n",
        "Full-length: ", is_full_length, "\n",
        "Telomeres: ", case_when(
          shape_status == "telo_both" ~ "both",
          shape_status == "telo_5p"   ~ "5\u2032 only",
          shape_status == "telo_3p"   ~ "3\u2032 only",
          TRUE ~ "none"
        ),
        ifelse(off_target_pct > 0,
               paste0("\nOff-target: ", sprintf("%.1f%%", off_target_pct)), "")
      )
    )

  # =====================================================================
  # X-axis layout: manual positioning with subgenome gaps
  # =====================================================================
  x_spacing   <- 1.0
  x_group_gap <- 1.8
  ref_layout <- ref_lengths %>%
    filter(!ref_id %in% organelle_ids) %>%
    mutate(
      chrom_num = as.numeric(str_extract(ref_id, "\\d+")),
      sg = if (has_subgenomes) subgenome else "G"
    ) %>%
    arrange(sg, chrom_num)
  sg_groups   <- unique(ref_layout$sg)
  n_sg_groups <- length(sg_groups)

  x_positions <- tibble(ref_id = character(), x_center = numeric(),
                         sg = character(), chrom_label = character())
  x_cursor <- 1.0
  for (sg_i in seq_along(sg_groups)) {
    sg_val  <- sg_groups[sg_i]
    sg_refs <- ref_layout %>% filter(sg == sg_val)
    for (j in seq_len(nrow(sg_refs))) {
      rid <- sg_refs$ref_id[j]
      lbl <- str_replace(rid, "^chr", "")
      if (has_subgenomes) lbl <- str_replace(lbl, "[A-Z]$", "")
      x_positions <- bind_rows(x_positions, tibble(
        ref_id = rid, x_center = x_cursor, sg = sg_val, chrom_label = lbl
      ))
      if (j < nrow(sg_refs)) x_cursor <- x_cursor + x_spacing
    }
    if (sg_i < n_sg_groups) x_cursor <- x_cursor + x_group_gap
  }

  # Y-axis layout
  y_spacing <- 1.0
  y_positions <- tibble(assembly = asm_names,
                         y_center = seq_along(asm_names) * y_spacing)

  # Join positions into hm_data
  hm_data <- hm_data %>%
    mutate(assembly = as.character(assembly),
           ref_id = as.character(ref_id)) %>%
    left_join(x_positions %>% select(ref_id, x_center), by = "ref_id") %>%
    left_join(y_positions, by = "assembly")

  # =====================================================================
  # Polygon construction helpers
  # =====================================================================
  pill_hw     <- 0.30
  rect_hh     <- 0.32   # half-height of core rectangle (straight edges)
  cap_r_val   <- 0.08   # cap extends beyond rect; ratio = 0.32/0.40 = 80%
  jagged_hh   <- 0.34   # slightly taller than rect for zigzag visibility
  n_teeth_val <- 2
  tooth_d_val <- 0.06   # deeper teeth for visibility

  # rect_hh = half-height of core rectangle; caps extend beyond by cr
  make_pill <- function(xc, yc, hw, rh, cr, cap_top, cap_bot, n_arc = 20) {
    rect_top <- yc - rh
    rect_bot <- yc + rh
    x <- c(); y <- c()
    if (cap_top) {
      theta <- seq(0, pi, length.out = n_arc)
      x <- c(x, xc + hw * cos(theta))
      y <- c(y, rect_top - cr * sin(theta))
    } else {
      x <- c(x, xc + hw, xc - hw)
      y <- c(y, rect_top, rect_top)
    }
    x <- c(x, xc - hw); y <- c(y, rect_bot)
    if (cap_bot) {
      theta <- seq(pi, 0, length.out = n_arc)
      x <- c(x, xc + hw * cos(theta))
      y <- c(y, rect_bot + cr * sin(theta))
    } else {
      x <- c(x, xc + hw); y <- c(y, rect_bot)
    }
    tibble(x = x, y = y)
  }

  make_pill_jagged <- function(xc, yc, hw, hh_rect, n_t, tooth_d) {
    top_y <- yc - hh_rect; bot_y <- yc + hh_rect
    tw <- (2 * hw) / n_t
    x <- c(); y <- c()
    for (i in 0:(n_t - 1)) {
      xl <- xc - hw + i * tw; xm <- xl + tw / 2; xr <- xl + tw
      x <- c(x, xl, xm, xr)
      y <- c(y, top_y, top_y + tooth_d, top_y)
    }
    x <- c(x, xc + hw); y <- c(y, bot_y)
    for (i in (n_t - 1):0) {
      xl <- xc - hw + i * tw; xm <- xl + tw / 2; xr <- xl + tw
      x <- c(x, xr, xm, xl)
      y <- c(y, bot_y, bot_y - tooth_d, bot_y)
    }
    tibble(x = x, y = y)
  }

  make_cell_polygon <- function(xc, yc, status) {
    switch(status,
      telo_both   = make_pill(xc, yc, pill_hw, rect_hh, cap_r_val, TRUE, TRUE),
      telo_5p     = make_pill(xc, yc, pill_hw, rect_hh, cap_r_val, TRUE, FALSE),
      telo_3p     = make_pill(xc, yc, pill_hw, rect_hh, cap_r_val, FALSE, TRUE),
      full_length = make_pill(xc, yc, pill_hw, rect_hh, 0, FALSE, FALSE),
      incomplete  = make_pill_jagged(xc, yc, pill_hw, jagged_hh,
                                     n_teeth_val, tooth_d_val),
      unassigned  = make_pill(xc, yc, pill_hw, rect_hh, 0, FALSE, FALSE)
    )
  }

  # =====================================================================
  # Build polygon data frame
  # =====================================================================
  poly_list <- list()
  for (i in seq_len(nrow(hm_data))) {
    row <- hm_data[i, ]
    if (is.na(row$x_center) || is.na(row$y_center)) next
    pts <- make_cell_polygon(row$x_center, row$y_center, row$shape_status)
    pts$cell_id <- i
    pts$fill_color <- if (row$shape_status == "unassigned") "#f0f0f0" else row$cell_fill
    pts$border_color <- if (row$shape_status == "unassigned") NA_character_ else row$border_color
    pts$border_lw <- if (row$shape_status == "unassigned") 0 else row$border_lw
    pts$tooltip <- row$tooltip
    pts$data_id <- paste0(row$ref_id, "_", row$assembly)
    pts$shape_status <- row$shape_status
    poly_list[[i]] <- pts
  }
  all_polys <- bind_rows(poly_list)
  unassigned_polys <- all_polys %>% filter(shape_status == "unassigned")
  assigned_polys   <- all_polys %>% filter(shape_status != "unassigned")
  multi_df <- hm_data %>% filter(multi_label != "")

  # =====================================================================
  # Annotation data
  # =====================================================================
  sg_extent <- x_positions %>%
    group_by(sg) %>%
    summarise(x_min = min(x_center), x_max = max(x_center),
              x_mid = mean(x_center), .groups = "drop")
  y_grid_top   <- min(y_positions$y_center) - 0.5
  chr_label_y1 <- y_grid_top - 0.10   # odd-numbered chromosomes (closer to grid)
  chr_label_y2 <- y_grid_top - 0.35   # even-numbered chromosomes (staggered up)
  x_positions  <- x_positions %>%
    mutate(label_y = ifelse(as.numeric(chrom_label) %% 2 == 0,
                            chr_label_y2, chr_label_y1))
  sep_line_y   <- y_grid_top - 0.65
  sg_label_y   <- y_grid_top - 0.90

  # =====================================================================
  # Compose ggplot
  # =====================================================================
  p_hm <- ggplot()

  # 1. Unassigned placeholders (grey, no border)
  if (nrow(unassigned_polys) > 0) {
    p_hm <- p_hm +
      geom_polygon(data = unassigned_polys,
                   aes(x = x, y = y, group = cell_id),
                   fill = "#f0f0f0", color = NA)
  }

  # 2. Assigned pills (precomputed fill/color via scale_*_identity)
  if (nrow(assigned_polys) > 0) {
    p_hm <- p_hm +
      geom_polygon_interactive(
        data = assigned_polys,
        aes(x = x, y = y, group = cell_id,
            fill = fill_color, color = border_color,
            linewidth = border_lw,
            tooltip = tooltip, data_id = data_id),
        show.legend = FALSE
      ) +
      scale_fill_identity() +
      scale_color_identity() +
      scale_linewidth_identity()
  }

  # 3. Multi-contig count labels (small number inside pill)
  # Adaptive color: white on dark fills, dark subgenome color on light fills
  if (nrow(multi_df) > 0) {
    multi_df <- multi_df %>%
      mutate(
        label_color = mapply(function(fill, sg_val) {
          if (is.na(fill) || fill == "#f0f0f0") return("#333333")
          rgb_vals <- col2rgb(fill)
          luminance <- (0.299 * rgb_vals[1] + 0.587 * rgb_vals[2] + 0.114 * rgb_vals[3]) / 255
          if (luminance > 0.55) {
            sg_key <- if (sg_val %in% names(sg_cols_dark)) sg_val else "NA"
            sg_cols_dark[sg_key]
          } else {
            "white"
          }
        }, cell_fill, sg)
      )
    p_hm <- p_hm +
      geom_text(data = multi_df,
                aes(x = x_center, y = y_center, label = multi_label),
                size = base_font_pt / .pt * 0.8, color = multi_df$label_color,
                fontface = "bold", family = base_family)
  }

  # 4. Subgenome separator lines + labels (polyploid only)
  if (has_subgenomes && n_sg_groups >= 2) {
    sg_line_df <- sg_extent %>%
      mutate(y = sep_line_y,
             color = unname(sg_cols_dark[sg]))
    p_hm <- p_hm +
      geom_segment(data = sg_line_df,
                   aes(x = x_min - 0.4, xend = x_max + 0.4,
                       y = y, yend = y),
                   color = sg_line_df$color, linewidth = 0.5) +
      geom_text(data = sg_extent,
                aes(x = x_mid, y = sg_label_y,
                    label = paste0("Subgenome ", sg)),
                size = base_font_pt / .pt, fontface = "bold",
                family = base_family)
  }

  # 5. Chromosome number labels (staggered odd/even to avoid overlap)
  n_chroms <- length(ref_order)
  p_hm <- p_hm +
    geom_text(data = x_positions,
              aes(x = x_center, y = label_y, label = chrom_label),
              size = base_font_pt / .pt, family = base_family,
              hjust = 0.5)

  # 6. Assembly labels (left of grid)
  x_left_edge <- min(x_positions$x_center) - pill_hw - 0.5
  p_hm <- p_hm +
    geom_text(data = y_positions,
              aes(x = x_left_edge, y = y_center, label = assembly),
              size = base_font_pt / .pt, hjust = 1, family = base_family)

  # Scales and theme
  p_hm <- p_hm +
    scale_y_reverse() +
    coord_cartesian(clip = "off") +
    labs(caption = paste0(
      "Pill = telomeres, rect = full-length, zigzag = incomplete, ",
      "grey = unassigned. Fill intensity = mean identity.",
      if (has_off_target) " Red border = off-target synteny." else ""
    )) +
    theme_void(base_size = base_font_pt, base_family = base_family) +
    theme(
      plot.margin = margin(8, 2, 5,
                           max(25, max(nchar(as.character(asm_names))) * 3)),
      plot.caption = element_text(size = 6, hjust = 0, color = "#666")
    )

  # --- Dynamic sizing ---
  n_chroms <- length(ref_order)
  n_sg_gaps <- max(0, n_sg_groups - 1)
  hm_width  <- max(6, n_chroms * 0.45 + n_sg_gaps * 0.8 + 2.5)
  hm_height <- max(3, n_asm * 0.5 + 2.0)

  save_pdf(pdf_heatmap, p_hm, width = min(7.2, hm_width), height = hm_height)
  girafe_styled(ggobj = p_hm,
                width_svg = min(13, hm_width * 1.4),
                height_svg = hm_height)
} else {
  cat("No chromosome-assigned contigs found in any assembly.")
}
```

<div class="note-box">
<b>Legend</b>:&ensp;
<svg width="14" height="24" style="vertical-align:middle"><rect x="1" y="1" width="12" height="22" rx="6" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&thinsp;Both telomeres&emsp;
<svg width="14" height="24" style="vertical-align:middle"><path d="M1,7 C1,1 13,1 13,7 L13,23 L1,23 Z" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&thinsp;5&#x2032; telomere&emsp;
<svg width="14" height="24" style="vertical-align:middle"><path d="M1,1 L13,1 L13,17 C13,23 1,23 1,17 Z" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&thinsp;3&#x2032; telomere&emsp;
<svg width="14" height="20" style="vertical-align:middle"><rect x="1" y="1" width="12" height="18" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&thinsp;Full-length&emsp;
<svg width="14" height="20" style="vertical-align:middle"><path d="M1,4 L4,1 L7,4 L10,1 L13,4 L13,16 L10,19 L7,16 L4,19 L1,16 Z" fill="#8cc5e3" stroke="#2a6496" stroke-width="0.8"/></svg>&thinsp;Incomplete&emsp;
<svg width="14" height="20" style="vertical-align:middle"><rect x="1" y="1" width="12" height="18" fill="#f0f0f0" stroke="none"/></svg>&thinsp;Unassigned<br>
Fill intensity = reference coverage. Red border = off-target synteny. Small numbers = multi-contig.
</div>

```{r scatter_setup, eval=(has_any_chrom && has_per_asm), include=FALSE}
scatter_df <- all_summary_df %>%
  filter(classification == "chrom_assigned") %>%
  mutate(
    length_mb = length / 1e6,
    assembly = factor(assembly, levels = asm_names)
  )
if (has_subgenomes && "assigned_subgenome" %in% names(scatter_df)) {
  scatter_df <- scatter_df %>%
    mutate(sg = ifelse(is.na(assigned_subgenome) | assigned_subgenome == "NA",
                       "Other", assigned_subgenome))
} else {
  scatter_df$sg <- "All"
}
has_scatter_ident <- "seq_identity_vs_ref" %in% names(scatter_df) &&
                     !all(is.na(scatter_df$seq_identity_vs_ref))
has_scatter_cov   <- "ref_coverage" %in% names(scatter_df) &&
                     !all(is.na(scatter_df$ref_coverage))
```

```{r scatter_identity_header, eval=(has_any_chrom && has_per_asm && exists("has_scatter_ident") && has_scatter_ident), results='asis'}
cat("## Length vs Identity\n\n")
```

```{r scatter_identity, eval=(has_any_chrom && has_per_asm && exists("has_scatter_ident") && has_scatter_ident), fig.width=10, fig.height=6}
si_df <- scatter_df %>% filter(!is.na(seq_identity_vs_ref))

p_si <- ggplot(si_df, aes(x = length_mb, y = seq_identity_vs_ref * 100,
                           color = assembly)) +
  geom_point_interactive(
    aes(tooltip = paste0(assembly, "\n", contig, "\n",
                         ifelse(!is.na(assigned_ref_id),
                                paste0("Assigned: ", assigned_ref_id, "\n"), ""),
                         sprintf("%.1f Mb, %.2f%% identity", length_mb,
                                 seq_identity_vs_ref * 100))),
    alpha = 0.6, size = 2
  ) +
  { if (has_subgenomes && length(unique(si_df$sg)) > 1) {
    facet_wrap(~ sg)
  }} +
  scale_color_manual(values = asm_colors, name = NULL) +
  labs(x = "Contig length (Mb)", y = "Sequence identity vs reference (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(legend.position = "bottom")
p_si <- apply_legend_theme(p_si)

girafe_styled(ggobj = p_si, width_svg = 10, height_svg = 6)
```

```{r scatter_coverage_header, eval=(has_any_chrom && has_per_asm && exists("has_scatter_cov") && has_scatter_cov), results='asis'}
cat("## Length vs Ref Coverage\n\n")
```

```{r scatter_coverage, eval=(has_any_chrom && has_per_asm && exists("has_scatter_cov") && has_scatter_cov), fig.width=10, fig.height=6}
sc_df <- scatter_df %>% filter(!is.na(ref_coverage))

p_sc <- ggplot(sc_df, aes(x = length_mb, y = ref_coverage * 100, color = assembly)) +
  geom_point_interactive(
    aes(tooltip = paste0(assembly, "\n", contig, "\n",
                         ifelse(!is.na(assigned_ref_id),
                                paste0("Assigned: ", assigned_ref_id, "\n"), ""),
                         sprintf("%.1f Mb, %.1f%% ref coverage", length_mb,
                                 ref_coverage * 100))),
    alpha = 0.6, size = 2
  ) +
  { if (has_subgenomes && length(unique(sc_df$sg)) > 1) {
    facet_wrap(~ sg)
  }} +
  scale_color_manual(values = asm_colors, name = NULL) +
  labs(x = "Contig length (Mb)", y = "Reference coverage (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(legend.position = "bottom")
p_sc <- apply_legend_theme(p_sc)

girafe_styled(ggobj = p_sc, width_svg = 10, height_svg = 6)
```

```{r completeness_table_setup, include=FALSE}
# Prepare completeness table data with subgenome info and natural sort
if (has_any_chrom) {
  compl_tbl_base <- compl_df %>%
    mutate(best_ref_coverage = as.numeric(best_ref_coverage),
           mean_identity = as.numeric(mean_identity)) %>%
    left_join(ref_lengths %>% select(ref_id, subgenome), by = "ref_id") %>%
    filter(!grepl("^chr[CM]$|^chrPt$|^chrMt$", ref_id)) %>%
    mutate(
      sg = if (has_subgenomes) subgenome else "G",
      chrom_num = as.numeric(str_extract(ref_id, "\\d+")),
      `Assigned (Mb)` = round(total_assigned_bp / 1e6, 1),
      `Ref cov (%)` = ifelse(is.na(best_ref_coverage), "\u2014",
                             sprintf("%.1f", best_ref_coverage * 100)),
      `Identity (%)` = ifelse(is.na(mean_identity), "\u2014",
                              sprintf("%.2f", mean_identity * 100)),
      `Full-length` = ifelse(is_full_length == "yes", "\u2713", ""),
      `Both telo` = ifelse(has_both_telomeres == "yes", "\u2713", "")
    ) %>%
    arrange(sg, chrom_num) %>%
    mutate(ref_id = factor(ref_id, levels = unique(ref_id)))

  compl_sg_levels <- if (has_subgenomes) sort(unique(compl_tbl_base$sg)) else "G"

  make_compl_gt <- function(df, title) {
    df %>%
      select(ref_id, assembly, n_contigs, `Assigned (Mb)`, `Ref cov (%)`,
             `Identity (%)`, `Full-length`, `Both telo`) %>%
      gt(groupname_col = "ref_id") %>%
      gt_theme() %>%
      tab_header(title = title) %>%
      cols_label(
        assembly = "Assembly", n_contigs = "Contigs",
        `Assigned (Mb)` = "Assigned (Mb)",
        `Ref cov (%)` = "Ref cov (%)",
        `Identity (%)` = "Identity (%)",
        `Full-length` = "Full-length",
        `Both telo` = "Both telo"
      ) %>%
      cols_align(align = "right", columns = c(n_contigs, `Assigned (Mb)`,
                                               `Ref cov (%)`, `Identity (%)`)) %>%
      cols_align(align = "center", columns = c(`Full-length`, `Both telo`))
  }
}
```

`r if (has_any_chrom && has_subgenomes && length(compl_sg_levels) >= 2) "## Completeness Table {.tabset}" else "## Completeness Table"`

```{r completeness_table_single, eval=(has_any_chrom && !(has_subgenomes && length(compl_sg_levels) >= 2))}
make_compl_gt(compl_tbl_base, "Chromosome Completeness by Reference")
```

```{r completeness_table_tabset, eval=(has_any_chrom && has_subgenomes && length(compl_sg_levels) >= 2), results='asis'}
for (sg_val in compl_sg_levels) {
  cat(sprintf("\n### Subgenome %s\n\n", sg_val))
  sg_df <- compl_tbl_base %>% filter(sg == sg_val)
  gt_out <- make_compl_gt(sg_df,
    sprintf("Chromosome Completeness \u2014 Subgenome %s", sg_val))
  # Render gt to HTML and emit inline
  cat(as.character(as_raw_html(gt_out)))
  cat("\n\n")
}
```

```{r completeness_table_none, eval=(!has_any_chrom)}
cat("No chromosome-assigned contigs found in any assembly.")
```


`r if (has_synteny_data) "# 3. Chromosome Synteny" else ""`

```{r synteny_setup, eval=has_synteny_data, include=FALSE}
# =========================================================================
# Section 3: Chromosome Synteny Riparian Plot
#
# Per-subgenome grouped x-axis layout (no faceting).
# Each subgenome gets its own horizontal group:
#   [Ref.A | asm1 | asm2]  ||  [Ref.P | asm3]  ||  [Ref.T | asm4]
# Within each group: ribbons connect adjacent columns.
# Chromosome rows (chr1..chr_n) are shared across all groups.
# =========================================================================

# --- Assembly contig lookup (incl. original_name for name mapping) ---
asm_contig_lookup <- list()
for (i in seq_along(asm_name_list)) {
  nm <- asm_name_list[i]
  df <- per_asm_summaries[[i]]
  if (!is.null(df)) {
    lookup <- df %>%
      filter(classification == "chrom_assigned") %>%
      select(contig, original_name, assigned_ref_id, reversed, length) %>%
      mutate(assembly = nm, reversed = (reversed == "yes"))
    asm_contig_lookup[[nm]] <- lookup
  }
}
asm_lookup_df <- bind_rows(asm_contig_lookup)

# --- Fix per-assembly macro_blocks: map original_name -> new_name ---
if (!is.null(per_asm_macro_df) && nrow(asm_lookup_df) > 0) {
  name_map <- asm_lookup_df %>%
    distinct(assembly, original_name, .keep_all = TRUE) %>%
    select(assembly, original_name, new_name = contig)
  per_asm_macro_df <- per_asm_macro_df %>%
    left_join(name_map, by = c("assembly", "contig" = "original_name")) %>%
    mutate(contig_new = coalesce(new_name, contig))
} else if (!is.null(per_asm_macro_df)) {
  per_asm_macro_df <- per_asm_macro_df %>% mutate(contig_new = contig)
}

# --- Nuclear chromosomes sorted by (subgenome, chrom_num) ---
has_named_subgenomes <- any(
  !is.na(ref_lengths$subgenome) & ref_lengths$subgenome != "NA"
)

if (has_named_subgenomes) {
  nuc_ref <- ref_lengths %>%
    filter(!is.na(subgenome) & subgenome != "NA") %>%
    mutate(chrom_num = as.numeric(str_extract(chrom_id, "\\d+"))) %>%
    arrange(subgenome, chrom_num)
} else {
  macro_rids <- unique(per_asm_macro_df$ref_id)
  nuc_ref <- ref_lengths %>%
    filter(ref_id %in% macro_rids) %>%
    mutate(
      subgenome = "all",
      chrom_num = as.numeric(str_extract(ref_id, "\\d+"))
    ) %>%
    arrange(chrom_num)
}

subgenome_order <- unique(nuc_ref$subgenome)
n_subgenomes <- length(subgenome_order)
ref_chr_order <- nuc_ref$ref_id

# --- Determine assemblies per subgenome ---
sg_assemblies <- list()
for (sg in subgenome_order) {
  sg_ref_ids <- nuc_ref %>% filter(subgenome == sg) %>% pull(ref_id)
  asms_with_sg <- asm_lookup_df %>%
    filter(assigned_ref_id %in% sg_ref_ids) %>%
    pull(assembly) %>% unique()
  sg_assemblies[[sg]] <- asm_name_list[asm_name_list %in% asms_with_sg]
}

# --- X-axis layout: per-subgenome groups with wider inter-group gaps ---
col_spacing <- 2
group_gap   <- 4

col_info <- tibble(
  column = character(), x_center = numeric(),
  subgenome = character(), is_ref = logical()
)
all_col_centers <- numeric()
all_col_labels  <- character()
gap_zones <- list()
gap_meta  <- list()

x_cursor <- 1
for (sg_i in seq_along(subgenome_order)) {
  sg <- subgenome_order[sg_i]
  sg_asms <- sg_assemblies[[sg]]

  ref_col_name <- if (n_subgenomes > 1) paste0("Ref.", sg) else "Reference"
  ref_x <- x_cursor
  col_info <- bind_rows(col_info, tibble(
    column = ref_col_name, x_center = ref_x, subgenome = sg, is_ref = TRUE
  ))
  all_col_centers <- c(all_col_centers, setNames(ref_x, ref_col_name))
  all_col_labels  <- c(all_col_labels, ref_col_name)

  prev_x <- ref_x
  prev_col <- ref_col_name

  for (asm in sg_asms) {
    asm_x <- prev_x + col_spacing
    asm_col_name <- if (n_subgenomes > 1) paste0(asm, ".", sg) else asm
    col_info <- bind_rows(col_info, tibble(
      column = asm_col_name, x_center = asm_x, subgenome = sg, is_ref = FALSE
    ))
    all_col_centers <- c(all_col_centers, setNames(asm_x, asm_col_name))
    all_col_labels  <- c(all_col_labels, asm)

    gap_zones[[length(gap_zones) + 1]] <- c(prev_x + 0.2, asm_x - 0.2)
    gap_meta[[length(gap_meta) + 1]] <- list(
      left_col = prev_col, right_col = asm_col_name,
      left_is_ref = (prev_col == ref_col_name), subgenome = sg,
      left_asm = if (prev_col == ref_col_name) NA_character_ else {
        sub(paste0("\\.", sg, "$"), "", prev_col)
      },
      right_asm = asm
    )
    prev_x <- asm_x
    prev_col <- asm_col_name
  }

  if (sg_i < n_subgenomes) {
    x_cursor <- prev_x + group_gap
  }
}

n_total_cols <- nrow(col_info)

# --- Y-axis layout: shared chromosome rows across all groups ---
chrom_nums <- sort(unique(nuc_ref$chrom_num))
max_len_per_row <- nuc_ref %>%
  group_by(chrom_num) %>%
  summarise(max_len = max(ref_len), .groups = "drop") %>%
  arrange(chrom_num)

total_max_len <- sum(max_len_per_row$max_len)
y_scale <- total_max_len / 15
row_gap <- 0.3

row_layout <- max_len_per_row %>%
  mutate(
    row_height = max_len / y_scale,
    cum_height = cumsum(row_height) - row_height,
    gap_above = (row_number() - 1) * row_gap,
    y_top = cum_height + gap_above
  )

chr_layout <- nuc_ref %>%
  left_join(row_layout %>% select(chrom_num, y_top_row = y_top),
            by = "chrom_num") %>%
  mutate(
    pill_height = ref_len / y_scale,
    y_top = y_top_row
  )

# Pill linewidth (reduced for cleaner look)
pill_lw <- 2.0

# --- Build pill data for all columns ---
# Create pills for ALL chromosome-assigned contigs (not just the largest)
pills_list <- list()

for (sg in subgenome_order) {
  sg_chr <- chr_layout %>% filter(subgenome == sg)
  sg_asms <- sg_assemblies[[sg]]
  ref_col_name <- if (n_subgenomes > 1) paste0("Ref.", sg) else "Reference"
  ref_x <- col_info %>% filter(column == ref_col_name) %>% pull(x_center)

  # Reference pills for this subgenome
  pills_list[[ref_col_name]] <- sg_chr %>%
    transmute(
      column = ref_col_name,
      ref_id, chrom_id, subgenome,
      contig = ref_id, contig_len = ref_len, reversed = FALSE,
      x_center = ref_x,
      y_top, y_bot = y_top + pill_height,
      pill_height,
      tooltip = sprintf("Reference\n%s\n%.2f Mb", ref_id, ref_len / 1e6)
    )

  # Assembly pills for this subgenome — ALL contigs, stacked vertically
  for (asm in sg_asms) {
    asm_col_name <- if (n_subgenomes > 1) paste0(asm, ".", sg) else asm
    asm_x <- col_info %>% filter(column == asm_col_name) %>% pull(x_center)

    lk <- asm_lookup_df %>%
      filter(assembly == asm, assigned_ref_id %in% sg_chr$ref_id) %>%
      arrange(assigned_ref_id, desc(length))

    if (nrow(lk) == 0) next

    # Build pills for each contig, stacking multiple contigs per ref_id
    asm_pills_list <- list()
    for (rid in unique(lk$assigned_ref_id)) {
      rid_contigs <- lk %>% filter(assigned_ref_id == rid)
      ref_row <- sg_chr %>% filter(ref_id == rid)
      if (nrow(ref_row) == 0) next

      y_cursor <- ref_row$y_top[1]
      for (ci in seq_len(nrow(rid_contigs))) {
        ctg <- rid_contigs[ci, ]
        ctg_h <- ctg$length / y_scale
        asm_pills_list[[length(asm_pills_list) + 1]] <- tibble(
          column = asm_col_name,
          ref_id = rid, chrom_id = ref_row$chrom_id[1],
          subgenome = ref_row$subgenome[1],
          contig = ctg$contig, contig_len = ctg$length,
          reversed = ctg$reversed,
          x_center = asm_x,
          y_top = y_cursor,
          pill_height = ctg_h,
          y_bot = y_cursor + ctg_h,
          tooltip = sprintf("%s\n%s\n%.2f Mb", asm, ctg$contig, ctg$length / 1e6)
        )
        y_cursor <- y_cursor + ctg_h
      }
    }
    if (length(asm_pills_list) > 0) {
      pills_list[[asm_col_name]] <- bind_rows(asm_pills_list)
    }
  }
}

all_pills <- bind_rows(pills_list)
all_pills$column <- factor(all_pills$column, levels = col_info$column)

# --- Chromosome color palette (by chrom_id, shared across subgenomes) ---
unique_chroms <- unique(nuc_ref$chrom_id)
unique_chroms <- unique_chroms[order(as.numeric(str_extract(unique_chroms, "\\d+")))]
n_unique_chroms <- length(unique_chroms)
if (n_unique_chroms <= 9) {
  chrom_pal <- setNames(oi[seq_len(n_unique_chroms)], unique_chroms)
} else {
  chrom_pal <- setNames(
    colorRampPalette(oi[1:9])(n_unique_chroms),
    unique_chroms
  )
}

# --- Helpers ---
pos_to_y <- function(pos, seq_len, pill_y_top, pill_h) {
  pill_y_top + (pos / seq_len) * pill_h
}

make_ribbon <- function(gap_left, gap_right,
                        left_y_top, left_y_bot,
                        right_y_top, right_y_bot,
                        chrom_id_val, alpha_val, tooltip_text,
                        subgenome_val, data_id_val, n_pts = 20) {
  t <- seq(0, 1, length.out = n_pts)
  s <- 0.5 - 0.5 * cos(pi * t)
  x_fwd <- gap_left + t * (gap_right - gap_left)
  top_fwd <- left_y_top + s * (right_y_top - left_y_top)
  bot_fwd <- left_y_bot + s * (right_y_bot - left_y_bot)
  tibble(
    x = c(x_fwd, rev(x_fwd)),
    y = c(top_fwd, rev(bot_fwd)),
    chrom_id = chrom_id_val,
    alpha = alpha_val,
    tooltip = tooltip_text,
    subgenome = subgenome_val,
    data_id = data_id_val
  )
}

# --- Build ribbon polygons ---
ribbon_polys <- list()
ribbon_idx <- 0

pill_lookup <- all_pills %>%
  select(column, ref_id, chrom_id, contig, contig_len, reversed,
         x_center, y_top, y_bot, pill_height, subgenome)

# Fixed min_span (50kb) — proportional scaling is too aggressive for large genomes
min_span <- 50000

for (gap_i in seq_along(gap_zones)) {
  gz <- gap_zones[[gap_i]]
  gm <- gap_meta[[gap_i]]
  left_col_name  <- gm$left_col
  right_col_name <- gm$right_col
  sg <- gm$subgenome

  sg_ref_ids <- nuc_ref %>% filter(subgenome == sg) %>% pull(ref_id)

  # --- Gather blocks for this gap ---
  if (gm$left_is_ref) {
    right_asm <- gm$right_asm
    if (!is.null(per_asm_macro_df)) {
      macro_src <- per_asm_macro_df %>%
        filter(assembly == right_asm, ref_id %in% sg_ref_ids)
    } else {
      macro_src <- NULL
    }
  } else {
    left_asm  <- gm$left_asm
    right_asm <- gm$right_asm
    if (!is.null(pairwise_macro_df)) {
      macro_src <- pairwise_macro_df %>%
        filter(left_assembly == left_asm, right_assembly == right_asm)
      if (n_subgenomes > 1) {
        macro_src <- macro_src %>% filter(subgenome == sg)
      }
    } else {
      macro_src <- NULL
    }
  }

  if (is.null(macro_src) || nrow(macro_src) == 0) next

  macro_src <- macro_src %>% filter(qspan_bp >= min_span)
  if (nrow(macro_src) == 0) next

  # --- Iterate per block ---
  for (b in seq_len(nrow(macro_src))) {
    blk <- macro_src[b, ]

    if (gm$left_is_ref) {
      lp <- pill_lookup %>% filter(column == left_col_name, contig == blk$ref_id)
      rp <- pill_lookup %>% filter(column == right_col_name, contig == blk$contig_new)
    } else {
      lp <- pill_lookup %>% filter(column == left_col_name, contig == blk$ref_id)
      rp <- pill_lookup %>% filter(column == right_col_name, contig == blk$contig)
    }
    if (nrow(lp) == 0 || nrow(rp) == 0) next

    left_len  <- lp$contig_len[1]
    left_y    <- lp$y_top[1]
    left_h    <- lp$pill_height[1]
    right_len <- rp$contig_len[1]
    right_y   <- rp$y_top[1]
    right_h   <- rp$pill_height[1]

    cid <- as.character(lp$chrom_id[1])

    l_start <- as.numeric(blk$ref_start)
    l_end   <- as.numeric(blk$ref_end)

    r_start <- as.numeric(blk$qstart)
    r_end   <- as.numeric(blk$qend)
    if (gm$left_is_ref && isTRUE(rp$reversed[1])) {
      r_start_new <- right_len - r_end
      r_end_new   <- right_len - r_start
      r_start <- r_start_new
      r_end   <- r_end_new
    }

    l_start <- max(0, l_start); l_end <- min(left_len, l_end)
    r_start <- max(0, r_start); r_end <- min(right_len, r_end)
    if (l_end <= l_start || r_end <= r_start) next

    ly_top <- pos_to_y(l_start, left_len, left_y, left_h)
    ly_bot <- pos_to_y(l_end, left_len, left_y, left_h)
    ry_top <- pos_to_y(r_start, right_len, right_y, right_h)
    ry_bot <- pos_to_y(r_end, right_len, right_y, right_h)

    strand_val <- blk$strand
    if (gm$left_is_ref && isTRUE(rp$reversed[1])) {
      strand_val <- if (strand_val == "+") "-" else "+"
    }
    if (strand_val == "-") {
      tmp <- ry_top; ry_top <- ry_bot; ry_bot <- tmp
    }

    ident_val <- as.numeric(blk$identity)
    # Higher base opacity for better visibility
    alpha_ribbon <- 0.8 + 0.15 * min(1, ident_val)

    left_key  <- lp$contig[1]
    right_key <- rp$contig[1]
    tooltip <- sprintf(
      "%s \u2192 %s\n%.2f Mb\nIdentity: %.1f%%\nStrand: %s",
      left_key, right_key,
      as.numeric(blk$union_bp) / 1e6,
      ident_val * 100, strand_val
    )

    ribbon_idx <- ribbon_idx + 1
    ribbon_polys[[ribbon_idx]] <- make_ribbon(
      gz[1], gz[2], ly_top, ly_bot, ry_top, ry_bot,
      chrom_id_val = cid, alpha_val = alpha_ribbon,
      tooltip_text = tooltip, subgenome_val = sg,
      data_id_val = cid
    ) %>% mutate(group_id = ribbon_idx)
  }
}

# --- Rescue ribbons: span across non-adjacent assemblies ---
# When an intermediate assembly lacks a chromosome, rescue pairwise alignments
# bridge the gap.  Rescue pairs are tagged with is_rescue == TRUE.
rescue_macro_df <- if (!is.null(pairwise_macro_df) && "is_rescue" %in% names(pairwise_macro_df)) {
  pairwise_macro_df %>% filter(is_rescue)
} else { NULL }

if (!is.null(rescue_macro_df) && nrow(rescue_macro_df) > 0) {
  rescue_pairs <- rescue_macro_df %>%
    distinct(left_assembly, right_assembly)

  for (rp_i in seq_len(nrow(rescue_pairs))) {
    r_left_asm  <- rescue_pairs$left_assembly[rp_i]
    r_right_asm <- rescue_pairs$right_assembly[rp_i]

    # Filter once per pair outside the subgenome loop
    pair_df <- rescue_macro_df %>%
      filter(left_assembly == r_left_asm, right_assembly == r_right_asm,
             qspan_bp >= min_span)
    if (nrow(pair_df) == 0) next

    for (sg in subgenome_order) {
      left_col_name  <- if (n_subgenomes > 1) paste0(r_left_asm, ".", sg) else r_left_asm
      right_col_name <- if (n_subgenomes > 1) paste0(r_right_asm, ".", sg) else r_right_asm

      # Look up x-coordinates; skip if either column is missing
      if (!(left_col_name %in% names(all_col_centers)) ||
          !(right_col_name %in% names(all_col_centers))) next
      left_x  <- all_col_centers[[left_col_name]]
      right_x <- all_col_centers[[right_col_name]]
      gz_rescue <- c(left_x + 0.2, right_x - 0.2)

      macro_src <- if (n_subgenomes > 1) filter(pair_df, subgenome == sg) else pair_df
      if (nrow(macro_src) == 0) next

      for (b in seq_len(nrow(macro_src))) {
        blk <- macro_src[b, ]

        lp <- pill_lookup %>% filter(column == left_col_name, contig == blk$ref_id)
        rp <- pill_lookup %>% filter(column == right_col_name, contig == blk$contig)
        if (nrow(lp) == 0 || nrow(rp) == 0) next

        left_len  <- lp$contig_len[1]
        left_y    <- lp$y_top[1]
        left_h    <- lp$pill_height[1]
        right_len <- rp$contig_len[1]
        right_y   <- rp$y_top[1]
        right_h   <- rp$pill_height[1]

        cid <- as.character(lp$chrom_id[1])

        l_start <- max(0, as.numeric(blk$ref_start))
        l_end   <- min(left_len, as.numeric(blk$ref_end))
        r_start <- max(0, as.numeric(blk$qstart))
        r_end   <- min(right_len, as.numeric(blk$qend))
        if (l_end <= l_start || r_end <= r_start) next

        ly_top <- pos_to_y(l_start, left_len, left_y, left_h)
        ly_bot <- pos_to_y(l_end, left_len, left_y, left_h)
        ry_top <- pos_to_y(r_start, right_len, right_y, right_h)
        ry_bot <- pos_to_y(r_end, right_len, right_y, right_h)

        strand_val <- blk$strand
        if (strand_val == "-") {
          tmp <- ry_top; ry_top <- ry_bot; ry_bot <- tmp
        }

        ident_val <- as.numeric(blk$identity)
        alpha_ribbon <- 0.8 + 0.15 * min(1, ident_val)

        left_key  <- lp$contig[1]
        right_key <- rp$contig[1]
        tooltip <- sprintf(
          "%s \u2192 %s\n%.2f Mb\nIdentity: %.1f%%\nStrand: %s",
          left_key, right_key,
          as.numeric(blk$union_bp) / 1e6,
          ident_val * 100, strand_val
        )

        ribbon_idx <- ribbon_idx + 1
        ribbon_polys[[ribbon_idx]] <- make_ribbon(
          gz_rescue[1], gz_rescue[2], ly_top, ly_bot, ry_top, ry_bot,
          chrom_id_val = cid, alpha_val = alpha_ribbon,
          tooltip_text = tooltip, subgenome_val = sg,
          data_id_val = cid
        ) %>% mutate(group_id = ribbon_idx, is_rescue = TRUE)
      }
    }
  }
}

has_ribbons <- length(ribbon_polys) > 0
if (has_ribbons) {
  all_ribbons <- bind_rows(ribbon_polys)
  # Ensure is_rescue column exists; fill NA (standard ribbons) with FALSE
  if (!"is_rescue" %in% names(all_ribbons)) all_ribbons$is_rescue <- FALSE
  all_ribbons$is_rescue[is.na(all_ribbons$is_rescue)] <- FALSE
  # Sort so rescue ribbons are drawn first (behind standard ribbons)
  all_ribbons <- all_ribbons %>% arrange(desc(is_rescue), group_id)
} else {
  all_ribbons <- NULL
}

# --- Plot sizing ---
n_chrom_rows <- length(chrom_nums)
syn_plot_h <- max(4, n_chrom_rows * 0.5 + 2)
syn_plot_w <- max(7.2, n_total_cols * 1.8 + 2)
```

```{r synteny_description, eval=has_synteny_data, results='asis'}
cat('<p class="section-intro">\n')
cat(sprintf("Chromosome synteny between the reference and each assembly (%s mode). ",
            params$synteny_mode))
if (exists("n_subgenomes") && n_subgenomes > 1) {
  cat("Assemblies are grouped by reference subgenome. ")
}
cat("Ribbons connect syntenic macro-blocks between adjacent columns. ")
cat("Color indicates the reference chromosome; opacity reflects alignment identity. ")
cat("Crossed ribbons indicate inversions.\n")
cat("</p>\n")
```

```{r synteny_riparian, eval=(has_synteny_data && exists("has_ribbons") && has_ribbons), fig.width=10, fig.height=10}
# --- Chromosome labels on far left ---
first_sg <- subgenome_order[1]
first_ref_col <- if (n_subgenomes > 1) paste0("Ref.", first_sg) else "Reference"
first_ref_x <- col_info %>% filter(column == first_ref_col) %>% pull(x_center)
chr_label_df <- chr_layout %>%
  filter(subgenome == first_sg) %>%
  mutate(
    x = first_ref_x - 0.4,
    y_mid = y_top + pill_height / 2,
    label = chrom_id
  )

# --- Compose the plot ---
p_syn <- ggplot()

# Ribbons (behind pills)
if (has_ribbons) {
  p_syn <- p_syn +
    geom_polygon_interactive(
      data = all_ribbons,
      aes(x = x, y = y, group = group_id, fill = chrom_id,
          alpha = alpha, tooltip = tooltip, data_id = data_id),
      show.legend = FALSE
    ) +
    scale_fill_manual(values = chrom_pal, guide = "none") +
    scale_alpha_identity()
}

# Reference pills (darker)
ref_pill_data <- all_pills %>%
  semi_join(col_info %>% filter(is_ref), by = "column")
p_syn <- p_syn +
  geom_segment(
    data = ref_pill_data,
    aes(x = x_center, xend = x_center, y = y_top, yend = y_bot),
    linewidth = pill_lw, lineend = "round", color = "grey55"
  )

# Assembly pills (lighter, interactive)
asm_pill_data <- all_pills %>%
  semi_join(col_info %>% filter(!is_ref), by = "column")
if (nrow(asm_pill_data) > 0) {
  p_syn <- p_syn +
    geom_segment_interactive(
      data = asm_pill_data,
      aes(x = x_center, xend = x_center, y = y_top, yend = y_bot,
          tooltip = tooltip, data_id = chrom_id),
      linewidth = pill_lw, lineend = "round", color = "grey78"
    )
}

# Chromosome labels on left
p_syn <- p_syn +
  geom_text(data = chr_label_df,
            aes(x = x, y = y_mid, label = label),
            size = 2, hjust = 1, family = base_family)

# Subgenome group labels
if (n_subgenomes > 1) {
  sg_label_df <- col_info %>%
    group_by(subgenome) %>%
    summarise(x_mid = mean(x_center), .groups = "drop") %>%
    mutate(y = min(row_layout$y_top) - 0.6,
           label = paste0("Subgenome ", subgenome))

  p_syn <- p_syn +
    geom_text(data = sg_label_df,
              aes(x = x_mid, y = y, label = label),
              size = 2.8, fontface = "bold", family = base_family)
}

# Axis scales
p_syn <- p_syn +
  scale_x_continuous(
    breaks = as.numeric(all_col_centers),
    labels = all_col_labels,
    expand = expansion(add = c(1.5, 0.5))
  ) +
  scale_y_reverse(expand = expansion(mult = c(0.02, 0.05)))

# Theme
p_syn <- p_syn +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  theme(
    axis.text.x = element_text(size = base_font_pt, face = "bold",
                                family = base_family, angle = 45, hjust = 1),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(5, 5, 5, 10)
  )

save_pdf(pdf_synteny, p_syn, width = 7.2, height = syn_plot_h)

# Use ggiraph directly to add opts_hover_inv for chromosome highlighting
g_syn <- ggiraph::girafe(ggobj = p_syn,
                         width_svg = min(13, syn_plot_w),
                         height_svg = syn_plot_h)
ggiraph::girafe_options(g_syn,
  ggiraph::opts_tooltip(css = paste0(
    "font-family: ", base_family, "; font-size: 9pt; ",
    "background: white; padding: 6px; border: 1px solid #666;"
  )),
  ggiraph::opts_hover(css = "opacity: 0.8;"),
  ggiraph::opts_hover_inv(css = "opacity: 0.15;"),
  ggiraph::opts_toolbar(saveaspng = FALSE)
)
```

```{r synteny_empty, eval=(has_synteny_data && !(exists("has_ribbons") && has_ribbons))}
cat("Synteny data available but no ribbon connections could be built. ",
    "This may indicate that assemblies have very different chromosome structures.")
```


`r if (has_any_chrom && has_per_asm) "# 4. Quality Comparison {.tabset}" else ""`

```{r quality_section, eval=(has_any_chrom && has_per_asm)}
chrom_df <- all_summary_df %>%
  filter(classification == "chrom_assigned")

# Join subgenome info
if (has_subgenomes && "assigned_subgenome" %in% names(chrom_df)) {
  chrom_df <- chrom_df %>%
    mutate(sg = ifelse(is.na(assigned_subgenome) | assigned_subgenome == "NA",
                       "Other", assigned_subgenome))
} else {
  chrom_df$sg <- "All"
}

has_identity <- "seq_identity_vs_ref" %in% names(chrom_df) &&
                !all(is.na(chrom_df$seq_identity_vs_ref))
has_collin   <- "collinearity_score" %in% names(chrom_df) &&
                !all(is.na(chrom_df$collinearity_score))
has_gc_dev   <- "gc_deviation" %in% names(chrom_df) &&
                !all(is.na(chrom_df$gc_deviation))

# Compute composite sort key for assembly ordering
# Primary: median identity across first subgenome, secondary: others
asm_sort_key <- chrom_df %>%
  filter(!is.na(seq_identity_vs_ref)) %>%
  group_by(assembly) %>%
  summarise(median_ident = median(seq_identity_vs_ref, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(desc(median_ident))
asm_quality_order <- if (nrow(asm_sort_key) > 0) asm_sort_key$assembly else asm_names

n_sg_facets <- length(unique(chrom_df$sg))
```

```{r quality_identity, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), results='asis'}
cat("## Sequence Identity\n\n")
```

```{r quality_identity_plot, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
ident_df <- chrom_df %>%
  filter(!is.na(seq_identity_vs_ref)) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_quality_order)),
         value = seq_identity_vs_ref * 100)

p_ident <- ggplot(ident_df, aes(x = value, y = assembly, fill = sg, color = sg)) +
  geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                      rel_min_height = 0.01,
                      quantile_lines = TRUE, quantiles = 2) +
  { if (has_subgenomes && n_sg_facets > 1) {
    list(
      facet_wrap(~ sg, ncol = n_sg_facets),
      scale_fill_manual(values = sg_cols_light, guide = "none"),
      scale_color_manual(values = sg_cols_dark, guide = "none")
    )
  } else {
    list(
      scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
      scale_color_manual(values = c("All" = oi[5]), guide = "none")
    )
  }} +
  labs(x = "Sequence identity vs reference (%)", y = NULL) +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_ident, height_svg = max(3, n_asm * 0.3 + 1.5))
```

```{r quality_collinearity, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), results='asis'}
cat("## Collinearity\n\n")
```

```{r quality_collinearity_plot, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
collin_df <- chrom_df %>%
  filter(!is.na(collinearity_score)) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_quality_order)),
         value = collinearity_score * 100)

p_collin <- ggplot(collin_df, aes(x = value, y = assembly, fill = sg, color = sg)) +
  geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                      rel_min_height = 0.01,
                      quantile_lines = TRUE, quantiles = 2) +
  { if (has_subgenomes && n_sg_facets > 1) {
    list(
      facet_wrap(~ sg, ncol = n_sg_facets),
      scale_fill_manual(values = sg_cols_light, guide = "none"),
      scale_color_manual(values = sg_cols_dark, guide = "none")
    )
  } else {
    list(
      scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
      scale_color_manual(values = c("All" = oi[5]), guide = "none")
    )
  }} +
  labs(x = "Collinearity (%)", y = NULL) +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_collin, height_svg = max(3, n_asm * 0.3 + 1.5))
```

```{r quality_gc, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), results='asis'}
cat("## GC Deviation\n\n")
```

```{r quality_gc_plot, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
gc_df <- chrom_df %>%
  filter(!is.na(gc_deviation)) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_quality_order)),
         value = abs(gc_deviation))

p_gc <- ggplot(gc_df, aes(x = value, y = assembly, fill = sg, color = sg)) +
  geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                      rel_min_height = 0.01,
                      quantile_lines = TRUE, quantiles = 2) +
  { if (has_subgenomes && n_sg_facets > 1) {
    list(
      facet_wrap(~ sg, ncol = n_sg_facets),
      scale_fill_manual(values = sg_cols_light, guide = "none"),
      scale_color_manual(values = sg_cols_dark, guide = "none")
    )
  } else {
    list(
      scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
      scale_color_manual(values = c("All" = oi[5]), guide = "none")
    )
  }} +
  labs(x = "|GC deviation| (std devs from reference)", y = NULL) +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_gc, height_svg = max(3, n_asm * 0.3 + 1.5))
```


`r if (has_any_depth) "# 5. Read Depth" else ""`

```{r depth_section, eval=(has_any_depth && has_per_asm), fig.width=10, fig.height=max(3, n_asm * 0.3 + 1.5)}
depth_df <- all_summary_df %>%
  filter(classification == "chrom_assigned", !is.na(depth_mean))

if (nrow(depth_df) > 0) {
  # Add subgenome info
  if (has_subgenomes && "assigned_subgenome" %in% names(depth_df)) {
    depth_df <- depth_df %>%
      mutate(sg = ifelse(is.na(assigned_subgenome) | assigned_subgenome == "NA",
                         "Other", assigned_subgenome))
  } else {
    depth_df$sg <- "All"
  }

  depth_df <- depth_df %>%
    mutate(assembly = factor(assembly, levels = rev(asm_names)))

  n_depth_sg <- length(unique(depth_df$sg))

  p_depth <- ggplot(depth_df, aes(x = depth_mean, y = assembly, fill = sg, color = sg)) +
    geom_density_ridges(alpha = 0.6, scale = 1.4, linewidth = 0.4,
                        rel_min_height = 0.01,
                        quantile_lines = TRUE, quantiles = 2) +
    { if (has_subgenomes && n_depth_sg > 1) {
      list(
        facet_wrap(~ sg, ncol = n_depth_sg),
        scale_fill_manual(values = sg_cols_light, guide = "none"),
        scale_color_manual(values = sg_cols_dark, guide = "none")
      )
    } else {
      list(
        scale_fill_manual(values = c("All" = lighten(oi[5], 0.4)), guide = "none"),
        scale_color_manual(values = c("All" = oi[5]), guide = "none")
      )
    }} +
    labs(x = "Mean read depth (chromosome contigs)", y = NULL) +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(panel.grid.major.y = element_blank())

  girafe_styled(ggobj = p_depth, height_svg = max(3, n_asm * 0.3 + 1.5))
} else {
  cat("No depth data available for chromosome-assigned contigs.")
}
```


`r if (has_per_asm && any(comp_df$n_rdna > 0, na.rm = TRUE)) "# 6. rDNA Comparison {.tabset}" else ""`

```{r rdna_setup, eval=(has_per_asm && any(comp_df$n_rdna > 0, na.rm = TRUE)), include=FALSE}
has_any_rdna <- any(comp_df$n_rdna > 0, na.rm = TRUE)
```

```{r rdna_summary_header, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna), results='asis'}
cat("## Summary\n\n")
```

```{r rdna_summary, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna)}
rdna_tbl <- comp_df %>%
  transmute(
    Assembly = assembly,
    rdna_mb = round(rdna_bp / 1e6, 2),
    `Contigs` = n_rdna,
    `Arrays` = n_rdna_arrays
  )

max_rdna_mb <- max(rdna_tbl$rdna_mb, na.rm = TRUE)

rdna_tbl %>%
  rowwise() %>%
  mutate(
    `Total (Mb)` = {
      pct <- if (max_rdna_mb > 0) min(100, rdna_mb / max_rdna_mb * 100) else 0
      bar_cell(pct, oi[1], sprintf("%.2f", rdna_mb), alpha = 0.3)
    }
  ) %>%
  ungroup() %>%
  select(Assembly, `Total (Mb)`, Contigs, Arrays) %>%
  gt() %>%
  fmt(columns = `Total (Mb)`, fns = function(x) lapply(x, html)) %>%
  cols_align(align = "right", columns = c(Contigs, Arrays)) %>%
  cols_width(`Total (Mb)` ~ px(140)) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  tab_header(title = md("**rDNA Summary**")) %>%
  gt_theme()
```

```{r rdna_arrays_header, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna && has_rdna_arrays), results='asis'}
cat("## Array Detail\n\n")
```

```{r rdna_arrays_table, eval=(has_per_asm && exists("has_any_rdna") && has_any_rdna && has_rdna_arrays)}
arr_tbl <- all_rdna_arrays_df %>%
  transmute(
    Assembly = assembly,
    `Array ID` = if ("array_id" %in% names(.)) array_id else row_number(),
    Contig = contig,
    `N copies` = if ("n_copies" %in% names(.)) n_copies else NA_integer_,
    Strand = if ("strand" %in% names(.)) strand else NA_character_,
    `Total bp` = if ("total_bp" %in% names(.)) total_bp else
                 if ("array_span" %in% names(.)) array_span else NA_integer_
  )

arr_tbl %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = md("**rDNA Array Details**")) %>%
  cols_align(align = "right", columns = c(`N copies`, `Total bp`))
```


`r if (has_per_asm && any(comp_df$chrC_found == "yes" | comp_df$chrM_found == "yes", na.rm = TRUE)) "# 7. Organelle Comparison" else ""`

```{r organelle_section, eval=(has_per_asm && any(comp_df$chrC_found == "yes" | comp_df$chrM_found == "yes", na.rm = TRUE))}
# Extract organelle contigs from per-assembly data
org_df <- all_summary_df %>%
  filter(classification == "organelle_complete") %>%
  mutate(
    org_type = case_when(
      grepl("chrC|Pt|chloro|plast", assigned_ref_id, ignore.case = TRUE) ~ "chrC",
      grepl("chrM|Mt|mito", assigned_ref_id, ignore.case = TRUE) ~ "chrM",
      TRUE ~ "Other"
    )
  ) %>%
  group_by(assembly, org_type) %>%
  summarise(
    n_contigs = n(),
    total_bp = sum(length, na.rm = TRUE),
    contigs = paste(contig, collapse = ", "),
    .groups = "drop"
  ) %>%
  mutate(
    assembly = as.character(assembly),
    total_mb = total_bp / 1e6
  )

# Build wide table with bar cells for chrC and chrM
org_wide <- comp_df %>%
  transmute(Assembly = assembly) %>%
  left_join(
    org_df %>% filter(org_type == "chrC") %>%
      transmute(Assembly = assembly, chrC_bp = total_bp, chrC_contigs = contigs),
    by = "Assembly"
  ) %>%
  left_join(
    org_df %>% filter(org_type == "chrM") %>%
      transmute(Assembly = assembly, chrM_bp = total_bp, chrM_contigs = contigs),
    by = "Assembly"
  ) %>%
  mutate(
    chrC_bp = replace_na(chrC_bp, 0),
    chrM_bp = replace_na(chrM_bp, 0),
    chrC_contigs = replace_na(chrC_contigs, "\u2014"),
    chrM_contigs = replace_na(chrM_contigs, "\u2014")
  )

max_chrC_bp <- max(org_wide$chrC_bp, na.rm = TRUE)
max_chrM_bp <- max(org_wide$chrM_bp, na.rm = TRUE)

# chrM unit: Kb if max >= 1 Mb, else plain bp with commas
chrM_use_kb <- max_chrM_bp >= 1e6
chrM_col_label <- if (chrM_use_kb) "Size (Kb)" else "Size (bp)"

# Chloroplast color: green-tinted, Mitochondria: red-tinted

org_gt_df <- org_wide %>%
  rowwise() %>%
  mutate(
    chrC_size = if (chrC_bp > 0) {
      pct <- min(100, chrC_bp / max_chrC_bp * 100)
      bar_cell(pct, oi[3], formatC(chrC_bp, format = "d", big.mark = ","))
    } else "\u2014",
    chrM_size = if (chrM_bp > 0) {
      pct <- min(100, chrM_bp / max_chrM_bp * 100)
      val_text <- if (chrM_use_kb) {
        formatC(round(chrM_bp / 1e3), format = "d", big.mark = ",")
      } else {
        formatC(chrM_bp, format = "d", big.mark = ",")
      }
      bar_cell(pct, oi[3], val_text)
    } else "\u2014",
    chrC_contig = chrC_contigs,
    chrM_contig = chrM_contigs
  ) %>%
  ungroup() %>%
  select(Assembly, chrC_contig, chrC_size, chrM_contig, chrM_size)

# Dynamic column labels
chrC_size_label <- "Size (bp)"

# Detect plastid visibility (reuse from overview table if available)
show_chrC_org <- exists("show_chrC") && show_chrC

if (show_chrC_org) {
  org_gt_df %>%
    gt() %>%
    fmt(columns = c(chrC_size, chrM_size), fns = function(x) lapply(x, html)) %>%
    cols_label(
      chrC_contig = "Contig", chrC_size = chrC_size_label,
      chrM_contig = "Contig", chrM_size = chrM_col_label
    ) %>%
    cols_width(chrC_size ~ px(140), chrM_size ~ px(140)) %>%
    tab_spanner(label = "Chloroplast", columns = c(chrC_contig, chrC_size)) %>%
    tab_spanner(label = "Mitochondria", columns = c(chrM_contig, chrM_size)) %>%
    tab_style(
      style = css(vertical_align = "middle"),
      locations = cells_body()
    ) %>%
    tab_header(title = md("**Organelle Summary**")) %>%
    gt_theme()
} else {
  org_gt_df %>%
    select(Assembly, chrM_contig, chrM_size) %>%
    gt() %>%
    fmt(columns = chrM_size, fns = function(x) lapply(x, html)) %>%
    cols_label(chrM_contig = "Contig", chrM_size = chrM_col_label) %>%
    cols_width(chrM_size ~ px(140)) %>%
    tab_spanner(label = "Mitochondria", columns = c(chrM_contig, chrM_size)) %>%
    tab_style(
      style = css(vertical_align = "middle"),
      locations = cells_body()
    ) %>%
    tab_header(title = md("**Organelle Summary**")) %>%
    gt_theme()
}
```


`r if (has_any_contam) "# 8. Contamination Comparison {.tabset}" else ""`

`r if (has_any_contam) "## Summary" else ""`

```{r contamination, eval=has_any_contam}
contam_tbl <- comp_df %>%
  filter(n_contaminants > 0) %>%
  transmute(
    Assembly = assembly,
    contam_mb = round(contaminant_bp / 1e6, 2),
    Contigs = n_contaminants,
    Species = n_contaminant_species
  )

max_contam_mb <- max(contam_tbl$contam_mb, na.rm = TRUE)

contam_tbl %>%
  rowwise() %>%
  mutate(
    `Total (Mb)` = {
      pct <- if (max_contam_mb > 0) min(100, contam_mb / max_contam_mb * 100) else 0
      bar_cell(pct, oi[7], sprintf("%.2f", contam_mb))
    }
  ) %>%
  ungroup() %>%
  select(Assembly, `Total (Mb)`, Contigs, Species) %>%
  gt() %>%
  fmt(columns = `Total (Mb)`, fns = function(x) lapply(x, html)) %>%
  cols_align(align = "right", columns = c(Contigs, Species)) %>%
  cols_width(`Total (Mb)` ~ px(140)) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  tab_header(title = md("**Contamination Summary**")) %>%
  gt_theme()
```

`r if (has_any_contam && has_contam_detail) "## Top Taxa" else ""`

```{r top_taxa, eval=(has_any_contam && has_contam_detail), results='asis'}
# --- Top genera by prevalence across assemblies ---
genus_prevalence <- all_contam_detail_df %>%
  filter(nzchar(genus) & !is.na(genus)) %>%
  distinct(assembly, genus) %>%
  group_by(genus) %>%
  summarise(n_assemblies = n(),
            asm_list = paste(sort(assembly), collapse = "\n"),
            .groups = "drop") %>%
  mutate(pct = n_assemblies / n_asm * 100) %>%
  arrange(desc(n_assemblies), genus) %>%
  head(10)

genus_asm_tips <- genus_prevalence$asm_list

genus_gt <- genus_prevalence %>%
  transmute(Genus = genus, Assemblies = n_assemblies, Prevalence = pct) %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = Genus),
    fn = function(x) sprintf(
      '<span title="%s" style="white-space:nowrap;overflow:hidden;text-overflow:ellipsis;display:block;max-width:120px"><i>%s</i> spp.</span>',
      x, x)
  ) %>%
  gt_plt_bar_pct(column = Prevalence, scaled = TRUE,
                 labels = TRUE, fill = oi[7], background = "#e8e8e8") %>%
  text_transform(
    locations = cells_body(columns = Assemblies),
    fn = function(x) mapply(
      function(v, tip) sprintf('<span title="%s">%s</span>', gsub('"', '&quot;', tip), v),
      x, genus_asm_tips[seq_along(x)], SIMPLIFY = TRUE)
  ) %>%
  text_transform(
    locations = cells_body(columns = Prevalence),
    fn = function(x) mapply(
      function(v, tip) sprintf('<span title="%s" style="display:block">%s</span>', gsub('"', '&quot;', tip), v),
      x, genus_asm_tips[seq_along(x)], SIMPLIFY = TRUE)
  ) %>%
  cols_label(Prevalence = "% assemblies") %>%
  cols_align(align = "right", columns = Assemblies) %>%
  cols_width(Genus ~ px(120), Assemblies ~ px(65), Prevalence ~ px(150)) %>%
  tab_header(title = md("**Top Genera**")) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()

# --- Top species by prevalence across assemblies ---
species_prevalence <- all_contam_detail_df %>%
  filter(nzchar(species) & !is.na(species)) %>%
  mutate(binomial = paste(genus, species)) %>%
  distinct(assembly, binomial) %>%
  group_by(binomial) %>%
  summarise(n_assemblies = n(),
            asm_list = paste(sort(assembly), collapse = "\n"),
            .groups = "drop") %>%
  mutate(pct = n_assemblies / n_asm * 100) %>%
  arrange(desc(n_assemblies), binomial) %>%
  head(10)

species_asm_tips <- species_prevalence$asm_list

species_gt <- species_prevalence %>%
  transmute(Species = binomial, Assemblies = n_assemblies, Prevalence = pct) %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = Species),
    fn = function(x) sprintf(
      '<span title="%s" style="white-space:nowrap;overflow:hidden;text-overflow:ellipsis;display:block;max-width:200px"><i>%s</i></span>',
      x, x)
  ) %>%
  gt_plt_bar_pct(column = Prevalence, scaled = TRUE,
                 labels = TRUE, fill = oi[7], background = "#e8e8e8") %>%
  text_transform(
    locations = cells_body(columns = Assemblies),
    fn = function(x) mapply(
      function(v, tip) sprintf('<span title="%s">%s</span>', gsub('"', '&quot;', tip), v),
      x, species_asm_tips[seq_along(x)], SIMPLIFY = TRUE)
  ) %>%
  text_transform(
    locations = cells_body(columns = Prevalence),
    fn = function(x) mapply(
      function(v, tip) sprintf('<span title="%s" style="display:block">%s</span>', gsub('"', '&quot;', tip), v),
      x, species_asm_tips[seq_along(x)], SIMPLIFY = TRUE)
  ) %>%
  cols_label(Prevalence = "% assemblies") %>%
  cols_align(align = "right", columns = Assemblies) %>%
  cols_width(Species ~ px(200), Assemblies ~ px(65), Prevalence ~ px(150)) %>%
  tab_header(title = md("**Top Species**")) %>%
  tab_style(
    style = css(vertical_align = "middle"),
    locations = cells_body()
  ) %>%
  gt_theme()

# Render side by side using a two-column HTML layout
cat('<div style="display: flex; gap: 12px; align-items: flex-start; flex-wrap: wrap;">\n')
cat('<div style="flex: 1; min-width: 300px;">\n')
cat(as.character(as_raw_html(genus_gt)))
cat('\n</div>\n')
cat('<div style="flex: 1; min-width: 300px;">\n')
cat(as.character(as_raw_html(species_gt)))
cat('\n</div>\n')
cat('</div>\n')
```


---

<div class="note-box">
Generated by **final_finalizer** multi-assembly comparison &mdash; `r Sys.Date()`
</div>
