---
title: "Multi-Assembly Comparison Report"
params:
  comparison_tsv: "__COMPARISON_TSV__"
  completeness_tsv: "__COMPLETENESS_TSV__"
  ref_lengths_tsv: "__REF_LENGTHS_TSV__"
  per_assembly_summary_tsvs: "__PER_ASM_TSVS__"
  assembly_names: "__ASM_NAMES__"
  per_assembly_macro_blocks_tsvs: "__PER_ASM_MACRO_TSVS__"
  pairwise_macro_blocks_tsvs: "__PAIRWISE_MACRO_TSVS__"
  pairwise_names: "__PAIRWISE_NAMES__"
  reference_name: "__REFNAME__"
  synteny_mode: "__SYNTENY_MODE__"
  chr_like_minlen: "__CHRLIKE__"
  out_prefix: "__OUTPREFIX__"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
---

```{css, echo=FALSE}
body {
  font-family: "Helvetica Neue", Helvetica, "Liberation Sans", Arial, sans-serif;
  font-size: 13px;
  max-width: 1300px;
  margin: auto;
}
h1.title { font-size: 24px; margin-bottom: 4px; }
h1 { font-size: 20px; border-bottom: 2px solid #2c3e50; padding-bottom: 4px; }
h2 { font-size: 16px; }
.gt_table { font-size: 11px !important; }
.gt_col_heading { font-size: 10px !important; font-weight: bold !important; }
.widget-container { margin: 16px 0; }
.note-box {
  background: #f8f9fa; border-left: 4px solid #6c757d;
  padding: 10px 14px; margin: 12px 0; font-size: 12px; color: #495057;
}
.section-intro { font-size: 13px; color: #555; margin-bottom: 12px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, dpi = 150)

if (!requireNamespace("pacman", quietly = TRUE))
  install.packages("pacman", repos = "https://cloud.r-project.org")
library(pacman)
pacman::p_load(readr, dplyr, stringr, tidyr, tibble,
               ggplot2, ggnewscale, patchwork, grid,
               ggiraph, htmlwidgets, ggrepel,
               ggokabeito, colorspace,
               gt, gtExtras, scales,
               showtext, sysfonts, knitr)

# ---------------------------------------------------------------------------
# Font setup (shared across all plots)
# ---------------------------------------------------------------------------
.resolve_font <- function(family) {
  slug <- tolower(gsub(" ", "", family))
  path <- systemfonts::match_fonts(family)$path
  if (nzchar(path) && grepl(slug, tolower(gsub("[- ]", "", basename(path)))))
    return(path)
  NULL
}
base_family <- "sans"
for (.fam in c("Helvetica Neue", "Helvetica", "Liberation Sans")) {
  .r <- .resolve_font(.fam)
  if (!is.null(.r)) {
    sysfonts::font_add(.fam,
      regular    = .r,
      bold       = systemfonts::match_fonts(.fam, weight = "bold")$path,
      italic     = systemfonts::match_fonts(.fam, italic = TRUE)$path,
      bolditalic = systemfonts::match_fonts(.fam, italic = TRUE, weight = "bold")$path
    )
    base_family <- .fam
    break
  }
}
showtext_auto()

# ---------------------------------------------------------------------------
# Shared theming
# ---------------------------------------------------------------------------
base_font_pt <- 8
axis_title_margin_pt <- 4
axis_text_margin_pt <- 4

axis_theme <- theme(
  axis.title.x = element_text(size = base_font_pt, family = base_family,
                               margin = margin(t = axis_title_margin_pt)),
  axis.title.y = element_text(size = base_font_pt, family = base_family,
                               margin = margin(r = axis_title_margin_pt)),
  axis.text.x = element_text(size = base_font_pt, family = base_family,
                               margin = margin(t = axis_text_margin_pt)),
  axis.text.y = element_text(size = base_font_pt, family = base_family)
)

apply_legend_theme <- function(p, text_pt = base_font_pt, key_pt = base_font_pt,
                               tight = TRUE) {
  base <- theme(
    legend.title = element_text(size = text_pt, family = base_family),
    legend.text  = element_text(size = text_pt, family = base_family),
    legend.key.height = grid::unit(key_pt, "pt"),
    legend.key.width  = grid::unit(key_pt, "pt"),
    legend.direction = "horizontal",
    legend.box = "horizontal"
  )
  if (tight) {
    base <- base + theme(
      legend.box.margin = margin(0, 0, 0, 0),
      legend.margin = margin(0, 0, 0, 0),
      legend.spacing.y = grid::unit(0, "pt"),
      legend.spacing.x = grid::unit(2, "pt")
    )
  }
  p + base
}

# ---------------------------------------------------------------------------
# Okabe-Ito palette and classification colors
# ---------------------------------------------------------------------------
oi <- palette_okabe_ito()

classification_colors <- c(
  "chrom_assigned"     = oi[5],
  "chrom_unassigned"   = oi[2],
  "organelle_complete" = oi[3],
  "organelle_debris"   = oi[4],
  "rDNA"               = oi[1],
  "contaminant"        = oi[7],
  "chrom_debris"       = lighten(oi[5], amount = 0.5),
  "debris"             = oi[9],
  "unclassified"       = oi[8]
)

classification_labels <- c(
  "chrom_assigned"     = "Chromosome (assigned)",
  "chrom_unassigned"   = "Chromosome (unassigned)",
  "organelle_complete" = "Organelle (complete)",
  "organelle_debris"   = "Organelle (debris)",
  "rDNA"               = "rDNA",
  "contaminant"        = "Contaminant",
  "chrom_debris"       = "Chromosome debris",
  "debris"             = "Debris",
  "unclassified"       = "Unclassified"
)

classification_levels <- c(
  "chrom_assigned", "chrom_unassigned", "organelle_complete", "organelle_debris",
  "rDNA", "contaminant", "chrom_debris", "debris", "unclassified"
)

# ---------------------------------------------------------------------------
# Helpers
# ---------------------------------------------------------------------------
save_pdf <- function(path, plot, width = 7.2, height = 5, dpi = 300) {
  dev <- if (capabilities("cairo")) cairo_pdf else "pdf"
  invisible(ggsave(path, plot, width = width, height = height, dpi = dpi, device = dev))
}

girafe_styled <- function(ggobj, width_svg = 10, height_svg = 6.9) {
  g <- ggiraph::girafe(ggobj = ggobj, width_svg = width_svg, height_svg = height_svg)
  ggiraph::girafe_options(g,
    ggiraph::opts_tooltip(css = paste0(
      "font-family: ", base_family, "; font-size: 9pt; ",
      "background: white; padding: 6px; border: 1px solid #666;"
    )),
    ggiraph::opts_hover(css = "opacity: 0.8;"),
    ggiraph::opts_toolbar(saveaspng = FALSE)
  )
}

gt_theme <- function(gt_obj) {
  gt_obj %>%
    tab_options(
      table.font.size = px(11),
      column_labels.font.weight = "bold",
      column_labels.border.top.style = "none",
      column_labels.border.bottom.width = px(2),
      column_labels.border.bottom.color = "#333",
      table.border.top.style = "none",
      table.border.bottom.style = "none",
      table_body.border.bottom.color = "#999",
      data_row.padding = px(4),
      heading.border.bottom.style = "none"
    ) %>%
    opt_row_striping(row_striping = FALSE)
}

fmt_mb <- function(x) sprintf("%.1f", x / 1e6)
fmt_pct <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.1f%%", x * 100))

# Assembly color palette (one color per assembly)
asm_pal <- c(oi[5], oi[1], oi[3], oi[6], oi[7], oi[4], oi[2], oi[8], oi[9])

# ---------------------------------------------------------------------------
# Load data
# ---------------------------------------------------------------------------
comp_df <- read_tsv(params$comparison_tsv, show_col_types = FALSE)
compl_df <- read_tsv(params$completeness_tsv, show_col_types = FALSE)
ref_lengths <- read_tsv(params$ref_lengths_tsv, show_col_types = FALSE,
                        col_types = cols(subgenome = col_character()))

n_asm <- nrow(comp_df)
asm_names <- comp_df$assembly
asm_colors <- setNames(rep_len(asm_pal, n_asm), asm_names)

# Load per-assembly summary TSVs
asm_tsv_paths <- strsplit(params$per_assembly_summary_tsvs, ";")[[1]]
asm_name_list <- strsplit(params$assembly_names, ";")[[1]]

per_asm_summaries <- list()
for (i in seq_along(asm_tsv_paths)) {
  p <- asm_tsv_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(assigned_subgenome = col_character()))
    df$assembly <- asm_name_list[i]
    per_asm_summaries[[i]] <- df
  }
}
all_summary_df <- if (length(per_asm_summaries) > 0) bind_rows(per_asm_summaries) else NULL
has_per_asm <- !is.null(all_summary_df) && nrow(all_summary_df) > 0

has_any_chrom <- any(comp_df$n_chrom_assigned > 0, na.rm = TRUE)
has_any_contam <- any(comp_df$n_contaminants > 0, na.rm = TRUE)
has_any_depth <- any(!is.na(comp_df$mean_chrom_depth) & comp_df$mean_chrom_depth > 0)

chr_like_minlen <- as.numeric(params$chr_like_minlen)

# PDF output paths
pdf_classbar <- paste0(params$out_prefix, ".classification_comparison.pdf")
pdf_heatmap  <- paste0(params$out_prefix, ".chromosome_completeness.pdf")
pdf_synteny  <- paste0(params$out_prefix, ".chromosome_synteny.pdf")

# ---------------------------------------------------------------------------
# Load per-assembly macro_blocks and pairwise macro_blocks for synteny plot
# ---------------------------------------------------------------------------
per_asm_macro_paths <- if (nzchar(params$per_assembly_macro_blocks_tsvs))
  strsplit(params$per_assembly_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_macro_paths <- if (nzchar(params$pairwise_macro_blocks_tsvs))
  strsplit(params$pairwise_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_pair_names <- if (nzchar(params$pairwise_names))
  strsplit(params$pairwise_names, ";")[[1]] else character(0)

# Load per-assembly macro_blocks (ref→asm) and tag with assembly name
per_asm_macro_list <- list()
for (i in seq_along(per_asm_macro_paths)) {
  p <- per_asm_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      per_asm_macro_list[[i]] <- df
    }
  }
}
per_asm_macro_df <- if (length(per_asm_macro_list) > 0) bind_rows(per_asm_macro_list) else NULL

# Load pairwise macro_blocks (asm→asm)
pairwise_macro_list <- list()
for (i in seq_along(pairwise_macro_paths)) {
  p <- pairwise_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0 && i <= length(pairwise_pair_names)) {
      df$pair_name <- pairwise_pair_names[i]
      # Extract left and right assembly names from pair name
      # Split only on first occurrence of "_vs_" to handle assembly names
      # that might contain "_vs_" (unlikely but defensive)
      parts <- strsplit(pairwise_pair_names[i], "_vs_", fixed = TRUE)[[1]]
      if (length(parts) >= 2) {
        df$left_assembly <- parts[1]
        df$right_assembly <- paste(parts[-1], collapse = "_vs_")
      } else {
        warning(sprintf("Cannot parse pairwise pair name: %s", pairwise_pair_names[i]))
        next
      }
      pairwise_macro_list[[i]] <- df
    }
  }
}
pairwise_macro_df <- if (length(pairwise_macro_list) > 0) bind_rows(pairwise_macro_list) else NULL

has_synteny_data <- !is.null(per_asm_macro_df) && nrow(per_asm_macro_df) > 0 &&
                    params$synteny_mode == "nucleotide"
```

---

<p class="section-intro">
**Reference**: `r if (nchar(params$reference_name) > 0) params$reference_name else "\u2014"` &emsp;
**Synteny mode**: `r params$synteny_mode` &emsp;
**Assemblies**: `r n_asm` &emsp;
**Date**: `r Sys.Date()`
</p>

# 1. Assembly Overview {.tabset}

## Summary Table

```{r overview_table}
tbl <- comp_df %>%
  transmute(
    Assembly = assembly,
    Contigs  = total_contigs,
    `Size (Mb)` = round(total_bp / 1e6, 1),
    N50 = round(n50 / 1e6, 2),
    L50 = l50,
    `Largest (Mb)` = round(largest_contig / 1e6, 1),
    `Chr assigned` = n_chrom_assigned,
    `Chr unassigned` = n_chrom_unassigned,
    `Full-length` = n_full_length,
    `Both telo` = n_both_telomeres,
    `Mean ref cov` = ifelse(is.na(mean_ref_coverage), "\u2014",
                            sprintf("%.1f%%", mean_ref_coverage * 100)),
    Chimeric = n_chimeric,
    `Mean identity` = ifelse(is.na(mean_identity), "\u2014",
                             sprintf("%.2f%%", mean_identity * 100)),
    Contaminants = n_contaminants,
    rDNA = n_rdna,
    chrC = ifelse(chrC_found == "yes", "\u2713", "\u2014"),
    chrM = ifelse(chrM_found == "yes", "\u2713", "\u2014")
  )

tbl %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Assembly Comparison Summary") %>%
  cols_align(align = "right", columns = where(is.numeric)) %>%
  tab_spanner(label = "Contiguity", columns = c(Contigs, `Size (Mb)`, N50, L50, `Largest (Mb)`)) %>%
  tab_spanner(label = "Chromosomes", columns = c(`Chr assigned`, `Chr unassigned`, `Full-length`, `Both telo`, `Mean ref cov`, Chimeric, `Mean identity`)) %>%
  tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrC, chrM))
```

## Classification Distribution

```{r classification_bar, fig.width=10, fig.height=max(3, n_asm * 0.8 + 1.5)}
if (has_per_asm) {
  clf_data <- all_summary_df %>%
    group_by(assembly, classification) %>%
    summarise(total_bp = sum(length, na.rm = TRUE), n_contigs = n(), .groups = "drop") %>%
    mutate(
      total_mb = total_bp / 1e6,
      classification = factor(classification, levels = rev(classification_levels)),
      assembly = factor(assembly, levels = rev(asm_names))
    )

  p_clf <- ggplot(clf_data, aes(x = assembly, y = total_mb, fill = classification)) +
    geom_col_interactive(
      aes(tooltip = paste0(assembly, "\n",
                           classification_labels[as.character(classification)], "\n",
                           sprintf("%.1f Mb (%d contigs)", total_mb, n_contigs))),
      width = 0.7
    ) +
    scale_fill_manual(values = classification_colors, labels = classification_labels,
                      drop = FALSE, name = NULL) +
    coord_flip() +
    labs(x = NULL, y = "Assembly size (Mb)") +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(legend.position = "bottom",
          panel.grid.major.y = element_blank())
  p_clf <- apply_legend_theme(p_clf)

  save_pdf(pdf_classbar, p_clf,
           width = 7.2, height = max(3, n_asm * 0.8 + 1.5))
  girafe_styled(ggobj = p_clf,
                width_svg = 10,
                height_svg = max(3, n_asm * 0.8 + 1.5))
} else {
  cat("Per-assembly contig summary data not available.")
}
```

## Size Comparison

```{r size_comparison, fig.width=10, fig.height=max(3, n_asm * 0.6 + 1.5)}
size_data <- comp_df %>%
  transmute(
    assembly = factor(assembly, levels = rev(asm_names)),
    `Chrom assigned`  = chrom_assigned_bp / 1e6,
    `Organelle`       = organelle_bp / 1e6,
    `rDNA`            = rdna_bp / 1e6,
    `Contaminant`     = contaminant_bp / 1e6,
    `Debris`          = debris_bp / 1e6,
    `Unclassified`    = unclassified_bp / 1e6
  ) %>%
  pivot_longer(-assembly, names_to = "category", values_to = "mb") %>%
  mutate(category = factor(category,
    levels = c("Chrom assigned", "Organelle", "rDNA", "Contaminant", "Debris", "Unclassified")))

cat_colors <- c(
  "Chrom assigned" = oi[5], "Organelle" = oi[3], "rDNA" = oi[1],
  "Contaminant" = oi[7], "Debris" = oi[9], "Unclassified" = oi[8]
)

p_size <- ggplot(size_data, aes(x = assembly, y = mb, fill = category)) +
  geom_col_interactive(
    aes(tooltip = paste0(assembly, "\n", category, ": ", sprintf("%.1f Mb", mb))),
    position = "dodge", width = 0.7
  ) +
  scale_fill_manual(values = cat_colors, name = NULL) +
  coord_flip() +
  labs(x = NULL, y = "Size (Mb)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(legend.position = "bottom", panel.grid.major.y = element_blank())
p_size <- apply_legend_theme(p_size)

girafe_styled(ggobj = p_size,
              width_svg = 10,
              height_svg = max(3, n_asm * 0.6 + 1.5))
```


# 2. Chromosome Completeness {.tabset}

## Heatmap

```{r completeness_heatmap, fig.width=10, fig.height=max(4, nrow(ref_lengths) * 0.3 + 2)}
if (has_any_chrom) {
  # Sort reference chromosomes by length
  ref_order <- ref_lengths %>%
    arrange(desc(ref_len)) %>%
    pull(ref_id)

  hm_data <- compl_df %>%
    mutate(
      ref_id = factor(ref_id, levels = rev(ref_order)),
      assembly = factor(assembly, levels = asm_names),
      best_ref_coverage = as.numeric(best_ref_coverage),
      coverage_pct = ifelse(is.na(best_ref_coverage), 0, best_ref_coverage * 100)
    )

  # Telomere/full-length status for overlay
  hm_data <- hm_data %>%
    mutate(
      status_label = case_when(
        is_full_length == "yes" & has_both_telomeres == "yes" ~ "\u2605",  # star
        is_full_length == "yes" ~ "\u25CF",  # filled circle
        n_contigs > 0 ~ "\u25CB",  # open circle
        TRUE ~ ""
      ),
      tooltip = paste0(
        ref_id, " \u2014 ", assembly, "\n",
        "Contigs: ", n_contigs, "\n",
        "Assigned bp: ", scales::comma(total_assigned_bp), "\n",
        "Best ref coverage: ", fmt_pct(best_ref_coverage), "\n",
        "Full-length: ", is_full_length, "\n",
        "Both telomeres: ", has_both_telomeres
      )
    )

  p_hm <- ggplot(hm_data, aes(x = assembly, y = ref_id, fill = coverage_pct)) +
    geom_tile_interactive(aes(tooltip = tooltip), color = "white", linewidth = 0.5) +
    geom_text(aes(label = status_label), size = 2.5, color = "white", fontface = "bold") +
    scale_fill_gradient2(
      low = "#f0f0f0", mid = "#74add1", high = "#1a9850",
      midpoint = 50, limits = c(0, 100),
      name = "Ref coverage (%)",
      na.value = "#f0f0f0"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = base_font_pt),
      legend.position = "bottom",
      panel.grid = element_blank()
    )
  p_hm <- apply_legend_theme(p_hm)

  save_pdf(pdf_heatmap, p_hm,
           width = max(4, n_asm * 0.8 + 2),
           height = max(4, length(ref_order) * 0.3 + 2))
  girafe_styled(ggobj = p_hm,
                width_svg = max(6, n_asm * 0.8 + 2),
                height_svg = max(4, length(ref_order) * 0.3 + 2))
} else {
  cat("No chromosome-assigned contigs found in any assembly.")
}
```

<div class="note-box">
**Legend**: &#9733; = full-length with both telomeres, &#9679; = full-length, &#9675; = assigned (not full-length), blank = no contigs assigned.
</div>

## Completeness Table

```{r completeness_table}
if (has_any_chrom) {
  tbl_data <- compl_df %>%
    mutate(best_ref_coverage = as.numeric(best_ref_coverage)) %>%
    select(ref_id, assembly, n_contigs, total_assigned_bp,
           best_ref_coverage, is_full_length, has_both_telomeres) %>%
    mutate(
      `Assigned (Mb)` = round(total_assigned_bp / 1e6, 1),
      `Ref cov (%)` = ifelse(is.na(best_ref_coverage), "\u2014",
                             sprintf("%.1f", best_ref_coverage * 100)),
      `Full-length` = ifelse(is_full_length == "yes", "\u2713", ""),
      `Both telo` = ifelse(has_both_telomeres == "yes", "\u2713", "")
    ) %>%
    select(ref_id, assembly, n_contigs, `Assigned (Mb)`, `Ref cov (%)`,
           `Full-length`, `Both telo`)

  tbl_data %>%
    gt(groupname_col = "ref_id") %>%
    gt_theme() %>%
    tab_header(title = "Chromosome Completeness by Reference") %>%
    cols_label(
      assembly = "Assembly", n_contigs = "Contigs",
      `Assigned (Mb)` = "Assigned (Mb)",
      `Ref cov (%)` = "Ref cov (%)",
      `Full-length` = "Full-length",
      `Both telo` = "Both telo"
    ) %>%
    cols_align(align = "right", columns = c(n_contigs, `Assigned (Mb)`, `Ref cov (%)`)) %>%
    cols_align(align = "center", columns = c(`Full-length`, `Both telo`))
} else {
  cat("No chromosome-assigned contigs found in any assembly.")
}
```


`r if (has_any_chrom && has_per_asm) "# 3. Quality Comparison {.tabset}" else ""`

```{r quality_section, eval=(has_any_chrom && has_per_asm)}
chrom_df <- all_summary_df %>%
  filter(classification == "chrom_assigned")

has_identity <- "seq_identity_vs_ref" %in% names(chrom_df) &&
                !all(is.na(chrom_df$seq_identity_vs_ref))
has_collin   <- "collinearity_score" %in% names(chrom_df) &&
                !all(is.na(chrom_df$collinearity_score))
has_gc_dev   <- "gc_deviation" %in% names(chrom_df) &&
                !all(is.na(chrom_df$gc_deviation))
```

```{r quality_identity, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), results='asis'}
cat("## Sequence Identity\n\n")
```

```{r quality_identity_plot, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
p_ident <- ggplot(chrom_df, aes(x = assembly, y = seq_identity_vs_ref * 100,
                                fill = assembly)) +
  geom_violin(alpha = 0.6, width = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Sequence identity vs reference (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_ident, height_svg = max(3, n_asm * 0.8 + 1))
```

```{r quality_collinearity, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), results='asis'}
cat("## Collinearity\n\n")
```

```{r quality_collinearity_plot, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
p_collin <- ggplot(chrom_df, aes(x = assembly, y = collinearity_score * 100,
                                  fill = assembly)) +
  geom_violin(alpha = 0.6, width = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Collinearity (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_collin, height_svg = max(3, n_asm * 0.8 + 1))
```

```{r quality_gc, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), results='asis'}
cat("## GC Deviation\n\n")
```

```{r quality_gc_plot, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
p_gc <- ggplot(chrom_df, aes(x = assembly, y = abs(gc_deviation), fill = assembly)) +
  geom_violin(alpha = 0.6, width = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "|GC deviation| (std devs from reference)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_gc, height_svg = max(3, n_asm * 0.8 + 1))
```


`r if (has_any_contam) "# 4. Contamination Comparison" else ""`

```{r contamination, eval=has_any_contam, fig.width=10, fig.height=max(3, n_asm * 0.6 + 1.5)}
contam_data <- comp_df %>%
  filter(n_contaminants > 0) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_names)))

p_contam <- ggplot(contam_data, aes(x = assembly, fill = assembly)) +
  geom_col_interactive(
    aes(y = contaminant_bp / 1e6,
        tooltip = paste0(assembly, "\n",
                         n_contaminants, " contaminants\n",
                         sprintf("%.2f Mb", contaminant_bp / 1e6), "\n",
                         n_contaminant_species, " species")),
    width = 0.6
  ) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Contaminant sequence (Mb)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_contam,
              height_svg = max(3, n_asm * 0.6 + 1.5))
```

```{r contamination_table, eval=has_any_contam}
comp_df %>%
  filter(n_contaminants > 0) %>%
  transmute(
    Assembly = assembly,
    `N contaminants` = n_contaminants,
    `Contam bp (Mb)` = round(contaminant_bp / 1e6, 2),
    `Unique species` = n_contaminant_species
  ) %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Contamination Summary") %>%
  cols_align(align = "right", columns = where(is.numeric))
```


`r if (has_any_depth) "# 5. Depth Comparison" else ""`

```{r depth_section, eval=(has_any_depth && has_per_asm), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
depth_df <- all_summary_df %>%
  filter(classification == "chrom_assigned", !is.na(depth_mean))

if (nrow(depth_df) > 0) {
  p_depth <- ggplot(depth_df, aes(x = assembly, y = depth_mean, fill = assembly)) +
    geom_violin(alpha = 0.6, width = 0.7) +
    geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
    scale_fill_manual(values = asm_colors, guide = "none") +
    coord_flip() +
    labs(x = NULL, y = "Mean read depth (chromosome contigs)") +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(panel.grid.major.y = element_blank())

  girafe_styled(ggobj = p_depth, height_svg = max(3, n_asm * 0.8 + 1))
} else {
  cat("No depth data available for chromosome-assigned contigs.")
}
```

`r if (has_synteny_data) "# 6. Chromosome Synteny {.tabset}" else ""`

```{r synteny_setup, eval=has_synteny_data, include=FALSE}
# =========================================================================
# Section 6: Chromosome Synteny Riparian Plot
#
# Per-subgenome grouped x-axis layout (no faceting).
# Each subgenome gets its own horizontal group:
#   [Ref.A | asm1 | asm2]  ||  [Ref.P | asm3]  ||  [Ref.T | asm4]
# Within each group: ribbons connect adjacent columns.
# Chromosome rows (chr1..chr_n) are shared across all groups.
# =========================================================================

# --- Assembly contig lookup (incl. original_name for name mapping) ---
asm_contig_lookup <- list()
for (i in seq_along(asm_name_list)) {
  nm <- asm_name_list[i]
  df <- per_asm_summaries[[i]]
  if (!is.null(df)) {
    lookup <- df %>%
      filter(classification == "chrom_assigned") %>%
      select(contig, original_name, assigned_ref_id, reversed, length) %>%
      mutate(assembly = nm, reversed = (reversed == "yes"))
    asm_contig_lookup[[nm]] <- lookup
  }
}
asm_lookup_df <- bind_rows(asm_contig_lookup)

# --- Fix per-assembly macro_blocks: map original_name -> new_name ---
if (!is.null(per_asm_macro_df) && nrow(asm_lookup_df) > 0) {
  name_map <- asm_lookup_df %>%
    distinct(assembly, original_name, .keep_all = TRUE) %>%
    select(assembly, original_name, new_name = contig)
  per_asm_macro_df <- per_asm_macro_df %>%
    left_join(name_map, by = c("assembly", "contig" = "original_name")) %>%
    mutate(contig_new = coalesce(new_name, contig))
} else if (!is.null(per_asm_macro_df)) {
  per_asm_macro_df <- per_asm_macro_df %>% mutate(contig_new = contig)
}

# --- Nuclear chromosomes sorted by (subgenome, chrom_num) ---
has_named_subgenomes <- any(
  !is.na(ref_lengths$subgenome) & ref_lengths$subgenome != "NA"
)

if (has_named_subgenomes) {
  nuc_ref <- ref_lengths %>%
    filter(!is.na(subgenome) & subgenome != "NA") %>%
    mutate(chrom_num = as.numeric(str_extract(chrom_id, "\\d+"))) %>%
    arrange(subgenome, chrom_num)
} else {
  macro_rids <- unique(per_asm_macro_df$ref_id)
  nuc_ref <- ref_lengths %>%
    filter(ref_id %in% macro_rids) %>%
    mutate(
      subgenome = "all",
      chrom_num = as.numeric(str_extract(ref_id, "\\d+"))
    ) %>%
    arrange(chrom_num)
}

subgenome_order <- unique(nuc_ref$subgenome)
n_subgenomes <- length(subgenome_order)
ref_chr_order <- nuc_ref$ref_id

# --- Determine assemblies per subgenome ---
# For each subgenome, find which assemblies have chromosomes assigned to it
sg_assemblies <- list()
for (sg in subgenome_order) {
  sg_ref_ids <- nuc_ref %>% filter(subgenome == sg) %>% pull(ref_id)
  asms_with_sg <- asm_lookup_df %>%
    filter(assigned_ref_id %in% sg_ref_ids) %>%
    pull(assembly) %>% unique()
  # Maintain original assembly order
  sg_assemblies[[sg]] <- asm_name_list[asm_name_list %in% asms_with_sg]
}

# --- X-axis layout: per-subgenome groups with wider inter-group gaps ---
col_spacing <- 2    # within group
group_gap   <- 4    # between groups

# Build column positions and metadata
col_info <- tibble(
  column = character(), x_center = numeric(),
  subgenome = character(), is_ref = logical()
)
all_col_centers <- numeric()   # named vector for x-axis breaks
all_col_labels  <- character() # for x-axis labels
gap_zones <- list()            # gap zone per adjacent-column pair within groups
gap_meta  <- list()            # metadata: left_col, right_col, subgenome

x_cursor <- 1
for (sg_i in seq_along(subgenome_order)) {
  sg <- subgenome_order[sg_i]
  sg_asms <- sg_assemblies[[sg]]

  # Reference column for this subgenome
  ref_col_name <- if (n_subgenomes > 1) paste0("Ref.", sg) else "Reference"
  ref_x <- x_cursor
  col_info <- bind_rows(col_info, tibble(
    column = ref_col_name, x_center = ref_x, subgenome = sg, is_ref = TRUE
  ))
  all_col_centers <- c(all_col_centers, setNames(ref_x, ref_col_name))
  all_col_labels  <- c(all_col_labels, ref_col_name)

  prev_x <- ref_x
  prev_col <- ref_col_name

  # Assembly columns for this subgenome
  for (asm in sg_asms) {
    asm_x <- prev_x + col_spacing
    # Column label: use assembly name (may appear in multiple groups)
    asm_col_name <- if (n_subgenomes > 1) paste0(asm, ".", sg) else asm
    col_info <- bind_rows(col_info, tibble(
      column = asm_col_name, x_center = asm_x, subgenome = sg, is_ref = FALSE
    ))
    all_col_centers <- c(all_col_centers, setNames(asm_x, asm_col_name))
    all_col_labels  <- c(all_col_labels, asm)

    # Gap zone between prev_col and this column
    gap_zones[[length(gap_zones) + 1]] <- c(prev_x + 0.2, asm_x - 0.2)
    gap_meta[[length(gap_meta) + 1]] <- list(
      left_col = prev_col, right_col = asm_col_name,
      left_is_ref = (prev_col == ref_col_name), subgenome = sg,
      left_asm = if (prev_col == ref_col_name) NA_character_ else {
        # Extract assembly name from compound col name
        sub(paste0("\\.", sg, "$"), "", prev_col)
      },
      right_asm = asm
    )
    prev_x <- asm_x
    prev_col <- asm_col_name
  }

  # Advance cursor for next group
  if (sg_i < n_subgenomes) {
    x_cursor <- prev_x + group_gap
  }
}

n_total_cols <- nrow(col_info)

# --- Y-axis layout: shared chromosome rows across all groups ---
# Row heights based on max ref chromosome length per chrom_num
chrom_nums <- sort(unique(nuc_ref$chrom_num))
max_len_per_row <- nuc_ref %>%
  group_by(chrom_num) %>%
  summarise(max_len = max(ref_len), .groups = "drop") %>%
  arrange(chrom_num)

total_max_len <- sum(max_len_per_row$max_len)
y_scale <- total_max_len / 15  # total maps to ~15 y-units
row_gap <- 0.3

# Compute y_top for each chrom_num row
row_layout <- max_len_per_row %>%
  mutate(
    row_height = max_len / y_scale,
    cum_height = cumsum(row_height) - row_height,
    gap_above = (row_number() - 1) * row_gap,
    y_top = cum_height + gap_above
  )

# Join row y_top back to chr_layout
chr_layout <- nuc_ref %>%
  left_join(row_layout %>% select(chrom_num, y_top_row = y_top),
            by = "chrom_num") %>%
  mutate(
    pill_height = ref_len / y_scale,
    y_top = y_top_row
  )

# Pill linewidth
pill_lw <- 3.5

# --- Build pill data for all columns ---
pills_list <- list()

for (sg in subgenome_order) {
  sg_chr <- chr_layout %>% filter(subgenome == sg)
  sg_asms <- sg_assemblies[[sg]]
  ref_col_name <- if (n_subgenomes > 1) paste0("Ref.", sg) else "Reference"
  ref_x <- col_info %>% filter(column == ref_col_name) %>% pull(x_center)

  # Reference pills for this subgenome
  pills_list[[ref_col_name]] <- sg_chr %>%
    transmute(
      column = ref_col_name,
      ref_id, chrom_id, subgenome,
      contig = ref_id, contig_len = ref_len, reversed = FALSE,
      x_center = ref_x,
      y_top, y_bot = y_top + pill_height,
      pill_height,
      tooltip = sprintf("Reference\n%s\n%.2f Mb", ref_id, ref_len / 1e6)
    )

  # Assembly pills for this subgenome
  for (asm in sg_asms) {
    asm_col_name <- if (n_subgenomes > 1) paste0(asm, ".", sg) else asm
    asm_x <- col_info %>% filter(column == asm_col_name) %>% pull(x_center)

    lk <- asm_lookup_df %>%
      filter(assembly == asm, assigned_ref_id %in% sg_chr$ref_id) %>%
      group_by(assigned_ref_id) %>%
      arrange(desc(length)) %>%
      slice(1) %>%
      ungroup()

    if (nrow(lk) == 0) next

    asm_pills <- sg_chr %>%
      inner_join(
        lk %>% select(assigned_ref_id, asm_contig = contig,
                       asm_len = length, asm_reversed = reversed),
        by = c("ref_id" = "assigned_ref_id")
      ) %>%
      transmute(
        column = asm_col_name,
        ref_id, chrom_id, subgenome,
        contig = asm_contig, contig_len = asm_len,
        reversed = asm_reversed,
        x_center = asm_x,
        y_top,
        pill_height = asm_len / y_scale,
        y_bot = y_top + asm_len / y_scale,
        tooltip = sprintf("%s\n%s\n%.2f Mb", asm, asm_contig, asm_len / 1e6)
      )
    pills_list[[asm_col_name]] <- asm_pills
  }
}

all_pills <- bind_rows(pills_list)
all_pills$column <- factor(all_pills$column, levels = col_info$column)

# --- Chromosome color palette (by chrom_id, shared across subgenomes) ---
unique_chroms <- unique(nuc_ref$chrom_id)
unique_chroms <- unique_chroms[order(as.numeric(str_extract(unique_chroms, "\\d+")))]
n_unique_chroms <- length(unique_chroms)
if (n_unique_chroms <= 9) {
  chrom_pal <- setNames(oi[seq_len(n_unique_chroms)], unique_chroms)
} else {
  chrom_pal <- setNames(
    colorRampPalette(oi[1:9])(n_unique_chroms),
    unique_chroms
  )
}

# --- Helpers ---
pos_to_y <- function(pos, seq_len, pill_y_top, pill_h) {
  pill_y_top + (pos / seq_len) * pill_h
}

make_ribbon <- function(gap_left, gap_right,
                        left_y_top, left_y_bot,
                        right_y_top, right_y_bot,
                        chrom_id_val, alpha_val, tooltip_text,
                        subgenome_val, n_pts = 20) {
  t <- seq(0, 1, length.out = n_pts)
  s <- 0.5 - 0.5 * cos(pi * t)
  x_fwd <- gap_left + t * (gap_right - gap_left)
  top_fwd <- left_y_top + s * (right_y_top - left_y_top)
  bot_fwd <- left_y_bot + s * (right_y_bot - left_y_bot)
  tibble(
    x = c(x_fwd, rev(x_fwd)),
    y = c(top_fwd, rev(bot_fwd)),
    chrom_id = chrom_id_val,
    alpha = alpha_val,
    tooltip = tooltip_text,
    subgenome = subgenome_val
  )
}

# --- Build ribbon polygons ---
ribbon_polys <- list()
ribbon_idx <- 0

pill_lookup <- all_pills %>%
  select(column, ref_id, chrom_id, contig, contig_len, reversed,
         x_center, y_top, y_bot, pill_height, subgenome)

max_ref_len <- max(nuc_ref$ref_len)
min_span <- max(50000, max_ref_len * 0.001)

for (gap_i in seq_along(gap_zones)) {
  gz <- gap_zones[[gap_i]]
  gm <- gap_meta[[gap_i]]
  left_col_name  <- gm$left_col
  right_col_name <- gm$right_col
  sg <- gm$subgenome

  sg_ref_ids <- nuc_ref %>% filter(subgenome == sg) %>% pull(ref_id)

  # --- Gather blocks for this gap ---
  if (gm$left_is_ref) {
    right_asm <- gm$right_asm
    if (!is.null(per_asm_macro_df)) {
      macro_src <- per_asm_macro_df %>%
        filter(assembly == right_asm, ref_id %in% sg_ref_ids)
    } else {
      macro_src <- NULL
    }
  } else {
    left_asm  <- gm$left_asm
    right_asm <- gm$right_asm
    if (!is.null(pairwise_macro_df)) {
      macro_src <- pairwise_macro_df %>%
        filter(left_assembly == left_asm, right_assembly == right_asm)
      if (n_subgenomes > 1) {
        macro_src <- macro_src %>% filter(subgenome == sg)
      }
    } else {
      macro_src <- NULL
    }
  }

  if (is.null(macro_src) || nrow(macro_src) == 0) next

  # Filter small blocks once for all blocks in this gap
  macro_src <- macro_src %>% filter(qspan_bp >= min_span)
  if (nrow(macro_src) == 0) next

  # --- Iterate per block (supports off-target blocks) ---
  for (b in seq_len(nrow(macro_src))) {
    blk <- macro_src[b, ]

    # Look up left and right pills by contig name
    if (gm$left_is_ref) {
      lp <- pill_lookup %>% filter(column == left_col_name, contig == blk$ref_id)
      rp <- pill_lookup %>% filter(column == right_col_name, contig == blk$contig_new)
    } else {
      lp <- pill_lookup %>% filter(column == left_col_name, contig == blk$ref_id)
      rp <- pill_lookup %>% filter(column == right_col_name, contig == blk$contig)
    }
    if (nrow(lp) == 0 || nrow(rp) == 0) next

    left_len  <- lp$contig_len[1]
    left_y    <- lp$y_top[1]
    left_h    <- lp$pill_height[1]
    right_len <- rp$contig_len[1]
    right_y   <- rp$y_top[1]
    right_h   <- rp$pill_height[1]

    # Color by left pill's chrom_id (source chromosome)
    cid <- as.character(lp$chrom_id[1])

    # Left side coordinates — never reversed
    # (ref coords are in reference frame; asm→asm coords are already oriented)
    l_start <- as.numeric(blk$ref_start)
    l_end   <- as.numeric(blk$ref_end)

    # Right side coordinates
    r_start <- as.numeric(blk$qstart)
    r_end   <- as.numeric(blk$qend)
    if (gm$left_is_ref && isTRUE(rp$reversed[1])) {
      # ref→asm: right assembly contig may be reversed relative to reference
      r_start_new <- right_len - r_end
      r_end_new   <- right_len - r_start
      r_start <- r_start_new
      r_end   <- r_end_new
    }
    # asm→asm: no reversal — pairwise alignment used already-oriented FASTAs

    # Clamp to valid ranges
    l_start <- max(0, l_start); l_end <- min(left_len, l_end)
    r_start <- max(0, r_start); r_end <- min(right_len, r_end)
    if (l_end <= l_start || r_end <= r_start) next

    # Map to y coordinates within each pill
    ly_top <- pos_to_y(l_start, left_len, left_y, left_h)
    ly_bot <- pos_to_y(l_end, left_len, left_y, left_h)
    ry_top <- pos_to_y(r_start, right_len, right_y, right_h)
    ry_bot <- pos_to_y(r_end, right_len, right_y, right_h)

    # Determine effective strand in the oriented frame
    strand_val <- blk$strand
    if (gm$left_is_ref && isTRUE(rp$reversed[1])) {
      # ref→asm: alignment was on unoriented query, but the pill shows the
      # oriented contig.  Reversing the contig flips the strand.
      strand_val <- if (strand_val == "+") "-" else "+"
    }
    # Inversions: strand = "-" -> swap top/bottom on right side
    if (strand_val == "-") {
      tmp <- ry_top; ry_top <- ry_bot; ry_bot <- tmp
    }

    ident_val <- as.numeric(blk$identity)
    alpha_ribbon <- 0.2 + 0.5 * min(1, ident_val)

    # Tooltip: show left → right contig/chromosome names
    left_key  <- lp$contig[1]
    right_key <- rp$contig[1]
    tooltip <- sprintf(
      "%s \u2192 %s\n%.2f Mb\nIdentity: %.1f%%\nStrand: %s",
      left_key, right_key,
      as.numeric(blk$union_bp) / 1e6,
      ident_val * 100, strand_val
    )

    ribbon_idx <- ribbon_idx + 1
    ribbon_polys[[ribbon_idx]] <- make_ribbon(
      gz[1], gz[2], ly_top, ly_bot, ry_top, ry_bot,
      chrom_id_val = cid, alpha_val = alpha_ribbon,
      tooltip_text = tooltip, subgenome_val = sg
    ) %>% mutate(group_id = ribbon_idx)
  }
}

has_ribbons <- length(ribbon_polys) > 0
if (has_ribbons) {
  all_ribbons <- bind_rows(ribbon_polys)
} else {
  all_ribbons <- NULL
}

# --- Plot sizing ---
n_chrom_rows <- length(chrom_nums)
syn_plot_h <- max(4, n_chrom_rows * 0.5 + 2)
syn_plot_w <- max(7.2, n_total_cols * 1.8 + 2)
```

```{r synteny_description, eval=has_synteny_data, results='asis'}
cat('<p class="section-intro">\n')
cat("Chromosome synteny between the reference and each assembly (nucleotide mode). ")
if (exists("n_subgenomes") && n_subgenomes > 1) {
  cat("Assemblies are grouped by reference subgenome. ")
}
cat("Ribbons connect syntenic macro-blocks between adjacent columns. ")
cat("Color indicates the reference chromosome; opacity reflects alignment identity. ")
cat("Crossed ribbons indicate inversions.\n")
cat("</p>\n")
```

## Riparian Plot

```{r synteny_riparian, eval=(has_synteny_data && exists("has_ribbons") && has_ribbons), fig.width=10, fig.height=10}
# --- Chromosome labels on far left of plot (one set, from first group) ---
first_sg <- subgenome_order[1]
first_ref_col <- if (n_subgenomes > 1) paste0("Ref.", first_sg) else "Reference"
first_ref_x <- col_info %>% filter(column == first_ref_col) %>% pull(x_center)
chr_label_df <- chr_layout %>%
  filter(subgenome == first_sg) %>%
  mutate(
    x = first_ref_x - 0.4,
    y_mid = y_top + pill_height / 2,
    label = chrom_id
  )

# --- Compose the plot (single flat plot, no faceting) ---
p_syn <- ggplot()

# Ribbons (behind pills)
if (has_ribbons) {
  p_syn <- p_syn +
    geom_polygon_interactive(
      data = all_ribbons,
      aes(x = x, y = y, group = group_id, fill = chrom_id,
          alpha = alpha, tooltip = tooltip),
      show.legend = FALSE
    ) +
    scale_fill_manual(values = chrom_pal, guide = "none") +
    scale_alpha_identity()
}

# Reference pills (darker)
ref_pill_data <- all_pills %>%
  semi_join(col_info %>% filter(is_ref), by = "column")
p_syn <- p_syn +
  geom_segment(
    data = ref_pill_data,
    aes(x = x_center, xend = x_center, y = y_top, yend = y_bot),
    linewidth = pill_lw, lineend = "round", color = "grey55"
  )

# Assembly pills (lighter, interactive)
asm_pill_data <- all_pills %>%
  semi_join(col_info %>% filter(!is_ref), by = "column")
if (nrow(asm_pill_data) > 0) {
  p_syn <- p_syn +
    geom_segment_interactive(
      data = asm_pill_data,
      aes(x = x_center, xend = x_center, y = y_top, yend = y_bot,
          tooltip = tooltip),
      linewidth = pill_lw, lineend = "round", color = "grey78"
    )
}

# Chromosome labels on left of each group
p_syn <- p_syn +
  geom_text(data = chr_label_df,
            aes(x = x, y = y_mid, label = label),
            size = 2, hjust = 1, family = base_family)

# Subgenome group labels (centered above each group, only if >1 subgenome)
if (n_subgenomes > 1) {
  sg_label_df <- col_info %>%
    group_by(subgenome) %>%
    summarise(x_mid = mean(x_center), .groups = "drop") %>%
    mutate(y = min(row_layout$y_top) - 0.6,
           label = paste0("Subgenome ", subgenome))

  p_syn <- p_syn +
    geom_text(data = sg_label_df,
              aes(x = x_mid, y = y, label = label),
              size = 2.8, fontface = "bold", family = base_family)
}

# Axis scales
p_syn <- p_syn +
  scale_x_continuous(
    breaks = as.numeric(all_col_centers),
    labels = all_col_labels,
    expand = expansion(add = c(1.5, 0.5))
  ) +
  scale_y_reverse(expand = expansion(mult = c(0.02, 0.05)))

# Theme (no faceting)
p_syn <- p_syn +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  theme(
    axis.text.x = element_text(size = base_font_pt, face = "bold",
                                family = base_family, angle = 45, hjust = 1),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = margin(5, 5, 5, 30)
  )

save_pdf(pdf_synteny, p_syn, width = 7.2, height = syn_plot_h)
girafe_styled(ggobj = p_syn, width_svg = syn_plot_w, height_svg = syn_plot_h)
```

```{r synteny_empty, eval=(has_synteny_data && !(exists("has_ribbons") && has_ribbons))}
cat("Synteny data available but no ribbon connections could be built. ",
    "This may indicate that assemblies have very different chromosome structures.")
```

```{r synteny_not_nucleotide, eval=(!has_synteny_data && params$synteny_mode != "nucleotide"), results='asis'}
# Don't show section header if not nucleotide mode — it's hidden by the conditional above
```

---

<div class="note-box">
Generated by **final_finalizer** multi-assembly comparison &mdash; `r Sys.Date()`
</div>
