---
title: "Multi-Assembly Comparison Report"
params:
  comparison_tsv: "__COMPARISON_TSV__"
  completeness_tsv: "__COMPLETENESS_TSV__"
  ref_lengths_tsv: "__REF_LENGTHS_TSV__"
  per_assembly_summary_tsvs: "__PER_ASM_TSVS__"
  assembly_names: "__ASM_NAMES__"
  per_assembly_macro_blocks_tsvs: "__PER_ASM_MACRO_TSVS__"
  pairwise_macro_blocks_tsvs: "__PAIRWISE_MACRO_TSVS__"
  pairwise_names: "__PAIRWISE_NAMES__"
  reference_name: "__REFNAME__"
  synteny_mode: "__SYNTENY_MODE__"
  chr_like_minlen: "__CHRLIKE__"
  out_prefix: "__OUTPREFIX__"
output:
  html_document:
    self_contained: true
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 2
    theme: flatly
---

```{css, echo=FALSE}
body {
  font-family: "Helvetica Neue", Helvetica, "Liberation Sans", Arial, sans-serif;
  font-size: 13px;
  max-width: 1300px;
  margin: auto;
}
h1.title { font-size: 24px; margin-bottom: 4px; }
h1 { font-size: 20px; border-bottom: 2px solid #2c3e50; padding-bottom: 4px; }
h2 { font-size: 16px; }
.gt_table { font-size: 11px !important; }
.gt_col_heading { font-size: 10px !important; font-weight: bold !important; }
.widget-container { margin: 16px 0; }
.note-box {
  background: #f8f9fa; border-left: 4px solid #6c757d;
  padding: 10px 14px; margin: 12px 0; font-size: 12px; color: #495057;
}
.section-intro { font-size: 13px; color: #555; margin-bottom: 12px; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, dpi = 150)

if (!requireNamespace("pacman", quietly = TRUE))
  install.packages("pacman", repos = "https://cloud.r-project.org")
library(pacman)
pacman::p_load(readr, dplyr, stringr, tidyr, tibble,
               ggplot2, ggnewscale, patchwork, grid,
               ggiraph, htmlwidgets, ggrepel,
               ggokabeito, colorspace,
               gt, gtExtras, scales,
               showtext, sysfonts, knitr)

# ---------------------------------------------------------------------------
# Font setup (shared across all plots)
# ---------------------------------------------------------------------------
.resolve_font <- function(family) {
  slug <- tolower(gsub(" ", "", family))
  path <- systemfonts::match_fonts(family)$path
  if (nzchar(path) && grepl(slug, tolower(gsub("[- ]", "", basename(path)))))
    return(path)
  NULL
}
base_family <- "sans"
for (.fam in c("Helvetica Neue", "Helvetica", "Liberation Sans")) {
  .r <- .resolve_font(.fam)
  if (!is.null(.r)) {
    sysfonts::font_add(.fam,
      regular    = .r,
      bold       = systemfonts::match_fonts(.fam, weight = "bold")$path,
      italic     = systemfonts::match_fonts(.fam, italic = TRUE)$path,
      bolditalic = systemfonts::match_fonts(.fam, italic = TRUE, weight = "bold")$path
    )
    base_family <- .fam
    break
  }
}
showtext_auto()

# ---------------------------------------------------------------------------
# Shared theming
# ---------------------------------------------------------------------------
base_font_pt <- 8
axis_title_margin_pt <- 4
axis_text_margin_pt <- 4

axis_theme <- theme(
  axis.title.x = element_text(size = base_font_pt, family = base_family,
                               margin = margin(t = axis_title_margin_pt)),
  axis.title.y = element_text(size = base_font_pt, family = base_family,
                               margin = margin(r = axis_title_margin_pt)),
  axis.text.x = element_text(size = base_font_pt, family = base_family,
                               margin = margin(t = axis_text_margin_pt)),
  axis.text.y = element_text(size = base_font_pt, family = base_family)
)

apply_legend_theme <- function(p, text_pt = base_font_pt, key_pt = base_font_pt,
                               tight = TRUE) {
  base <- theme(
    legend.title = element_text(size = text_pt, family = base_family),
    legend.text  = element_text(size = text_pt, family = base_family),
    legend.key.height = grid::unit(key_pt, "pt"),
    legend.key.width  = grid::unit(key_pt, "pt"),
    legend.direction = "horizontal",
    legend.box = "horizontal"
  )
  if (tight) {
    base <- base + theme(
      legend.box.margin = margin(0, 0, 0, 0),
      legend.margin = margin(0, 0, 0, 0),
      legend.spacing.y = grid::unit(0, "pt"),
      legend.spacing.x = grid::unit(2, "pt")
    )
  }
  p + base
}

# ---------------------------------------------------------------------------
# Okabe-Ito palette and classification colors
# ---------------------------------------------------------------------------
oi <- palette_okabe_ito()

classification_colors <- c(
  "chrom_assigned"     = oi[5],
  "chrom_unassigned"   = oi[2],
  "organelle_complete" = oi[3],
  "organelle_debris"   = oi[4],
  "rDNA"               = oi[1],
  "contaminant"        = oi[7],
  "chrom_debris"       = lighten(oi[5], amount = 0.5),
  "debris"             = oi[9],
  "unclassified"       = oi[8]
)

classification_labels <- c(
  "chrom_assigned"     = "Chromosome (assigned)",
  "chrom_unassigned"   = "Chromosome (unassigned)",
  "organelle_complete" = "Organelle (complete)",
  "organelle_debris"   = "Organelle (debris)",
  "rDNA"               = "rDNA",
  "contaminant"        = "Contaminant",
  "chrom_debris"       = "Chromosome debris",
  "debris"             = "Debris",
  "unclassified"       = "Unclassified"
)

classification_levels <- c(
  "chrom_assigned", "chrom_unassigned", "organelle_complete", "organelle_debris",
  "rDNA", "contaminant", "chrom_debris", "debris", "unclassified"
)

# ---------------------------------------------------------------------------
# Helpers
# ---------------------------------------------------------------------------
save_pdf <- function(path, plot, width = 7.2, height = 5, dpi = 300) {
  dev <- if (capabilities("cairo")) cairo_pdf else "pdf"
  invisible(ggsave(path, plot, width = width, height = height, dpi = dpi, device = dev))
}

girafe_styled <- function(ggobj, width_svg = 10, height_svg = 6.9) {
  g <- ggiraph::girafe(ggobj = ggobj, width_svg = width_svg, height_svg = height_svg)
  ggiraph::girafe_options(g,
    ggiraph::opts_tooltip(css = paste0(
      "font-family: ", base_family, "; font-size: 9pt; ",
      "background: white; padding: 6px; border: 1px solid #666;"
    )),
    ggiraph::opts_hover(css = "opacity: 0.8;"),
    ggiraph::opts_toolbar(saveaspng = FALSE)
  )
}

gt_theme <- function(gt_obj) {
  gt_obj %>%
    tab_options(
      table.font.size = px(11),
      column_labels.font.weight = "bold",
      column_labels.border.top.style = "none",
      column_labels.border.bottom.width = px(2),
      column_labels.border.bottom.color = "#333",
      table.border.top.style = "none",
      table.border.bottom.style = "none",
      table_body.border.bottom.color = "#999",
      data_row.padding = px(4),
      heading.border.bottom.style = "none"
    ) %>%
    opt_row_striping(row_striping = FALSE)
}

fmt_mb <- function(x) sprintf("%.1f", x / 1e6)
fmt_pct <- function(x) ifelse(is.na(x), "\u2014", sprintf("%.1f%%", x * 100))

# Assembly color palette (one color per assembly)
asm_pal <- c(oi[5], oi[1], oi[3], oi[6], oi[7], oi[4], oi[2], oi[8], oi[9])

# ---------------------------------------------------------------------------
# Load data
# ---------------------------------------------------------------------------
comp_df <- read_tsv(params$comparison_tsv, show_col_types = FALSE)
compl_df <- read_tsv(params$completeness_tsv, show_col_types = FALSE)
ref_lengths <- read_tsv(params$ref_lengths_tsv, show_col_types = FALSE,
                        col_types = cols(subgenome = col_character()))

n_asm <- nrow(comp_df)
asm_names <- comp_df$assembly
asm_colors <- setNames(asm_pal[seq_len(min(n_asm, length(asm_pal)))], asm_names)

# Load per-assembly summary TSVs
asm_tsv_paths <- strsplit(params$per_assembly_summary_tsvs, ";")[[1]]
asm_name_list <- strsplit(params$assembly_names, ";")[[1]]

per_asm_summaries <- list()
for (i in seq_along(asm_tsv_paths)) {
  p <- asm_tsv_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(assigned_subgenome = col_character()))
    df$assembly <- asm_name_list[i]
    per_asm_summaries[[i]] <- df
  }
}
all_summary_df <- if (length(per_asm_summaries) > 0) bind_rows(per_asm_summaries) else NULL
has_per_asm <- !is.null(all_summary_df) && nrow(all_summary_df) > 0

has_any_chrom <- any(comp_df$n_chrom_assigned > 0, na.rm = TRUE)
has_any_contam <- any(comp_df$n_contaminants > 0, na.rm = TRUE)
has_any_depth <- any(!is.na(comp_df$mean_chrom_depth) & comp_df$mean_chrom_depth > 0)

chr_like_minlen <- as.numeric(params$chr_like_minlen)

# PDF output paths
pdf_classbar <- paste0(params$out_prefix, ".classification_comparison.pdf")
pdf_heatmap  <- paste0(params$out_prefix, ".chromosome_completeness.pdf")
pdf_synteny  <- paste0(params$out_prefix, ".chromosome_synteny.pdf")

# ---------------------------------------------------------------------------
# Load per-assembly macro_blocks and pairwise macro_blocks for synteny plot
# ---------------------------------------------------------------------------
per_asm_macro_paths <- if (nzchar(params$per_assembly_macro_blocks_tsvs))
  strsplit(params$per_assembly_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_macro_paths <- if (nzchar(params$pairwise_macro_blocks_tsvs))
  strsplit(params$pairwise_macro_blocks_tsvs, ";")[[1]] else character(0)
pairwise_pair_names <- if (nzchar(params$pairwise_names))
  strsplit(params$pairwise_names, ";")[[1]] else character(0)

# Load per-assembly macro_blocks (ref→asm) and tag with assembly name
per_asm_macro_list <- list()
for (i in seq_along(per_asm_macro_paths)) {
  p <- per_asm_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0) {
      df$assembly <- asm_name_list[i]
      per_asm_macro_list[[i]] <- df
    }
  }
}
per_asm_macro_df <- if (length(per_asm_macro_list) > 0) bind_rows(per_asm_macro_list) else NULL

# Load pairwise macro_blocks (asm→asm)
pairwise_macro_list <- list()
for (i in seq_along(pairwise_macro_paths)) {
  p <- pairwise_macro_paths[i]
  if (nzchar(p) && file.exists(p)) {
    df <- read_tsv(p, show_col_types = FALSE,
                   col_types = cols(subgenome = col_character()))
    if (nrow(df) > 0 && i <= length(pairwise_pair_names)) {
      df$pair_name <- pairwise_pair_names[i]
      # Extract left and right assembly names from pair name
      # Split only on first occurrence of "_vs_" to handle assembly names
      # that might contain "_vs_" (unlikely but defensive)
      parts <- strsplit(pairwise_pair_names[i], "_vs_", fixed = TRUE)[[1]]
      if (length(parts) >= 2) {
        df$left_assembly <- parts[1]
        df$right_assembly <- paste(parts[-1], collapse = "_vs_")
      } else {
        warning(sprintf("Cannot parse pairwise pair name: %s", pairwise_pair_names[i]))
        next
      }
      pairwise_macro_list[[i]] <- df
    }
  }
}
pairwise_macro_df <- if (length(pairwise_macro_list) > 0) bind_rows(pairwise_macro_list) else NULL

has_synteny_data <- !is.null(per_asm_macro_df) && nrow(per_asm_macro_df) > 0 &&
                    params$synteny_mode == "nucleotide"
```

---

<p class="section-intro">
**Reference**: `r if (nchar(params$reference_name) > 0) params$reference_name else "\u2014"` &emsp;
**Synteny mode**: `r params$synteny_mode` &emsp;
**Assemblies**: `r n_asm` &emsp;
**Date**: `r Sys.Date()`
</p>

# 1. Assembly Overview {.tabset}

## Summary Table

```{r overview_table}
tbl <- comp_df %>%
  transmute(
    Assembly = assembly,
    Contigs  = total_contigs,
    `Size (Mb)` = round(total_bp / 1e6, 1),
    N50 = round(n50 / 1e6, 2),
    L50 = l50,
    `Largest (Mb)` = round(largest_contig / 1e6, 1),
    `Chr assigned` = n_chrom_assigned,
    `Chr unassigned` = n_chrom_unassigned,
    `Full-length` = n_full_length,
    `Both telo` = n_both_telomeres,
    `Mean ref cov` = ifelse(is.na(mean_ref_coverage), "\u2014",
                            sprintf("%.1f%%", mean_ref_coverage * 100)),
    Chimeric = n_chimeric,
    `Mean identity` = ifelse(is.na(mean_identity), "\u2014",
                             sprintf("%.2f%%", mean_identity * 100)),
    Contaminants = n_contaminants,
    rDNA = n_rdna,
    chrC = ifelse(chrC_found == "yes", "\u2713", "\u2014"),
    chrM = ifelse(chrM_found == "yes", "\u2713", "\u2014")
  )

tbl %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Assembly Comparison Summary") %>%
  cols_align(align = "right", columns = where(is.numeric)) %>%
  tab_spanner(label = "Contiguity", columns = c(Contigs, `Size (Mb)`, N50, L50, `Largest (Mb)`)) %>%
  tab_spanner(label = "Chromosomes", columns = c(`Chr assigned`, `Chr unassigned`, `Full-length`, `Both telo`, `Mean ref cov`, Chimeric, `Mean identity`)) %>%
  tab_spanner(label = "Other", columns = c(Contaminants, rDNA, chrC, chrM))
```

## Classification Distribution

```{r classification_bar, fig.width=10, fig.height=max(3, n_asm * 0.8 + 1.5)}
if (has_per_asm) {
  clf_data <- all_summary_df %>%
    group_by(assembly, classification) %>%
    summarise(total_bp = sum(length, na.rm = TRUE), n_contigs = n(), .groups = "drop") %>%
    mutate(
      total_mb = total_bp / 1e6,
      classification = factor(classification, levels = rev(classification_levels)),
      assembly = factor(assembly, levels = rev(asm_names))
    )

  p_clf <- ggplot(clf_data, aes(x = assembly, y = total_mb, fill = classification)) +
    geom_col_interactive(
      aes(tooltip = paste0(assembly, "\n",
                           classification_labels[as.character(classification)], "\n",
                           sprintf("%.1f Mb (%d contigs)", total_mb, n_contigs))),
      width = 0.7
    ) +
    scale_fill_manual(values = classification_colors, labels = classification_labels,
                      drop = FALSE, name = NULL) +
    coord_flip() +
    labs(x = NULL, y = "Assembly size (Mb)") +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(legend.position = "bottom",
          panel.grid.major.y = element_blank())
  p_clf <- apply_legend_theme(p_clf)

  save_pdf(pdf_classbar, p_clf,
           width = 7.2, height = max(3, n_asm * 0.8 + 1.5))
  girafe_styled(ggobj = p_clf,
                width_svg = 10,
                height_svg = max(3, n_asm * 0.8 + 1.5))
} else {
  cat("Per-assembly contig summary data not available.")
}
```

## Size Comparison

```{r size_comparison, fig.width=10, fig.height=max(3, n_asm * 0.6 + 1.5)}
size_data <- comp_df %>%
  transmute(
    assembly = factor(assembly, levels = rev(asm_names)),
    `Chrom assigned`  = chrom_assigned_bp / 1e6,
    `Organelle`       = organelle_bp / 1e6,
    `rDNA`            = rdna_bp / 1e6,
    `Contaminant`     = contaminant_bp / 1e6,
    `Debris`          = debris_bp / 1e6,
    `Unclassified`    = unclassified_bp / 1e6
  ) %>%
  pivot_longer(-assembly, names_to = "category", values_to = "mb") %>%
  mutate(category = factor(category,
    levels = c("Chrom assigned", "Organelle", "rDNA", "Contaminant", "Debris", "Unclassified")))

cat_colors <- c(
  "Chrom assigned" = oi[5], "Organelle" = oi[3], "rDNA" = oi[1],
  "Contaminant" = oi[7], "Debris" = oi[9], "Unclassified" = oi[8]
)

p_size <- ggplot(size_data, aes(x = assembly, y = mb, fill = category)) +
  geom_col_interactive(
    aes(tooltip = paste0(assembly, "\n", category, ": ", sprintf("%.1f Mb", mb))),
    position = "dodge", width = 0.7
  ) +
  scale_fill_manual(values = cat_colors, name = NULL) +
  coord_flip() +
  labs(x = NULL, y = "Size (Mb)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(legend.position = "bottom", panel.grid.major.y = element_blank())
p_size <- apply_legend_theme(p_size)

girafe_styled(ggobj = p_size,
              width_svg = 10,
              height_svg = max(3, n_asm * 0.6 + 1.5))
```


# 2. Chromosome Completeness {.tabset}

## Heatmap

```{r completeness_heatmap, fig.width=10, fig.height=max(4, nrow(ref_lengths) * 0.3 + 2)}
if (has_any_chrom) {
  # Sort reference chromosomes by length
  ref_order <- ref_lengths %>%
    arrange(desc(length)) %>%
    pull(ref_id)

  hm_data <- compl_df %>%
    mutate(
      ref_id = factor(ref_id, levels = rev(ref_order)),
      assembly = factor(assembly, levels = asm_names),
      best_ref_coverage = as.numeric(best_ref_coverage),
      coverage_pct = ifelse(is.na(best_ref_coverage), 0, best_ref_coverage * 100)
    )

  # Telomere/full-length status for overlay
  hm_data <- hm_data %>%
    mutate(
      status_label = case_when(
        is_full_length == "yes" & has_both_telomeres == "yes" ~ "\u2605",  # star
        is_full_length == "yes" ~ "\u25CF",  # filled circle
        n_contigs > 0 ~ "\u25CB",  # open circle
        TRUE ~ ""
      ),
      tooltip = paste0(
        ref_id, " \u2014 ", assembly, "\n",
        "Contigs: ", n_contigs, "\n",
        "Assigned bp: ", scales::comma(total_assigned_bp), "\n",
        "Best ref coverage: ", fmt_pct(best_ref_coverage), "\n",
        "Full-length: ", is_full_length, "\n",
        "Both telomeres: ", has_both_telomeres
      )
    )

  p_hm <- ggplot(hm_data, aes(x = assembly, y = ref_id, fill = coverage_pct)) +
    geom_tile_interactive(aes(tooltip = tooltip), color = "white", linewidth = 0.5) +
    geom_text(aes(label = status_label), size = 2.5, color = "white", fontface = "bold") +
    scale_fill_gradient2(
      low = "#f0f0f0", mid = "#74add1", high = "#1a9850",
      midpoint = 50, limits = c(0, 100),
      name = "Ref coverage (%)",
      na.value = "#f0f0f0"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = base_font_pt),
      legend.position = "bottom",
      panel.grid = element_blank()
    )
  p_hm <- apply_legend_theme(p_hm)

  save_pdf(pdf_heatmap, p_hm,
           width = max(4, n_asm * 0.8 + 2),
           height = max(4, length(ref_order) * 0.3 + 2))
  girafe_styled(ggobj = p_hm,
                width_svg = max(6, n_asm * 0.8 + 2),
                height_svg = max(4, length(ref_order) * 0.3 + 2))
} else {
  cat("No chromosome-assigned contigs found in any assembly.")
}
```

<div class="note-box">
**Legend**: &#9733; = full-length with both telomeres, &#9679; = full-length, &#9675; = assigned (not full-length), blank = no contigs assigned.
</div>

## Completeness Table

```{r completeness_table}
if (has_any_chrom) {
  tbl_data <- compl_df %>%
    mutate(best_ref_coverage = as.numeric(best_ref_coverage)) %>%
    select(ref_id, assembly, n_contigs, total_assigned_bp,
           best_ref_coverage, is_full_length, has_both_telomeres) %>%
    mutate(
      `Assigned (Mb)` = round(total_assigned_bp / 1e6, 1),
      `Ref cov (%)` = ifelse(is.na(best_ref_coverage), "\u2014",
                             sprintf("%.1f", best_ref_coverage * 100)),
      `Full-length` = ifelse(is_full_length == "yes", "\u2713", ""),
      `Both telo` = ifelse(has_both_telomeres == "yes", "\u2713", "")
    ) %>%
    select(ref_id, assembly, n_contigs, `Assigned (Mb)`, `Ref cov (%)`,
           `Full-length`, `Both telo`)

  tbl_data %>%
    gt(groupname_col = "ref_id") %>%
    gt_theme() %>%
    tab_header(title = "Chromosome Completeness by Reference") %>%
    cols_label(
      assembly = "Assembly", n_contigs = "Contigs",
      `Assigned (Mb)` = "Assigned (Mb)",
      `Ref cov (%)` = "Ref cov (%)",
      `Full-length` = "Full-length",
      `Both telo` = "Both telo"
    ) %>%
    cols_align(align = "right", columns = c(n_contigs, `Assigned (Mb)`, `Ref cov (%)`)) %>%
    cols_align(align = "center", columns = c(`Full-length`, `Both telo`))
} else {
  cat("No chromosome-assigned contigs found in any assembly.")
}
```


`r if (has_any_chrom && has_per_asm) "# 3. Quality Comparison {.tabset}" else ""`

```{r quality_section, eval=(has_any_chrom && has_per_asm)}
chrom_df <- all_summary_df %>%
  filter(classification == "chrom_assigned")

has_identity <- "seq_identity_vs_ref" %in% names(chrom_df) &&
                !all(is.na(chrom_df$seq_identity_vs_ref))
has_collin   <- "collinearity_score" %in% names(chrom_df) &&
                !all(is.na(chrom_df$collinearity_score))
has_gc_dev   <- "gc_deviation" %in% names(chrom_df) &&
                !all(is.na(chrom_df$gc_deviation))
```

```{r quality_identity, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), results='asis'}
cat("## Sequence Identity\n\n")
```

```{r quality_identity_plot, eval=(has_any_chrom && has_per_asm && exists("has_identity") && has_identity), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
p_ident <- ggplot(chrom_df, aes(x = assembly, y = seq_identity_vs_ref * 100,
                                fill = assembly)) +
  geom_violin(alpha = 0.6, width = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Sequence identity vs reference (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_ident, height_svg = max(3, n_asm * 0.8 + 1))
```

```{r quality_collinearity, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), results='asis'}
cat("## Collinearity\n\n")
```

```{r quality_collinearity_plot, eval=(has_any_chrom && has_per_asm && exists("has_collin") && has_collin), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
p_collin <- ggplot(chrom_df, aes(x = assembly, y = collinearity_score * 100,
                                  fill = assembly)) +
  geom_violin(alpha = 0.6, width = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Collinearity (%)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_collin, height_svg = max(3, n_asm * 0.8 + 1))
```

```{r quality_gc, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), results='asis'}
cat("## GC Deviation\n\n")
```

```{r quality_gc_plot, eval=(has_any_chrom && has_per_asm && exists("has_gc_dev") && has_gc_dev), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
p_gc <- ggplot(chrom_df, aes(x = assembly, y = abs(gc_deviation), fill = assembly)) +
  geom_violin(alpha = 0.6, width = 0.7) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "|GC deviation| (std devs from reference)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_gc, height_svg = max(3, n_asm * 0.8 + 1))
```


`r if (has_any_contam) "# 4. Contamination Comparison" else ""`

```{r contamination, eval=has_any_contam, fig.width=10, fig.height=max(3, n_asm * 0.6 + 1.5)}
contam_data <- comp_df %>%
  filter(n_contaminants > 0) %>%
  mutate(assembly = factor(assembly, levels = rev(asm_names)))

p_contam <- ggplot(contam_data, aes(x = assembly, fill = assembly)) +
  geom_col_interactive(
    aes(y = contaminant_bp / 1e6,
        tooltip = paste0(assembly, "\n",
                         n_contaminants, " contaminants\n",
                         sprintf("%.2f Mb", contaminant_bp / 1e6), "\n",
                         n_contaminant_species, " species")),
    width = 0.6
  ) +
  scale_fill_manual(values = asm_colors, guide = "none") +
  coord_flip() +
  labs(x = NULL, y = "Contaminant sequence (Mb)") +
  theme_minimal(base_size = base_font_pt, base_family = base_family) +
  axis_theme +
  theme(panel.grid.major.y = element_blank())

girafe_styled(ggobj = p_contam,
              height_svg = max(3, n_asm * 0.6 + 1.5))
```

```{r contamination_table, eval=has_any_contam}
comp_df %>%
  filter(n_contaminants > 0) %>%
  transmute(
    Assembly = assembly,
    `N contaminants` = n_contaminants,
    `Contam bp (Mb)` = round(contaminant_bp / 1e6, 2),
    `Unique species` = n_contaminant_species
  ) %>%
  gt() %>%
  gt_theme() %>%
  tab_header(title = "Contamination Summary") %>%
  cols_align(align = "right", columns = where(is.numeric))
```


`r if (has_any_depth) "# 5. Depth Comparison" else ""`

```{r depth_section, eval=(has_any_depth && has_per_asm), fig.width=10, fig.height=max(3, n_asm * 0.8 + 1)}
depth_df <- all_summary_df %>%
  filter(classification == "chrom_assigned", !is.na(depth_mean))

if (nrow(depth_df) > 0) {
  p_depth <- ggplot(depth_df, aes(x = assembly, y = depth_mean, fill = assembly)) +
    geom_violin(alpha = 0.6, width = 0.7) +
    geom_boxplot(width = 0.15, outlier.size = 0.5, fill = "white", alpha = 0.7) +
    scale_fill_manual(values = asm_colors, guide = "none") +
    coord_flip() +
    labs(x = NULL, y = "Mean read depth (chromosome contigs)") +
    theme_minimal(base_size = base_font_pt, base_family = base_family) +
    axis_theme +
    theme(panel.grid.major.y = element_blank())

  girafe_styled(ggobj = p_depth, height_svg = max(3, n_asm * 0.8 + 1))
} else {
  cat("No depth data available for chromosome-assigned contigs.")
}
```

`r if (has_synteny_data) "# 6. Chromosome Synteny {.tabset}" else ""`

```{r synteny_setup, eval=has_synteny_data, include=FALSE}
# =========================================================================
# Section 6: Chromosome Synteny Riparian Plot
#
# Horizontal-bar layout: one row per reference chromosome.
# Columns: Reference, Asm1, Asm2, ... with S-curve ribbons between
# adjacent columns showing syntenic blocks.
# =========================================================================

# --- Build assembly contig → assigned_ref_id lookup from contig_summary ---
asm_contig_lookup <- list()
for (i in seq_along(asm_name_list)) {
  nm <- asm_name_list[i]
  df <- per_asm_summaries[[i]]
  if (!is.null(df)) {
    lookup <- df %>%
      filter(classification == "chrom_assigned") %>%
      select(contig, assigned_ref_id, reversed, length) %>%
      mutate(assembly = nm)
    asm_contig_lookup[[nm]] <- lookup
  }
}
asm_lookup_df <- bind_rows(asm_contig_lookup)

# --- Reference chromosome order (by length descending) ---
ref_chr_order <- ref_lengths %>%
  arrange(desc(length)) %>%
  pull(ref_id)

# Column order: Reference, then assemblies in order
col_order <- c("Reference", asm_name_list)
n_cols <- length(col_order)

# --- Build contig records per column per chromosome ---
# For reference: one bar per ref_id with ref length
# For assemblies: join with contig_summary to get assigned contigs

# Compute bar data: for each column and each chromosome row,
# what contig(s) and lengths are present?
bar_data <- list()

# Reference bars
for (rid in ref_chr_order) {
  rlen <- ref_lengths$length[ref_lengths$ref_id == rid]
  if (length(rlen) > 0) {
    bar_data[[length(bar_data) + 1]] <- tibble(
      column = "Reference", ref_id = rid,
      contig = rid, contig_len = rlen[1], reversed = FALSE
    )
  }
}

# Assembly bars
for (nm in asm_name_list) {
  lk <- asm_lookup_df %>% filter(assembly == nm)
  for (rid in ref_chr_order) {
    contigs_for_ref <- lk %>% filter(assigned_ref_id == rid)
    if (nrow(contigs_for_ref) > 0) {
      # Use the longest contig assigned to this ref chromosome
      best <- contigs_for_ref %>% arrange(desc(length)) %>% slice(1)
      bar_data[[length(bar_data) + 1]] <- tibble(
        column = nm, ref_id = rid,
        contig = best$contig, contig_len = best$length,
        reversed = as.logical(best$reversed)
      )
    }
  }
}
bar_df <- bind_rows(bar_data)
bar_df$column <- factor(bar_df$column, levels = col_order)

# --- Layout geometry ---
# Each column has an x-center; bars extend left/right proportional to Mb
# Gap zones between columns hold the ribbons

# Compute maximum chromosome length across all columns for shared scale
max_bar_len <- max(bar_df$contig_len, na.rm = TRUE)

# Column widths (proportional to max chromosome length in that column)
col_max_lens <- bar_df %>%
  group_by(column) %>%
  summarise(max_len = max(contig_len, na.rm = TRUE), .groups = "drop") %>%
  arrange(factor(column, levels = col_order))

# Use shared Mb scale: bar width in x-units = contig_len / scale_factor
# Set total plot width budget (arbitrary units)
total_bar_width <- 80  # total x-units for all bars
gap_width <- 8         # x-units per gap

# Scale: longest bar maps to proportional share of total_bar_width
scale_factor <- max_bar_len / (total_bar_width / n_cols)

# Compute column x-centers
col_x <- numeric(n_cols)
col_x[1] <- col_max_lens$max_len[1] / scale_factor / 2
for (i in 2:n_cols) {
  prev_half <- col_max_lens$max_len[i - 1] / scale_factor / 2
  curr_half <- col_max_lens$max_len[i] / scale_factor / 2
  col_x[i] <- col_x[i - 1] + prev_half + gap_width + curr_half
}
col_centers <- setNames(col_x, col_order)

# Gap zones (for ribbon S-curves)
gap_zones <- list()
for (i in 1:(n_cols - 1)) {
  left_col <- col_order[i]
  right_col <- col_order[i + 1]
  left_right_edge <- col_centers[left_col] + col_max_lens$max_len[i] / scale_factor / 2
  right_left_edge <- col_centers[right_col] - col_max_lens$max_len[i + 1] / scale_factor / 2
  gap_zones[[i]] <- c(left_right_edge, right_left_edge)
}

# Row heights proportional to reference chromosome length
# y from 0 (top, chr1) downward
row_gap <- 0.3
row_heights <- ref_lengths$length[match(ref_chr_order, ref_lengths$ref_id)] / max_bar_len * 10
names(row_heights) <- ref_chr_order

row_y_top <- numeric(length(ref_chr_order))
names(row_y_top) <- ref_chr_order
cum_y <- 0
for (i in seq_along(ref_chr_order)) {
  row_y_top[i] <- cum_y
  cum_y <- cum_y + row_heights[i] + row_gap
}
total_y <- cum_y - row_gap

# --- Build chromosome bar rectangles ---
bar_rects <- bar_df %>%
  mutate(
    x_center = col_centers[as.character(column)],
    half_w = contig_len / scale_factor / 2,
    xmin = x_center - half_w,
    xmax = x_center + half_w,
    ymin = row_y_top[ref_id],
    ymax = row_y_top[ref_id] + row_heights[ref_id]
  )

# --- Chromosome color palette ---
n_chroms <- length(ref_chr_order)
if (n_chroms <= 9) {
  chrom_pal <- setNames(oi[seq_len(n_chroms)], ref_chr_order)
} else {
  chrom_pal <- setNames(
    colorRampPalette(c(oi[1:9]))(n_chroms),
    ref_chr_order
  )
}

# --- Build ribbon polygons ---
# Helper: map genomic position to y within a chromosome row
pos_to_y <- function(pos, seq_len, y_top, height) {
  y_top + (pos / seq_len) * height
}

# Helper: make S-curve polygon between two vertical extents across a gap
make_ribbon <- function(gap_left, gap_right,
                         left_y_top, left_y_bot,
                         right_y_top, right_y_bot,
                         fill_col, alpha_val, tooltip_text,
                         n_pts = 20) {
  t <- seq(0, 1, length.out = n_pts)
  s <- 0.5 - 0.5 * cos(pi * t)
  x_fwd <- gap_left + t * (gap_right - gap_left)
  top_fwd <- left_y_top + s * (right_y_top - left_y_top)
  bot_fwd <- left_y_bot + s * (right_y_bot - left_y_bot)
  tibble(
    x = c(x_fwd, rev(x_fwd)),
    y = c(top_fwd, rev(bot_fwd)),
    fill = fill_col,
    alpha = alpha_val,
    tooltip = tooltip_text
  )
}

# Build ref→asm ribbons (gap index 1: Reference → Asm1)
ribbon_polys <- list()
ribbon_idx <- 0

for (gap_i in seq_len(n_cols - 1)) {
  left_col_name <- col_order[gap_i]
  right_col_name <- col_order[gap_i + 1]
  gz <- gap_zones[[gap_i]]

  if (gap_i == 1) {
    # Reference → first assembly: use per-assembly macro_blocks
    macro_src <- per_asm_macro_df %>%
      filter(assembly == right_col_name)
  } else {
    # Assembly → Assembly: use pairwise macro_blocks
    if (!is.null(pairwise_macro_df)) {
      macro_src <- pairwise_macro_df %>%
        filter(left_assembly == left_col_name, right_assembly == right_col_name)
    } else {
      macro_src <- NULL
    }
  }

  if (is.null(macro_src) || nrow(macro_src) == 0) next

  # For ref→asm: ref_id is the reference chromosome, contig is the assembly contig
  # For asm→asm: ref_id is left assembly contig name, contig is right assembly contig name
  # We need to map both to chromosome rows via assigned_ref_id

  for (row_i in seq_along(ref_chr_order)) {
    rid <- ref_chr_order[row_i]
    rh <- row_heights[rid]
    ry <- row_y_top[rid]

    # Get left column bar for this chromosome
    left_bar <- bar_rects %>%
      filter(column == left_col_name, ref_id == rid)
    right_bar <- bar_rects %>%
      filter(column == right_col_name, ref_id == rid)

    if (nrow(left_bar) == 0 || nrow(right_bar) == 0) next

    left_len <- left_bar$contig_len[1]
    left_contig <- left_bar$contig[1]
    left_reversed <- left_bar$reversed[1]
    right_len <- right_bar$contig_len[1]
    right_contig <- right_bar$contig[1]
    right_reversed <- right_bar$reversed[1]

    if (gap_i == 1) {
      # ref→asm: ref_id in macro_blocks = reference chromosome ID
      blocks <- macro_src %>%
        filter(ref_id == rid, contig == right_contig)
    } else {
      # asm→asm: ref_id in pairwise = left assembly contig name
      blocks <- macro_src %>%
        filter(ref_id == left_contig, contig == right_contig)
    }

    if (nrow(blocks) == 0) next

    # Filter small blocks
    min_span <- max(50000, max_bar_len * 0.001)  # at least 50kb or 0.1% of max
    blocks <- blocks %>% filter(qspan_bp >= min_span)
    if (nrow(blocks) == 0) next

    for (b in seq_len(nrow(blocks))) {
      blk <- blocks[b, ]

      # Left side coordinates (ref_start/ref_end on left column)
      if (gap_i == 1) {
        # ref→asm: ref_start/ref_end are reference coordinates
        l_start <- as.numeric(blk$ref_start)
        l_end   <- as.numeric(blk$ref_end)
      } else {
        # asm→asm: ref_start/ref_end are left assembly coordinates
        l_start <- as.numeric(blk$ref_start)
        l_end   <- as.numeric(blk$ref_end)
        # If left contig is reversed relative to reference, flip coordinates
        if (left_reversed) {
          l_start_new <- left_len - l_end
          l_end_new   <- left_len - l_start
          l_start <- l_start_new
          l_end   <- l_end_new
        }
      }

      # Right side coordinates (qstart/qend on right column)
      r_start <- as.numeric(blk$qstart)
      r_end   <- as.numeric(blk$qend)
      # If right contig is reversed relative to reference, flip
      if (right_reversed) {
        r_start_new <- right_len - r_end
        r_end_new   <- right_len - r_start
        r_start <- r_start_new
        r_end   <- r_end_new
      }

      # Handle negative coordinates from flipping
      l_start <- max(0, l_start)
      l_end   <- min(left_len, l_end)
      r_start <- max(0, r_start)
      r_end   <- min(right_len, r_end)

      if (l_end <= l_start || r_end <= r_start) next

      # Map to y coordinates within the row
      ly_top <- pos_to_y(l_start, left_len, ry, rh)
      ly_bot <- pos_to_y(l_end, left_len, ry, rh)
      ry_top <- pos_to_y(r_start, right_len, ry, rh)
      ry_bot <- pos_to_y(r_end, right_len, ry, rh)

      # Color by reference chromosome; alpha by identity
      ident_val <- as.numeric(blk$identity)
      alpha_ribbon <- 0.2 + 0.5 * min(1, ident_val)

      # Determine if this block indicates an inversion (strand = "-")
      strand_val <- blk$strand
      if (strand_val == "-") {
        # Swap top/bottom on right side to show crossing
        tmp <- ry_top
        ry_top <- ry_bot
        ry_bot <- tmp
      }

      # Check if this is a cross-chromosome block
      # (ref_id for pairwise may map to a different chromosome)
      fill_chr <- rid
      tooltip <- sprintf(
        "%s → %s\n%s: %.2f Mb\nIdentity: %.1f%%\nStrand: %s",
        left_col_name, right_col_name, rid,
        as.numeric(blk$union_bp) / 1e6,
        ident_val * 100, strand_val
      )

      ribbon_idx <- ribbon_idx + 1
      ribbon_polys[[ribbon_idx]] <- make_ribbon(
        gz[1], gz[2], ly_top, ly_bot, ry_top, ry_bot,
        fill_col = fill_chr, alpha_val = alpha_ribbon,
        tooltip_text = tooltip
      ) %>% mutate(group_id = ribbon_idx)
    }
  }
}

has_ribbons <- length(ribbon_polys) > 0
if (has_ribbons) {
  all_ribbons <- bind_rows(ribbon_polys)
} else {
  all_ribbons <- NULL
}
```

```{r synteny_description, eval=has_synteny_data, results='asis'}
cat('<p class="section-intro">\n')
cat("Chromosome synteny between the reference and each assembly (nucleotide mode). ")
cat("Ribbons connect syntenic macro-blocks between adjacent columns. ")
cat("Color indicates the reference chromosome; opacity reflects alignment identity. ")
cat("Crossed ribbons indicate inversions.\n")
cat("</p>\n")
```

## Riparian Plot

```{r synteny_riparian, eval=(has_synteny_data && exists("has_ribbons") && has_ribbons), fig.width=10, fig.height=max(5, total_y * 0.45 + 2)}
# --- Compose the plot ---

# Base plot with coordinate system
p_syn <- ggplot() +
  coord_cartesian(
    xlim = c(-1, max(col_x) + col_max_lens$max_len[n_cols] / scale_factor / 2 + 1),
    ylim = c(total_y + 0.5, -0.5),  # reversed y: chr1 at top
    expand = FALSE
  )

# Draw ribbons first (behind bars)
p_syn <- p_syn +
  geom_polygon_interactive(
    data = all_ribbons,
    aes(x = x, y = y, group = group_id, fill = fill,
        alpha = alpha, tooltip = tooltip),
    show.legend = FALSE
  ) +
  scale_fill_manual(values = chrom_pal, guide = "none") +
  scale_alpha_identity()

# Draw chromosome bars
bar_rects_plot <- bar_rects %>%
  mutate(
    fill_col = ifelse(column == "Reference", "grey70", "grey85"),
    tooltip = paste0(column, "\n", contig, "\n",
                     sprintf("%.2f Mb", contig_len / 1e6))
  )

p_syn <- p_syn +
  ggnewscale::new_scale_fill() +
  geom_rect_interactive(
    data = bar_rects_plot,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
        fill = fill_col, tooltip = tooltip),
    color = "grey40", linewidth = 0.3
  ) +
  scale_fill_identity()

# Column labels at top
col_label_df <- tibble(
  x = col_x,
  label = col_order
)
p_syn <- p_syn +
  geom_text(data = col_label_df,
            aes(x = x, y = -0.3, label = label),
            size = 2.5, fontface = "bold", family = base_family, vjust = 1)

# Chromosome labels on the left
chr_label_df <- tibble(
  ref_id = ref_chr_order,
  y_mid = row_y_top[ref_chr_order] + row_heights[ref_chr_order] / 2
)
p_syn <- p_syn +
  geom_text(data = chr_label_df,
            aes(x = -0.5, y = y_mid, label = ref_id),
            size = 2, hjust = 1, family = base_family)

# Theme
p_syn <- p_syn +
  theme_void(base_size = base_font_pt, base_family = base_family) +
  theme(plot.margin = margin(5, 5, 5, 40))

# Dynamic sizing
plot_h <- max(5, total_y * 0.45 + 2)
plot_w <- min(10, max(7.2, n_cols * 2.5 + 2))

save_pdf(pdf_synteny, p_syn, width = 7.2, height = min(10, plot_h))
girafe_styled(ggobj = p_syn, width_svg = plot_w, height_svg = plot_h)
```

```{r synteny_empty, eval=(has_synteny_data && !(exists("has_ribbons") && has_ribbons))}
cat("Synteny data available but no ribbon connections could be built. ",
    "This may indicate that assemblies have very different chromosome structures.")
```

```{r synteny_not_nucleotide, eval=(!has_synteny_data && params$synteny_mode != "nucleotide"), results='asis'}
# Don't show section header if not nucleotide mode — it's hidden by the conditional above
```

---

<div class="note-box">
Generated by **final_finalizer** multi-assembly comparison &mdash; `r Sys.Date()`
</div>
